---
output:
  #bookdown::html_document2: default
  #bookdown::word_document2: default
  bookdown::pdf_document2:
    template: templates/brief_template.tex
    citation_package: biblatex
bib-humanities: true
documentclass: book
bibliography: [bibliography/references.bib]
---

```{r , include=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
```


# Eutrophication and greenhouse gases in an urbanized tropical estuary{#cites-and-refs} 
\chaptermark{Eutrophication and greenhouse gases}
\minitoc <!-- this will include a mini table of contents-->


Urbanized tropical estuaries are known as hot spots of eutrophication and greenhouse gases (GHGs) in emerging countries. This chapter presents the nutrients, organic carbon, phytoplankton, and greenhouse gases concentrations in the Saigon River Estuary and compares them with other estuaries in different climates. Several internal processes (e.g., nitrification, denitrification, aerobic degradation of organic carbon) are interpreted to identify the main processes affecting water quality and GHGs emissions at urbanized estuaries.

There was a clear separation of nutrients, organic carbon, GHGs, phytoplankton abundance between the urban area and the remaining area of Saigon River because of polluted urban emissions of Ho Chi Minh Megacity (HCMC). The major phytoplankton were diatoms (e.g., *Cyclotella* cf. *meneghiniana, Leptocylindrus danicus*), cyanobacteria (e.g., *Microcystis* spp). The dominance of potentially toxic cyanobacteria in the Saigon River possess health risk to local residents, especially upon the increasing temperature context and nutrient loading into the river in the next decades. Phytoplankton densities in dry season were about 100 times higher than in the rainy season. Calculation of the eutrophication index shows that 30-50% of the Saigon River (based on river length) was in the eutrophic state and 10% was in the hypertrophic state, mainly in the urban section. While the concentrations of nutrients did not differ significantly between the dry and wet seasons, the concentrations of GHGs ($\mathrm{N_2O}$, $\mathrm{CH_4}$) and phytoplankton densities were markedly different.

High eutrophication and phytoplankton abundance may enhance carbon dioxide ($\mathrm{CO_2}$) consumption by photosynthesis in the water column. However, the GHGs results at this urbanized estuary were always in supersaturation state. The average concentrations of $\mathrm{CO_2}$, $\mathrm{CH_4}$ and $\mathrm{N_2O}$ at Saigon River in 2019-2020 were 3174 $\pm$ 1725 $\mu$gC-$\mathrm{CO_2}$ $L^{-1}$, 5.9 $\pm$ 16.8 $\mu$gC-$\mathrm{CH_4}$ $L^{-1}$ and 3.0 $\pm$ 4.8 $\mu$gN-$\mathrm{N_2O}$ $L^{-1}$, respectively. Their concentrations were 13 – 18 times, 52 – 332 times, and 9 – 37 times higher than the global mean concentrations of GHGs, respectively. The increase in eutrophication status along the dense urban area was linearly correlated with increased GHGs concentrations. Both nitrification and denitrification resulted in elevated $\mathrm{N_2O}$ concentrations in this urban area of the estuary. The high concentration of $\mathrm{CO_2}$ was contributed by the high concentration of organic carbon and mineralization process. GHGs fluxes at Saigon River Estuary were comparable to other urbanized estuaries regardless of climatic condition. Control of eutrophication in urbanized estuaries through the implantation of efficient wastewater treatment facilities will effectively mitigate the global warming potential caused by estuarine emissions.

Chapter 3 uses the information presented in two articles:

Nguyen AT, Dao TS, Strady E, Nguyen TTN, Aimé J, Gratiot N, Némery J. (2021). Phytoplankton characterization in a tropical tidal river impacted by a megacity: the case of the Saigon River (Southern Vietnam). Environ Sci Pollut Res Int. <https://doi.org/10.1007/s11356-021-15850-x>

An Truong Nguyen, Julien Némery, Nicolas Gratiot, Thanh-Son Dao, Thi Minh Tam Le, Christine Baduel, Josette Garnier (*in press*). Does eutrophication enhance greenhouse gas emissions in urbanized tropical estuaries? Environmental Pollution journal. <https://doi.org/10.1016/j.envpol.2022.119105>

## Introduction {.unnumbered}

Saigon River Estuary is an urbanized tropical estuary in Southeast Asia. This river receives a large amount of untreated wastewater from Ho Chi Minh megacity (HCMC), resulting in high concentrations of nutrients, organic matter, chlorophyll-a, and oxygen deficiency in the water column. Several studies have revealed that this estuary was experiencing eutrophication in the urban area [@Nguyen2019; @Strady2017]. Eutrophication can significantly affect GHGs emissions [@Li2021; @DelSontro2019; @Dinauer2017]. This chapter presents the results of water quality monitoring (Figure \@ref(fig:C3-CARE-sampling) for sampling locations) in 5 years 2015 at Saigon River Estuary (Part 1), spatio-temporal variation of phytoplankton (Part 2) and interaction between GHGs and eutrophication (Part 3). Descriptions of the monitoring program, analysis methods, and data interpretation are detailed in Chapter 2. The contents presented below focus on results and discussion.

           
```{r C3-CARE-sampling, echo=FALSE, out.width='75%', fig.cap='Sampling sites of bi-weekly monitoring and longitudinal profiles', fig.align='center'}
knitr::include_graphics("figures/C3-CARE-sampling.pdf")
```


## Part 1: Spatio-temporal of environmental parameters 

### Physico-chemical parameters

Figure \@ref(fig:C3-physical) presents the monitoring data of physico-chemical parameters (temperature, pH, TSS and salinity). The water temperature at Saigon River ranges from about 28$^{\circ}$C to 33$^{\circ}$C, the highest temperature is usually in March and April of the dry seasons and the lowest at the end of the rainy season or the beginning of the dry season (December) (Figure \@ref(fig:C3-physical)a). While the temperature did not differ between the three monitoring stations, the pH concentration was significantly different between Phu Cuong station (PC) and Bach Dang station (BD), Binh Khanh station (BK) with lower pH concentration at PC. The lower pH at PC shows that this area is slightly acidic (pH \< 7). There was no difference in pH between BD and BK stations with a range from 6 to 7.5 and no difference between rainy and dry seasons. The pH levels at PC, BD and BK are 6.5, 6.8 and 6.8, respectively.

TSS concentrations ranged from about 20 mgTSS$L^{-1}$ to more than 300 mgTSS$L^{-1}$ (Figure \@ref(fig:C3-physical)c). The evolution of TSS at all three stations is usually not seasonal, however in general TSS concentration at PC - upstream is the lowest. The mean TSS concentrations at PC, BD and BK were 32, 95 and 75 mgTSS$L^{-1}$, respectively.

```{r C3-physical, echo=FALSE, out.width='100%', fig.cap='Temporal variations of (a) temperature, (b) pH, (c) TSS and (d) salinity at three sampling stations from July 2015 to December 2020. PC, BD and BK are Phu Cuong station at upstream, Bach Dang station at urban section and Binh Khanh station at downstream of Saigon River Estuary', fig.align='center'}
knitr::include_graphics("figures/C3-physical.pdf")
```

### Dissolved oxygen, nutrients and organic carbon

Monitoring data of DO, nutrients ($\mathrm{NH_4^{+}}$, $\mathrm{NO_{3}^{-}}$, $\mathrm{PO_{4}^{3-}}$, $\mathrm{DSi}$), organic carbon at three monitoring stations in five years along the Saigon River are shown in Figure \@ref(fig:C3-chemical). The evolution of DO, nutrient and TOC concentrations at all three monitoring stations fluctuated strongly and did not have a seasonal trend. However, there was a clear difference in concentration between monitoring stations.

```{r C3-chemical, echo=FALSE, out.width='100%', fig.cap='Temporal variations of (a) DO, (b) $\\mathrm{NH_4^{+}}$, (c) $\\mathrm{NO_{3}^{-}}$, (d) $\\mathrm{PO_{4}^{3-}}$, (e) $\\mathrm{DSi}$ and (f) TOC at three sampling stations from July 2015 to December 2020.', fig.align='center'}
knitr::include_graphics("figures/C3-chemical.pdf")
```

The lowest oxygen concentration was at the BD - urban section station, with an average value of 1.5 $\pm$ 0.8 mg$L^{-1}$ (Figure \@ref(fig:C3-chemical)a). DO concentrations can range from less than 1 mg$L^{-1}$ to about 4 mg$L^{-1}$. DO concentrations at PC and BK were 2.1 $\pm$ 1.0 and 3.3 $\pm$ 1.1 mg$L^{-1}$, respectively. While DO was not significantly different between PC and BK, $\mathrm{NH_4^{+}}$ concentration depicted this distinct phenomenon. The $\mathrm{NH_4^{+}}$ concentration at BK was only about 0.1 $\pm$ 0.1 $mgNL^{-1}$, while the $\mathrm{NH_4^{+}}$ concentrations at PC and BD were 0.5 $\pm$ 0.2 and 0.7 $\pm$ 0.5 $mgNL^{-1}$, respectively. During the 5 years of monitoring, the concentration of $\mathrm{NH_4^{+}}$ at BD exceeded 2 $mgNL^{-1}$ four times, the remaining time periods were lower than 1 $mgNL^{-1}$ (Figure \@ref(fig:C3-chemical)b).

While BK- downstream has the lowest concentration of $\mathrm{NH_4^{+}}$, the lowest concentration of $\mathrm{NO_{3}^{-}}$ was observed at PC - upstream, reaching about 0.2 $\pm$ 0.2 $mgNL^{-1}$. The $\mathrm{NO_{3}^{-}}$ concentrations at BD and BK were quite similar at 0.7 $\pm$ 0.3 and 0.6 $\pm$ 0.3 mg$L^{-1}$, respectively (Figure \@ref(fig:C3-chemical)c). The concentration of $\mathrm{PO_{4}^{3-}}$ is a special pollutant in the Saigon River, its average concentration is only about 0.05 mgP$L^{-1}$ with the concentration at BD being 0.07 $\pm$ 0.05. The concentration of $\mathrm{PO_{4}^{3-}}$ at PC and BK is almost not different, reaching about 0.04 $\pm$ 0.02 mgP$L^{-1}$. In the dry season of 2020, the concentration of $\mathrm{PO_{4}^{3-}}$ decreased markedly, almost not detected in the analysis, while the concentrations of $\mathrm{PO_{4}^{3-}}$ in were highest in rainy season of 2019 and 2020, the, reaching about 0.17 mgP for three monitoring stations (Figure \@ref(fig:C3-chemical)d).

$\mathrm{DSi}$ concentration in three years from 2017 to 2019 was very low, only 1mgSi$L^{-1}$, however from the end of 2019 $\mathrm{DSi}$ concentration gradually increased from 1 to over 9 mgSi$L^{-1}$ in 2020. In the rainy season of 2020, the concentration of $\mathrm{DSi}$ decreased rapidly which was similar to $\mathrm{PO_{4}^{3-}}$. TOC monitoring results showed that TOC concentrations at PC, BD and BK were 4.8 $\pm$ 1.5, 6.3 $\pm$ 2.6 and 3.9 $\pm$ 2.0 mgC$L^{-1}$, respectively. TOC concentrations did not differ by seasons and did not tend to increase or decrease as strongly as found in $\mathrm{DSi}$ and $\mathrm{PO_{4}^{3-}}$.

### Eutrophication status

The Chl-a monitoring for about 5 years from 2015 to 2020 shows that differences between the two seasons, with higher Chl-a concentrations in the dry seasons (Figure \@ref(fig:C3-PC) and Figure \@ref(fig:C3-HA)). Besides, Chl-a was the parameter with the most different results among the three stations along the Saigon River. The average Chl-a concentrations at PC, BD and BK were 3.3 $\pm$ 6.0, 21.2 $\pm$ 26.4 and 1.0 $\pm$ 1.0 $\mu g$Chl-a$L^{-1}$, respectively. Besides, a monitoring station at Dongnai River, a major tributary of Saigon River was also reported with a concentration of 3.7 $\pm$ 3.9 $\mu g$Chl-a$L^{-1}$. Trophic status was also calculated based on the concentrations of Chl-a, DO, DIN ($\mathrm{NH_4^{+}}$ + $\mathrm{NO_{3}^{-}}$) and $\mathrm{PO_{4}^{3-}}$ to assess the risk of eutrophication in the Saigon River.

```{r C3-PC, echo=FALSE, out.width='100%', fig.cap='Evolution of Chl-a and trophic index in the Saigon River from 2015 to 2020, at (a) Phu Cuong station and (b) Bach Dang station. Trophic index indicates several states of eutrophication, including ultraoligotrophic (0-2), oligotrophic (2-4), mesotrophic (4-6), eutrophic (6-8), and hypertrophic (8-10) states, respectively.', fig.align='center'}
knitr::include_graphics("figures/C3-PC.pdf")
```

Figure \@ref(fig:C3-PC) depicts Chl-a concentrations and trophic index at Phu Cuong (PC) and Bach Dang (BD). Chl-a concentration in PC only once in 5 years of survey exceeded 30 $\mu g$Chl-a$L^{-1}$ in the dry season 2019. Meanwhile, Chl-a concentration in BD in dry season was often higher than 30 $\mu g$Chl-a$L^{-1}$, peaked in the dry season of 2017 with a concentration of 150 $\mu g$Chl-a$L^{-1}$. According to the calculation of the trophic status at Saigon River, the PC station is mainly in the oligotrophic (2–4), mesotrophic (4–6) state in the dry season and sometimes ultraoligotrophic (0–2) in the rainy season (eg. 2017-2018 and 2020). At the BD station, hypertrophic status (8–10) was determined during the wet season of 2015, 2020 and summer 2017. The remaining time, urban sections of Saigon River were mainly mesotrophic (4–6) and eutrophic (6–8) stages.

```{r C3-HA, echo=FALSE, out.width='100%', fig.cap='Evolution of Chl-a and trophic index in the Saigon River from 2015 to 2020, at (a) Hoa An station and (b) Binh Khanh station', fig.align='center'}
knitr::include_graphics("figures/C3-HA.pdf")
```

Figure \@ref(fig:C3-HA) depicts Chl-a concentration and trophic index at Binh Khanh (BK) and Hoa An (HA) station. Both of these stations have low Chl-a concentration, especially at HA – Dongnai River. Therefore, the trophic indexes of these two stations are often in the state of ultraoligotrophic (0–2), oligotrophic (2–4).

## Part 2: Phytoplankton dynamics

### Spatio-seasonal variations

The HCA on phytoplankton densities during the dry season shown three major site groups in which phytoplankton community and densities were closely similar: SG01-SG04, SG05-SG12 and SG13-SG18 (Figure \@ref(fig:C3-HCA)). It also illustrated a higher similarity of phytoplankton’s characteristics between two groups SG01-SG04 and SG05-SG12, than downstream group (SG13-SG18). Three groups were identified during the rainy season: SG01-SG07, SG08-SG12 and SG13-SG18 (Figure \@ref(fig:C3-HCA)). Unlike HCA for the dry season, the similarity in phytoplankton compositions in the rainy season at upstream SG01 stretched to SG07, which is located in the urban area of HCMC. The close similarity of phytoplankton’s characteristics among the sites SG13-SG18 was comparable between the dry and the rainy seasons and had the most difference with the two groups in the upstream and urban areas. In general, HCA results show that phytoplankton composition and densities along the Saigon River varied along three sections of Saigon River (i) upstream, (ii) urban area and (iii) downstream of HCMC. The temporal variation of phytoplankton communities at these sections will be analyzed in turn based on bi-weekly monitoring over a year at three representative sites (SG01, SG10 and SG18).

```{r C3-HCA, echo=FALSE, out.width='100%', fig.cap='Spatial cluster analysis of phytoplankton abundance for longitudinal profile monitoring in dry and rainy season. Red, green and blue colors describe the upstream, urban area and downstream section of Saigon River, respectively.', fig.align='center'}
knitr::include_graphics("figures/C3-HCA.pdf")
```

### Structure and density of phytoplankton

```{r C3-profile, echo=FALSE, out.width='100%', fig.cap='Densities of dominant phytoplankton species and Chl-a concentrations in longitudinal profile monitoring in (a) dry season, April 2017 and in (b) rainy season, October 2017. Legend: “Total” is total phytoplankton abundance.', fig.align='center'}
knitr::include_graphics("figures/C3-profile.pdf")
```

The dominant phytoplankton communities identified along the two longitudinal profiles are presented in Figure \@ref(fig:C3-profile). It consisted in mainly species of diatoms (e.g., *Cyclotella* cf*. meneghiniana*, *Leptocylindrus danicus* Cleve 1889, *Synedra* sp, *Amphiprora* sp, *Aulacoseira granulata* (Ehrenberg) Simonsen 1979 , *Nitzschia* cf*. palea*; *Navicula* sp), cyanobacteria (e.g., *Raphidiopsis raciborskii* (Woloszynska) Aguilera et al. 2018, *Microcystis* spp, *Pseudanabaena* sp, *Dolichospermum* sp, *Oscillatoria* sp), green algae (e.g, *Scenedesmus acuminatus* (Lagerheim) Chodat 1902), and euglenoids (e.g., *Euglena* sp, *Trachelomonas* *volvocina* (Ehrenberg) Ehrenberg 1834) (Table \@ref(tab:C2-table1-domiant)).

```{r C2-table1-domiant, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Dominant species of phytoplankton based on the spatial and temporal monitoring in the Saigon River") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "Hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH3-table1-Dominant species of phytoplankton.pdf") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

The longitudinal profiles along the Saigon River revealed contrasted phytoplankton densities, varying from 48,200 to 5,296,000 individuals $L^{-1}$ during the dry season (April 2017) and from 1,400 to 35,120 individuals $L^{-1}$ during the rainy season (October 2017) (Figure \@ref(fig:C3-profile)). During the dry season, the phytoplankton densities shown a similar range of lower densities (\< 0.5 million individuals $L^{-1}$) in upstream (SG01-SG04) and downstream (SG11-SG18), but they drastically increased up to around 5.0 million individuals $L^{-1}$ from SG05 to SG10. In contrast, during the rainy season longitudinal profile, there was no sudden increase in phytoplankton densities, but they gradually increased from SG01 to SG14 and decreased in downstream of Saigon River (Figure \@ref(fig:C3-profile)b). Beside the significant difference in phytoplankton densities between dry and rainy seasons (lower than 100 times), the dominant species along the Saigon River also differed between the two distinct seasons. The dominant phytoplankton species in dry season longitudinal profile were *Leptocylindru*s *danicus* (SG01 to SG04, km 0 to km 13) and *Cyclotella* cf. *meneghiniana* (SG05 to SG18, km 17 to km 61) (Figure \@ref(fig:C3-profile)a, Table \@ref(tab:C2-table1-domiant)). The dominant species in rainy season longitudinal profile were *Trachelomonas volvocina* (SG01 to SG12, km 0 to km 45), *Raphidiopsis raciborskii* (SG13 to SG15, km 49 to km 53), and *Microcystis* spp., *Psedudanabaena* sp (SG13 to SG18).

```{r C3-season, echo=FALSE, out.width='100%', fig.cap='Temporal phytoplankton densities at the sites (left panels) and the relationship between total phytoplankton density and Chl-a (right panels) at each sampling site. Legend: “Total” is total phytoplankton abundance, “Others” is the other dominant species.', fig.align='center'}
knitr::include_graphics("figures/C3-season.pdf")
```

The phytoplankton communities at the three sampling sites (SG01; SG10; SG18) are presented in Figure \@ref(fig:C3-season). The dominant phytoplankton species, genera and their densities varied among the sites and were globally characterized by: diatoms, (e.g. *A. granulata, Navicula* spp., *L. danicus*, *Cyclotella* cf. *meneghiniana*), cyanobacteria (e.g. *Microcystis* sp., *R. raciborskii, Pseudanabaena* sp.) green algae (e.g. *S. acuminatus*), euglenoids, (e.g. *T. volvocina*) and dinoflagellates (e.g. *Peridinium* sp.) (Figure \@ref(fig:C3-season)). A high correlation between total phytoplankton density and chlorophyll-*a* at SG01 ($\mathrm{R^2}$= 0.91) and SG10 ($\mathrm{R^2}$= 0.93) was observed (Figure \@ref(fig:C3-season)). The phytoplankton densities varied between sites with several orders of magnitude, from 400–327,500 individuals $L^{-1}$ at SG01, 9,330–2,733,000 individuals $L^{-1}$ at SG10, and 1,360–45,700 individuals $L^{-1}$ at SG18 (Figure \@ref(fig:C3-season)). At SG01, the densities were elevated in March and April 2017 (mainly *L. danicus* and *A. granulata*) and very low (\< 50,000 individuals $L^{-1}$) the rest of the survey (Figure \@ref(fig:C3-season)a). At SG10, phytoplankton densities were extremely high from January to May 2017 (70% *C. meneghiniana* and 30% *L. danicus*) and lower than 200,000 individuals $L^{-1}$ the remaining months (Figure \@ref(fig:C3-season)b). At SG18 – downstream of Saigon River, there was no significant difference in phytoplankton density between rainy and dry seasons (p> 0.1). Phytoplankton densities were the highest during two periods, from January to March 2017 (mainly *C. meneghiniana*) and from June to November 2017 (mainly *Microcystis* spp), and lower than 10,000 individuals $L^{-1}$ the rest of the year (mainly *L. danicus*, *Pseudanabaena* sp and *R. raciborskii*).

### Drivers of eutrophication

#### Impact of urban discharge on phytoplankton communities

Based on bi-weekly monitoring data at SG01, SG10 and SG18, the PCA was used to identify the various patterns of environmental parameters and similarities between upstream, urban area and downstream sections of the Saigon River. The PCA results could be used to identify the impact of urban discharge on the phytoplankton biomass, supported by the high correlation of Chl-a and phytoplankton abundances ($\mathrm{R^2}$= 0.9, Figure \@ref(fig:C3-season)). Similar to the HCA results on phytoplankton abundances along the Saigon River, the PCA results on environmental parameters separate the three monitoring locations (SG01, SG10 and SG18) (Figure \@ref(fig:C3-PCA)). The PCA results shown that the water quality in SG10 suffered from the most diverse environmental parameters (TSS, TOC, nutrients, Chl-*a* and dissolved metals), which were mainly related to pollution from the urban area of HCMC. During the dry season, the first principal component (PC1) mainly represented metal, pH and salinity variables mainly linked to SG18 (downstream section) (Figure \@ref(fig:C3-PCA)). In PC2 of the dry season and PC1 of the rainy season, the most important variables were the nutrients, TOC, Chl-a, which are associated with SG10 (urban section of the Saigon River). This illustrated that the urban area was affected by urban discharges containing high pollutant concentrations from the connecting canals and creeks during both seasons.

```{r C3-PCA, echo=FALSE, out.width='100%', fig.cap='PCA of 20 environmental variables (physicochemical parameters, Chl-a, metals) in (a) dry season and (b) rainy seasons in the Saigon River. Upstream, urban area and downstream are SG01, SG10, and SG18, respectively. The environmental variables with longer length are the controlling parameters. Note that the axes of PCA for rainy season were rotated for easier comparison.', fig.align='center'}
knitr::include_graphics("figures/C3-PCA.pdf")
```

RDA method allowed to evaluate the influence of environmental variables on the phytoplankton composition and abundances [@Cavalcanti2020; @Varol2018]. Based on RDA, 18 environmental parameters can explain 72% ($\mathrm{R^2}$adjusted = 0.35) and 68% ($\mathrm{R^2}_{adjusted}$= 0.31) of variation of phytoplankton abundances during dry and rainy seasons, respectively. Only abundances of *L. danicus*, *Cyclotella* cf. *meneghiniana*, *Microcystis* spp, *Eunotia* sp had a clear relationship with environmental parameters while other dominant species (mainly in rainy season) such as *Peridinium* sp, *A.* *granulata*, *T.* *volvocina*, *R. raciborskii, Pseudanabaena* sp have not been explained by RDA. It is noticed that even though total phytoplankton abundance in the Saigon River was highly correlated with phytoplankton biomass (assessed by chlorophyll *a*), the use of phytoplankton abundances had some limitations in assessing ecological relationship because some phytoplankton species, with lower densities, did not have a high correlation with chlorophyll *a* such as *R. raciborskii*, *Peridinium sp*.

```{r C3-RDA, echo=FALSE, out.width='100%', fig.cap='RDA on the relationships between environmental variables (as driving factors) and phytoplankton abundances in (a) dry and (b) rainy season in the Saigon River. Lptcy – L. danicus, Prdnm – Peridinium sp, Alcsr – A. granulata, Mcrcy - Microcystis spp, Eunot - Eunotia sp, Rphdp – R. raciborskii, Trchl – T. volvocina, Cyclt - Cyclotella cf. meneghiniana, Psdnb - Pseudanabaena sp.', fig.align='center'}
knitr::include_graphics("figures/C3-RDA.pdf")
```


During the dry season, the abundance of *Cyclotella* cf. *meneghiniana* was related with pH, salinity, TN, DIN, Chl-a, Mn, Mg and Mo (Figure \@ref(fig:C3-RDA)a). This indicates that the majority of *Cyclotella* cf. *meneghiniana* abundance contributed to the biomass of Chl-a measured at the urban area of the Saigon River. While Mn plays an important role in the photosynthesis, Mg is an essential element in algal cellular proliferation which reduces the loss of chlorophyll pigments [@Bogorad1966]. Therefore, the increase of Mn and Mg in the urban area could support the growth of phytoplankton in the Saigon River. However, according to the longitudinal profile during the dry season, there was no difference in Mn, Mg concentrations between at SG04 and SG05, while diatom abundance suddenly increased from about 500,000 to 4,000,000 ind $L^{-1}$. *Cyclotella* cf. *meneghiniana* became the highest densely phytoplankton, replacing *L*. *danicus* at SG05. Therefore, the metals (Mn and Mg) were not factors leading to the predominance of *Cyclotella* cf. *meneghiniana* in the Saigon River. Relationship between *Cyclotella* cf. *meneghiniana* and salinity clarified this phenomenon. The salinity intrusion of the Saigon River during the dry season was about 90-100 km from the estuary mouth (SG05-SG06), the salinity reached a maximum of about 0.5 during the dry season at SG05. Therefore, the salinity at SG04 was rarely greater than zero. This interesting finding shown that freshwater diatom *Cyclotella* cf. *meneghiniana* can also dominate in non-zero salinity areas (urban area and downstream section). The second dominant diatom in the Saigon River, *L. danicus* was not associated with nutrients but with TOC and inversely correlated with $\mathrm{DSi}$ and Mg. The silicate is a macronutrient for diatoms because it is the main component of the frustules of diatoms [@Wetzel2001]. However, it is out of our expectation that there was no correlation between diatoms and $\mathrm{DSi}$. The sufficient levels of silicate in the Saigon River (> 0.5 mg $L^{-1}$) helps to explain the lack of this relationship [@Nguyen2019]. Besides that, the increase of Mg concentration from SG01 to SG18 (1224 to 14092 $\mu g$$L^{-1}$) has led *Cyclotella* cf. *meneghiniana* outgrown *L. danicus* which fell sharply at SG18. *Leptocylindrus danicus* and TOC were positively linked in this study because high phytoplankton abundance contributed to the TOC in surface water, as previously stated in @Nguyen2020c.

During the rainy season, RDA results generally show that diatom densities (*Cyclotella* cf. *meneghiniana*, *L. danicus*) were positively related to variance of TOC, nutrients (DIN, TN, DIP) and metals (Mn, Mo, Co). The positive relationship of phytoplankton divisions with nitrogen and phosphorus compounds in this study is supported by previous investigations [@Varol2018; @Liu2020]. The abundance of *Eunotia* sp. during the rainy season was highly related with the Cu, Zn, Co concentrations but negatively related with pH. This species might be sensitive lower concentration of pH in SG01 where the pH was about 6.2 – 6.4. This was in contrast to *Cyclotella* cf. *meneghiniana* or *L. danicus* that dominated at SG10 (pH range 6.7 - 6.8). This result shows that the small difference in pH between upstream (SG01) and urban area (SG10) may result in the change of dominant species along the Saigon River.

#### Spatiotemporal effect on phytoplankton densities

In the longitudinal profiles along the Saigon River (SG01 - SG18) and during the bi-weekly monitoring (SG01, SG10 and SG18), we found significant difference of phytoplankton’s characteristics between the urban area and the upstream and downstream parts of HCMC (Figure \@ref(fig:C3-profile)), suggesting influence of localized environmental factors. The phytoplankton abundance in the urban area were much higher than in upstream and downstream of Saigon River. The phytoplankton abundances at SG01 and SG18 (Figure \@ref(fig:C3-season)a, c) were within the range of the phytoplankton abundances in the Vam Co River (an affluent of the downstream Saigon River) 920 – 383,600 individuals $L^{-1}$ [@Dao2016]. During the dry season longitudinal profile, the diatom *L. danicus,* a brackish water species, was dominated at the first four sites, SG01-SG04, which could be related to the slight increase of salinity up to SG01 observed at these stations and could support the conditions for the occurrence and out competition of this diatom to other phytoplankton. The freshwater species (except *L. danicus –* a brackish diatom) dominated during the dry period in March and April 2017 at the site SG01, which is coherent with the freshwater characteristic (salinity \~ 0) nearly all year around at the site SG01. The abundance of phytoplankton at SG10 during the dry season was around 10 and 100 times higher than that at SG01 and SG18, respectively (Figure \@ref(fig:C3-season)). These results were comparable with the phytoplankton abundance in the Tigris and Paraguay Rivers which have been also under effect of anthropogenic pressures [@Domitrovic2002; @Varol2018]. At SG10, the dominance of *Cyclotella* cf. *meneghiniana* and *L. danicus* could be driven by environmental parameters such as salinity, trace elements (e.g. Mo, Mn) and high concentrations of nutrients from urban discharge. @Dao2016 also observed the dominance of diatoms (*Cyclotella* spp and *Eunotia* spp) in the surface waters of Vam Co River (15 km from Saigon River Estuary mouth), which is in accordance with the dominance of diatoms species (e.g. *Navicula* sp.*, A. granulata, Cyclotella* cf. *meneghiniana,* and *L. danicus)* in our study (Table \@ref(tab:C2-table1-domiant)).

Unlike upstream and urban area of HCMC, phytoplankton abundance was quite low in the downstream of HCMC (Figure \@ref(fig:C3-profile)) and dominated by both diatoms (*Cyclotella* cf. *meneghiniana*) and cyanobacteria (e.g. *Microcystis* spp, *Pseudanabaena* sp, *Raphidiopsis raciborskii*) (Table \@ref(tab:C2-table1-domiant)). The site SG16 is the meeting point of the Saigon and Dongnai Rivers, and the water discharge of the Dongnai River is around 12 times higher than the one of the Saigon River [@Nguyen2019]. The phytoplankton community and abundance would thus be much more influenced by the water from the Dongnai River than by water from the Saigon River at site SG18. Proliferation and bloom forming of cyanobacteria have been commonly observed in the Tri An Reservoir, upstream of the Dongnai River [@Dao2016; @Nguyen2020c]. We found that there was a high similarity of cyanobacteria groups (*Microcystis* spp) found in downstream HCMC and at Tri An Reservoir [@Dao2016]. Therefore, the commonly presence and dominance of cyanobacteria (Table \@ref(tab:C2-table1-domiant)) especially at SG18 (Figure \@ref(fig:C3-season)c) could result of (i) the consequence of the diffusion of waters enriched with cyanobacteria, from Tri An Reservoir [@Nguyen2020c]; and/or (ii) the saline intrusion and anthropogenic emission which enhanced the cyanobacterial development in the most active estuarine section [@Paerl2009] .

Beside the influence of localized environmental factors, seasonal effects lead to a significant change in phytoplankton communities in tropical estuaries [@Cavalcanti2020; @vanChu2014; @Bledsoe2004] . The results of bi-weekly monitoring show that there was a clear difference in the abundance and phytoplankton species between the dry and rainy seasons in the Saigon River. In contrast to the rainy season, the residual water discharge in the Saigon River was very low (30$m^{3}$ $s^{-1}$) during the dry season [@Camenen2021]. The residence time of the water body in this urban section of Saigon River was estimated to be around two months [@Nguyen2021]. High load of nutrients from HCMC and long residence time may strongly enhance the phytoplankton development during the dry season in the urban section of the Saigon River. @Ferreira2005 found that higher phytoplankton biomass is more prevalent in estuaries with long residence time even without anthropogenic impact. However, during the rainy season, stronger water discharge from upstream of the Saigon River had shorten residence time of water body. This leads to a decrease of phytoplankton development before being flushed out of the estuary during the rainy season. The seasonal variation in phytoplankton abundance in the Saigon River has also been found in other tropical estuaries such as Bach Dang Estuary (Vietnam) [@vanChu2014], Red River (Vietnam) [@Duong2019] and Paciencia River Estuary (Brazil) [@Cavalcanti2020]. During the dry season*, L.* *danicus* and *A. granulata* were dominant in upstream of HCMC (SG01), while the freshwater algae *Peridinium* was dominant there in rainy season. At SG10, there was the dominance of another freshwater species (*T. volvocina*) beside the dominance of *Cyclotella* cf. *meneghiniana* and *L. danicus* (Figure \@ref(fig:C3-season)). The appearance of freshwater algae could be due to the prevention of saline intrusion supported by the high discharge in the rainy season. This could also be related to the Saigon River morphometry, which is very deep at this location (> 12 m), and has a lower water current locally, freshwater algae from inland, and eutrophic canals and creeks connecting to the river.

During the rainy season longitudinal profile, the similarity of phytoplankton characteristics among the sites did not clearly separate as in dry season (Figure \@ref(fig:C3-profile)). This may be linked with the higher water discharge upstream of Saigon River during the rainy season, resulting in the influence of inland water down to the urban area. The supporting evidences for this hypothesis are that (i) there was not a significant increase of phytoplankton abundance along Saigon River (max of 35,120 ind $L^{-1}$), and (ii) typical inland algae (*T. volvocina*; Reynolds, 2006) dominated during the rainy season longitudinal profile. Besides, the dominance of the freshwater species of *T. volvocina* and *Pseudanabaena* sp. at the site urban section of Saigon River by the end of the rainy season (Oct, Nov 2017) is supported by the dominance of *T. volvocina* at 12 sites during the rainy season (Figure \@ref(fig:C3-profile)). The euglenoids *Trachelomonas* and dinoflagellates *Peridinium* are typical genera of nutrient enrich trophic lakes observed in the Saigon River.

Another noteworthy difference is the formation of cyanobacterial species during the rainy season. Bi-weekly monitoring and longitudinal profile during rainy season both detected cyanobacterium *Microcystis* spp and *R*. *raciborskii* in the downstream of the Saigon River. While *Microcystis* spp could have originated from Tri An Reservoir [@Dao2016], Cyanobacterium *R. raciborskii* may be related to the high TSS concentrations in the downstream area during the rainy season. Some cyanobacterial species such as *R. raciborskii* and *Microcystis* spp. have air aerotopes in their cells, giving them a buoyant capacity to surfaceater [@Mhlanga2006]. The cyanobacterium *R. raciborskii* has a filamentous form which helps to capture more light in turbid waters [@Reynolds2006]. Therefore, the species *R. raciborskii* could withstand the low radiation in water column in high TSS conditions and maintain their normal photosynthesis activities in topwater.

## Part 3: Interaction of eutrophication and GHGs

### Longitudinal profiles

The spatial variations of water quality variables in the Saigon River in dry and rainy seasons are shown in Figure \@ref(fig:C3-gas-profile). Temporal variation of several water quality parameters was observed between the dry and rainy seasons. Spatial variation was also observed between upstream sites and downstream sites of HCMC. TSS and $\mathrm{NO_{3}^{-}}$ concentrations were not significantly different between seasons (p > 0.05). In the dry season, salinity, water temperature, pH, DO, Chl-a and alkalinity were higher (p \< 0.05), while TOC, $\mathrm{PO_{4}^{3-}}$, $\mathrm{NH_4^{+}}$ were lower in comparison to the rainy season (p \< 0.05). Besides, the difference in concentrations between seasons and water quality variables varied along the Saigon River. Salinity, TSS, pH and DO gradually increased from 0 to 23 ppt, 11 to 340 mg$L^{-1}$, 5.7 to 8.1 pH, 0.5 to 6.2 mg$\mathrm{O_{2}}$ $L^{-1}$, respectively.

```{r C3-gas-profile, echo=FALSE, out.width='100%', fig.cap='Water quality variables and GHG concentrations in the Saigon River Estuary 2019-2020. The lines and color bands represent the mean values and maximum/minimum values of each parameter. The histogram represents water samples collected in the four urban canals of HCMC. Orange and blue colors indicate the samples collected during the dry and rainy seasons, respectively.', fig.align='center'}
knitr::include_graphics("figures/C3-gas-profile.pdf")
```

In the urban section (SG03 to SG11), Chl-a increased from 0.1 to 26.2 $\mu g$$L^{-1}$ in the dry season. $\mathrm{NH_4^{+}}$, $\mathrm{PO_{4}^{3-}}$ increased significantly in rainy season, 0.1 - 1.1 mgN $L^{-1}$ and 0.01 – 0.61 mgP $L^{-1}$, respectively. TOC had a slight increase while POC had a stronger increase in this area up to 15% and then decreased sharply to nearly 0% at the estuary mouth. $\mathrm{NO_{3}^{-}}$ fluctuated from 0.1 – 0.8 mgN $L^{-1}$ between sampling sites in both seasons. The total alkalinity in upstream of Saigon River to the urban influences was relatively low (5 – 45 mg$CaCO_3$$L^{-1}$), with even lower values observed during the rainy season (Figure \@ref(fig:C3-gas-profile)).

### Greenhouse gases

#### GHG concentrations

The spatial distributions of the GHGs in dry and rainy seasons in the Saigon River are illustrated in Figure \@ref(fig:C3-gas-flux). In general, GHGs tended to increase gradually in the urban section and to decrease downstream. The concentrations of $\mathrm{CO_2}$ were 404 – 5947 $\mu$gC $L^{-1}$ (mean 2697 $\pm$ 1792) in the dry season and 739 – 5620 $\mu$gC $L^{-1}$ (mean 3641 $\pm$ 1567) in the rainy season. The difference in $\mathrm{CO_2}$ concentration between the two seasons was about 30%, with a higher $\mathrm{CO_2}$ concentration in the rainy season. $\mathrm{CH_4}$ concentrations were 0.6 – 3.6 $\mu$gC $L^{-1}$ (mean 1.8 $\pm$ 1.1) and 0.9 – 117 $\mu$gC $L^{-1}$ (mean 11.3 $\pm$ 22.7) respectively, meaning that the concentration of $\mathrm{CH_4}$ in the dry season was about seven times lower than in the rainy season. $\mathrm{N_2O}$ concentrations ranged from 0.1 to 20.1 $\mu$gN $L^{-1}$ (mean 4.6 $\pm$ 6.2) in dry season and from 0.0 to 2.9 $\mu$gN $L^{-1}$ (mean 1.2 $\pm$ 0.9) in rainy season. The concentration of $\mathrm{N_2O}$ in the dry season was around four times higher than in the rainy season. These results clearly show that there were important seasonal differences in GHGs in the Saigon River. Compared to the global mean atmospheric concentration, the average saturation concentrations of $\mathrm{CO_2}$, $\mathrm{CH_4}$, $\mathrm{N_2O}$ in the dry and rainy seasons were 13, 52, 37 and 18, 332, 9 times higher, respectively. Calculations based on global warming potential, i.e., $\mathrm{CO_2}$, $\mathrm{CH_4}$, $\mathrm{N_2O}$ expressed in total $\mathrm{CO_2}$ equivalents ($\mathrm{CO_2}$-eq) in the Saigon River, led to concentrations of 14311 $\mu g$$\mathrm{CO_2}$-eq$L^{-1}$ in dry season and 14935 $\mu g$$\mathrm{CO_2}$eq$L^{-1}$ in the rainy season. Their respective contributions are 69.1%, 0.5%, 30.4% and 89.4%, 2.8%, 7.8%, respectively for the three gases ($\mathrm{CO_2}$, $\mathrm{CH_4}$, $\mathrm{N_2O}$).

```{r C3-gas-flux, echo=FALSE, out.width='100%', fig.cap='Spatial and seasonal concentrations (in $\\mu g$ $L^{-1}$), water-air fluxes of GHGs (in g $m^{-2}$ $d^{-1}$ or mg $m^{-2}$ $d^{-1}$) and eutrophication index (TRIX) in dry and rainy season in the Saigon River Estuary. The water-air fluxes were calculated for each section (i) upstream, (ii) urban area of HCMC and (iii) downstream of HCMC. It is noticed that the maps were built by using the Inverse Distance Weighting Interpolation (QGIS 3.18.1, 19-Mar-2021) between 27 sampling sites over the 141 km of the Saigon River. No data were acquired in the upper 55 km of the Saigon river, from Dau Tieng Reservoir to UP01 (the river is then underlined with white background).', fig.align='center'}
knitr::include_graphics("figures/C3-gas-flux.pdf")
```

In urban canals of HCMC, the concentrations of GHGs (except $\mathrm{N_2O}$) were much higher than in the Saigon River. The mean concentrations of $\mathrm{CO_2}$), $\mathrm{CH_4}$ and $\mathrm{N_2O}$ were 5878 $\pm$ 2739 $\mu$gC $L^{-1}$, 304 $\pm$ 298 $\mu$gC $L^{-1}$ and 3.0 $\pm$ 3.0 $\mu$gN $L^{-1}$ respectively. During the monitoring campaign, we observed plenty of gas bubbles appearing on the water surface of urban canals (Photo S2 in Appendix). At Ben Nghe Canal (close to SG10), $\mathrm{CH_4}$ concentrations were more than 600 and 1000 $\mu$gC-$\mathrm{CH_4}$ $L^{-1}$ in dry and rainy seasons, respectively. On the opposite, the $\mathrm{N_2O}$ concentrations in this canal were 2.3 $\mu$gN-$\mathrm{N_2O}$ $L^{-1}$ in dry season, and very low in rainy season (Figure \@ref(fig:C3-gas-flux)). For other canals, the $\mathrm{CO_2}$ concentrations were within the same range of $\mathrm{CO_2}$ concentration in the Saigon River, while $\mathrm{CH_4}$ concentrations were 2 – 25 times higher than those in the river.

#### GHGs fluxes

The high supersaturation values of GHGs suggest that this tropical estuary contributed significantly to GHGs emissions. With the normalized gas transfer velocity of 12.3 cm $h^{-1}$ (10 - 18 cm $h^{-1}$), the mean density fluxes of $\mathrm{CO_2}$, $\mathrm{CH_4}$, $\mathrm{N_2O}$ in the Saigon River were 31870 $\pm$ 11224, 26.5 $\pm$ 42.5 mg C $m^{-2}$ $d^{-1}$ and 26.0 $\pm$ 29.6 mg N $m^{-2}$ $d^{-1}$. Similar to the concentrations of GHGs (Figure \@ref(fig:C3-gas-flux)), the distribution of GHGs fluxes varied between the upstream and downstream sections of the Saigon River and between the rainy and dry seasons. During the dry season, $\mathrm{CO_2}$ fluxes were higher in the upstream (about 78000 mgC $m^{-2}$ $d^{-1}$ at UP03 in 2020) and much lower in the estuary mouth (about 1700 mgC $m^{-2}$ $d^{-1}$ at SG23). $\mathrm{CH_4}$ flux decreases gradually from about 8 mgC $m^{-2}$ $d^{-1}$ in the upstream and urban section to about 5 mgC $m^{-2}$ $d^{-1}$ in the estuary mouth. $\mathrm{N_2O}$ flux was the highest at urban section SG05 (5km landward from an urban canal) with 341 mgN $m^{-2}$ $d^{-1}$ and undersaturated at estuary mouth with -1.5 mgN $m^{-2}$ $d^{-1}$. $\mathrm{CO_2}$ fluxes in the upstream section did not differ significantly between the two seasons, while the $\mathrm{CO_2}$ flux in the urban section in the rainy season was about 1.5 times higher than in the dry season. The higher $\mathrm{CO_2}$ flux tended to move further downstream in the rainy season, compared to the dry season. The $\mathrm{CO_2}$ flux at the estuary mouth thus reached more than 18000 mgC $m^{-2}$ $d^{-1}$, about 10 times higher than in the dry season. $\mathrm{CH_4}$ fluxes reached the highest level in urban section with up to 466 mgC $m^{-2}$ $d^{-1}$ in SG07, where urban wastewater is directly discharged. It is about 70 times higher than $\mathrm{CH_4}$ fluxes values observed in upstream and downstream sections. In contrast to $\mathrm{CH_4}$, there was no increase of $\mathrm{N_2O}$ flux in the urban section in the rainy season but instead, a decrease (about ten times) compared to the dry season. The $\mathrm{N_2O}$ fluxes at upstream and downstream in rainy season were not significantly different from dry season.

#### Comparison with other estuarine systems

GHGs in the Saigon River Estuary were supersaturated, indicating that this tropical urbanized estuary is a source of $\mathrm{CO_2}$), $\mathrm{CH_4}$, and $\mathrm{N_2O}$ to the atmosphere during both seasons. We compared our present GHGs concentrations and associated fluxes between Saigon River with other subtropical and temperate systems experiencing different environmental conditions (Table \@ref(tab:C3-GHGs-compare)). In addition, the emission fluxes of $\mathrm{CH_4}$ and $\mathrm{N_2O}$ were converted to $\mathrm{CO_2}$ equivalents to evaluate the total impact of GHGs on global warming potential. Overall, this comparison shows that the concentrations/fluxes of GHGs in the estuaries, heavily affected by emissions from urbanization, industry, and agriculture, were much higher than other areas, whatever the climatic region considered, revealing that anthropogenic activity is the main driver.

GHGs concentrations in the Saigon River were at the same level as in other urbanized estuaries such as Nanfei River, China, [@Zhang2021], Zambezi River, Africa [@Teodoru2015], Guadalete Estuary, Spain, [@Burgos2015] and Seine River, France [@Marescaux2018] (Table \@ref(tab:C3-GHGs-compare)). Zambezi River had TOC concentration of about 5.8 mgC$L^{-1}$, close to the 5.2 mgC$L^{-1}$ of the Saigon River. The $\mathrm{CO_2}$ concentrations of the two systems in their urban areas were similar, but the $\mathrm{CH_4}$ concentration in the Zambezi River was about two times higher while the $\mathrm{N_2O}$ concentration was about 10 times smaller than that of the Saigon River. Similarly, the Indian Adyar River, highly polluted because of untreated domestic wastewater in Chennai city, also shown high $\mathrm{CH_4}$ and low $\mathrm{N_2O}$ concentrations [@NirmalRajkumar2008]. The high $\mathrm{CH_4}$ and low $\mathrm{N_2O}$ concentrations were also observed in the urban area of Saigon River in the rainy season (Figure \@ref(fig:C3-gas-flux)). Based on $\mathrm{CO_2}$ equivalents calculation, the reduction of $\mathrm{N_2O}$ will contribute more efficiently to the reduction in global warming potential than the reduction of $\mathrm{CH_4}$. Therefore, controlling the concentration of nutrients in wastewater and in the river can be thus considered in priority to improve the effectiveness of reducing global warming.

Concentrations of GHGs in non-urban estuaries can be about 10-100 times lower than in urbanized estuaries, especially for $\mathrm{CO_2}$ and $\mathrm{N_2O}$. The rivers, peat-dominated or mangrove-dominated regions (e.g., @Hu2018; @Muller2016; @Reithmaier2020) revealed low GHGs concentrations regardless of climatic region (tropical or temperate estuaries). This may suggest that despite the existence of higher rates of nitrification and aerobic/anaerobic respiration in tropical/subtropical estuaries than in temperate estuaries, linked to high temperature, these rates might be masked by polluted water from urban discharge. For example, Guadalete Estuary (Spain), an urban estuary in a south Mediterranean region and Po River (Italy) with continental temperate climate still had $\mathrm{CO_2}$), $\mathrm{CH_4}$ and $\mathrm{N_2O}$ concentrations several times higher than the peat-dominated river of the tropical region in Malaysia. Moreover, the Po River (Italy) under the impact of nitrate pollution can lead to $\mathrm{N_2O}$ concentrations/fluxes much higher than the ones in urbanized tropical estuaries such as Saigon River (Vietnam), Adyar River (India).

```{r C3-GHGs-compare, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Comparison of mean GHGs concentrations and fluxes ($\\mathrm{CO_2}$ equivalent) in estuaries under different environmental conditions (Tropical, Temperate, Mediterranean) and human pressures.") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "Hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH3-table2-Comparison of mean GHGs concentrations and fluxes.pdf") %>%
  landscape() %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

### Nitrogen and carbon cycling

The statistical methods (RDA, HP) allowed evaluating the relationship and contribution of eutrophication, physicochemical variables to the variations of GHGs in the Saigon River Estuary. We searched here to identify more precisely the processes involved in the production and consumption of GHGs to evaluate the behavior of GHGs in this urbanized tropical estuary.

#### Nitrification and denitrification

$\mathrm{N_2O}$ is predominantly released via microbial processes, such as nitrification and denitrification, occurring in the water column, sediments or in suspended particles [@Bange2008]. It is known that $\mathrm{N_2O}$ production from these processes is highly dependent on DO concentration [@Bie2002]. In water column with the presence of $\mathrm{O_{2}}$, $\mathrm{N_2O}$ is mainly formed during the first step of nitrification, as a by-product of ammonium oxidation, from $\mathrm{NH_4^{+}}$ to $NO_2^{-}$ [@Garnier2007], but also as nitrifier denitrification at low oxygenation [@Wrage2001; @Zhu2013]. Besides, denitrification can play the main role in $\mathrm{N_2O}$ production under hypereutrophic condition and associated anoxia [@Galantini2021]. @Manjrekar2020 found a positive relationship between $\mathrm{N_2O}$ and $\mathrm{NH_4^{+}}$ in several tropical estuaries, which suggested that nitrification was the possible mechanism for $\mathrm{N_2O}$ production in these systems. in the Saigon River, we did not detect this correlation (Figure \@ref(fig:C3-season)). In dry season, the $\mathrm{N_2O}$ production in the Saigon River can be due to nitrifier denitrification under low DO concentration condition (2mg $L^{-1}$). In rainy season, denitrification may account mostly for the $\mathrm{N_2O}$ production with lower DO concentration (1 mg $L^{-1}$) (Figure \@ref(fig:C3-gas-profile)).

```{r C3-AOU-N2O, echo=FALSE, out.width='100%', fig.cap='Relationship between $\\mathrm{N_2O}$ excess ($\\mathrm{N_2O}$xs) and apparent oxygen utilization (AOU=DOair -DOwater, i.e., difference between measured DO in water and its equilibrium saturation concentration) in (a) dry and (b) rainy season 2019 – 2020 in the Saigon River. Upstream (UP01-UP04, SG01-SG03); Urban (SG04-SG15); Downstream (SG16-SG23). The linear correlations of AOU-$\\mathrm{N_2O}$xs indicate nitrification as the main contributor to $\\mathrm{N_2O}$ production. Correlation breaks indicate the impact of denitrification on $\\mathrm{N_2O}$ production.', fig.align='center'}
knitr::include_graphics("figures/C3-AOU-N2O.pdf")
```

Several studies determined $\mathrm{N_2O}$ production by nitrification process based on a linear relationship between $\mathrm{N_2O}$ excess ($\mathrm{N_2O}$xs, i.e., $\mathrm{N_2O}$ supersaturation) and apparent oxygen utilization (AOU) (e.g., @Nevison2003; @Santoro2021; @Brase2017). We explored such a linear relationship in upstream and downstream of Saigon River in the dry season. These positive correlations are indicative of $\mathrm{O_{2}}$ consumption linked to $\mathrm{N_2O}$ production. Therefore, most of $\mathrm{N_2O}$ production in upstream, downstream sections and some parts in the urban area of HCMC was mainly derived from nitrification/nitrifier denitrification processes, as shown by a modeling approach we applied in previous work [@Nguyen2021]. We also observed a break of linearity with the highest $\mathrm{N_2O}$ supersaturation values at the highest AOU (i.e. low DO condition) in the urban stretch of Saigon River (SG04 – SG07 with $\mathrm{N_2O_{xs}}$ of 4.5 – 35  $\mu gNL^{-1}$, Figure \@ref(fig:C3-AOU-N2O)a). The broken linear relationship of AOU and $\mathrm{N_2O}$xs was mainly observed at highly urbanized estuaries, which provide the most intense $\mathrm{N_2O}$ production, supported by the denitrification process [@Murray2015]. Besides, $\mathrm{N_2O}$ emissions in the estuarine system are generally higher due to the propensity of nitrifying bacteria to attach onto suspended particulates, supporting high activities due to their longer residence times than the water itself [@Brion2000; @Murray2015]. There were some stations in urban areas aligned with the AOU-$\mathrm{N_2O}$xs slope in upstream and downstream stations. We hypothesized that the $\mathrm{N_2O}$ in these locations could be the result of nitrification/nitrifier denitrification (Figure \@ref(fig:C3-AOU-N2O)a).

Besides nitrifier denitrification, the combined nitrification and denitrification under moderate oxygen conditions (1–2 mg$L^{-1}$) in the urban area of Saigon River might also explain the higher increase in $\mathrm{N_2O}$ than in the upstream and downstream areas. @Bie2002 and @Murray2015 found that at intermediate $\mathrm{O_{2}}$ saturation with high concentrations of $\mathrm{NH_4^{+}}$, estuaries can produce a large amount of $\mathrm{N_2O}$ by the coupled nitrification and denitrification. in the Saigon River, this can be supported by the dominance of $\mathrm{NH_4^{+}}$ in DIN. Nevertheless, the $\mathrm{N_2O}$xs vs. AOU relationship does not allow calculating the proportion of nitrification, nitrifier denitrification or denitrification contributing to $\mathrm{N_2O}$ production [@Nevison2003].

In rainy season in contrast, the linear relationship $\mathrm{N_2O_{xs}}$ vs. AOU was broken at the lowest $\mathrm{N_2O_{xs}}$. The maximum $\mathrm{N_2O_{xs}}$in the urban area in rainy season was 10 – 20 times lower than in the dry season (Figure \@ref(fig:C3-AOU-N2O)). The oxygen depletion in the urban and downstream sectors was much higher (a difference of about 2 mg $L^{-1}$) than in the dry season caused by organic matter inputs from bottom sediment of the canals and the river, and edges (Figure \@ref(fig:C3-gas-profile)). The slope of $\mathrm{N_2O}$xs vs AOU in the upstream stations was lower than for the dry season despite higher ammonium substrate, but it would also indicate nitrification in this period. In the urban sector, high TOC level with the low oxygenation (0.5 – 1.0 mg$L^{-1}$ in the urban canals, Figure \@ref(fig:C3-gas-profile)) would promote denitrification as shown by the decrease in nitrate. Denitrification was even complete, as shown by the null values of $\mathrm{N_2O}$ (Figure \@ref(fig:C3-AOU-N2O)b). The $\mathrm{N_2O}$ production by nitrification in the downstream sector in rainy season was weaker than in dry season (i.e. the $\mathrm{N_2O}$xs/AOU slope was less steep) even with similar $\mathrm{NH_4^{+}}$ concentrations. The lower temperature, together with a high dilution in rainy season, might account for the lower activities of nitrifying bacteria. However, the difference in $\mathrm{N_2O}$ concentration downstream between the two seasons was not significant.

#### Photosynthesis and respiration

Most of estuaries are net heterotrophic systems because bacterial degradation of organic carbon exceeds gross primary production [@Robin2016; @Gupta2009; @Daniel2013]. Nevertheless, other studies have reported that some systems with potential eutrophication can reduce $\mathrm{CO_2}$ emissions when photosynthesis uptake is greater than respiration [@Grasset2020]. In tropical systems with high impacts from urban discharge (low DO, high TOC), the respiration of these anthropogenic organic carbon sources is mainly responsible for the $\mathrm{CO_2}$ supersaturation [@Gupta2009; @Nguyen2018]. Such a $\mathrm{CO_2}$ supersaturation was also observed in the Saigon River. The relationship between excess $\mathrm{CO_2}$ ($\mathrm{CO_2}$xs) and AOU can also be used to describe mechanisms responsible for $\mathrm{CO_2}$ production and DO consumption in the water column [@Cotovicz2021; @Han2017;@Zhai2005].

We observed a linear relationship between $\mathrm{CO_2}_{xs}$ and AOU, mainly in urban and downstream of Saigon River (Figure \@ref(fig:C3-AOU-CO2)). The correlation of AOU and $\mathrm{CO_2}_{xs}$ indicated that the high $\mathrm{CO_2}$ concentration derived mainly from aerobic respiration of organic matter in urban and downstream of Saigon River. The high concentration of $\mathrm{CO_2}_{xs}$ in upstream of Saigon River could also be due to allochthonous inputs of inorganic and organic carbon such as groundwater discharges, leading to an increase of $\mathrm{CO_2}$. The urban area might have additional contributions by anaerobic bacterial metabolism, supported by very low DO concentration (Figure \@ref(fig:C3-AOU-CO2)).

```{r C3-AOU-CO2, echo=FALSE, out.width='100%', fig.cap='Relationship between $\\mathrm{CO_2}$ excess ($\\mathrm{CO_2}_{xs}$) and apparent oxygen utilization (AOU) in dry and rainy seasons in the Saigon River. The linear correlations of AOU-$\\mathrm{CO_2}$xs indicate respiration as a major cause for $\\mathrm{CO_2}$ supersaturation. The allochthonous inputs of $\\mathrm{CO_2}$ can be either from rivers or lateral inputs.', fig.align='center'}
knitr::include_graphics("figures/C3-AOU-CO2.pdf")
```

Based on the phytoplankton respiration equation [@Redfield1963] of the degradation of organic carbon into $\mathrm{CO_2}$, the ratio $\mathrm{CO_2}_{xs}$/AOU (mol/mol, phytoplanktonic respiration quotient in aerobic) should be 0.77 (106/138) or 1.0 (106/106) in lower DO and pH conditions. In dry season, our study's ratios of $\mathrm{CO_2}_{xs}$/AOU were 0.7 – 1.0 mainly in urban and downstream sections, 1.1 – 1.6 in the upstream and some parts of urban area (Figure \@ref(fig:C3-AOU-CO2)). The respiration quotient was close to the range of theoretical Redfield respiration, indicating that aerobic respiration was the main factor producing $\mathrm{CO_2}$ in most urban areas and downstream. Meanwhile, potentially lateral inputs of $\mathrm{CO_2}$ or another internal process might occur in upper parts of urban and upstream sections because of higher $\mathrm{CO_2}_{xs}$/AOU quotient (>1.0) [@Smith2017]. For example, higher $\mathrm{CO_2}$ concentrations upstream in the Seine Riverere due to organic carbon inputs from soils, carbonate-enriched and enriched $\mathrm{CO_2}$ from groundwater discharges [@Marescaux2018; @Marescaux2018]. These patterns were similar to the $\mathrm{CO_2}$ simulation results in Red River (Northern Vietnam), where a large contribution of allochthonous inputs of $\mathrm{CO_2}$ was found [@Nguyen2018]. The $\mathrm{CO_2}$ concentrations in rivers can be affected by the catchment geology [@Duvert2018]. However, the Saigon River is dominated by igneous rocks with a low carbonate content [@David2018]. With a very low total alkalinity (mean 32 mgCaCO3 $L^{-1}$), the $\mathrm{CO_2}$ concentrations in surface water of the Saigon River would not thus be affected by buffering effect of bicarbonate concentration.

#### Methanogenesis and methanotrophy

While dissolved $\mathrm{CO_2}$ can be produced by both aerobic and anaerobic respiration of organic matter, $\mathrm{CH_4}$ is mainly produced by the methanogenesis process under anaerobic conditions and consumed by methanotrophy [@Hanson1996]. In estuarine systems, the methanogenesis process in water column only proceeds when oxidants such as $\mathrm{O_{2}}$, $\mathrm{NO_{3}^{-}}$ are completely depleted, the $\mathrm{CH_4}$ production thus mainly occurs in the sediment [@Borges2011]. The positive trend between $\mathrm{CH_4}$ excess and AOU can show how much $\mathrm{CH_4}$ can be left with increasing AOU in the water column, i.e., less oxygen for methane oxidation (Figure \@ref(fig:C3-AOU-CH4)).

```{r C3-AOU-CH4, echo=FALSE, out.width='100%', fig.cap='Relationship between $\\mathrm{CH_4}$ excess ($\\mathrm{CH_4}$xs) and apparent oxygen utilization (AOU) in dry and rainy seasons in the Saigon River. The negative correlations of AOU-$\\mathrm{CH_4}$xs indicate methane oxidation ($\\mathrm{CH_4}$ consumption/methanotrophy).', fig.align='center'}
knitr::include_graphics("figures/C3-AOU-CH4.pdf")
```

Similar to AOU-$\mathrm{CO_2}$xs (Figure \@ref(fig:C3-AOU-CO2)), extreme supersaturations of $\mathrm{CH_4}$ in the AOU-$\mathrm{CH_4}_{xs}$ were distributed in the areas with the lowest DO concentration (Figure \@ref(fig:C3-AOU-CH4)b). However, no correlation was found based on AOU-$\mathrm{CH_4}_{xs}$ analysis for the Saigon River. This result confirmed that high oxygen utilization (AOU) in this urbanized estuary was not conducive to $\mathrm{CH_4}$ oxidation in the water column. This finding was similar to an urbanized tropical lagoon in Brazil (Jacarepagua Lagoon) where $\mathrm{CH_4}$ concentrations could reach more than 1000 $\mu g$C$L^{-1}$ [@Cotovicz2021]. This study suggested that nitrifiers outcompeted the methanotroph community because of the strong nitrification process. Indeed, the analysis of AOU-$\mathrm{N_2O}_{xs}$ (Figure \@ref(fig:C3-AOU-N2O)) and AOU-$\mathrm{CO_2}_{xs}$ (Figure \@ref(fig:C3-AOU-CO2)) shown that DO concentration in the water column was affected by oxic respiration and nitrification, leading to DO depletion in the Saigon River. Our result differs from the studies of @Buioanh2018 or @Sierra2020b which observed a negative correlation between AOU-$\mathrm{CH_4}_{xs}$, indicating a $\mathrm{CH_4}$ decrease by methane oxidation in aerobic conditions. Similarly, @Zhang2020 found that 90% of $\mathrm{CH_4}$ formed from sediment was microbially consumed in the water column under aerobic conditions. In other words, the higher DO in water column tends to increase the methanotrophy and decrease the methanogenesis rate. This was consistent with lower concentrations of $\mathrm{CH_4}$ (about 1 $\mu g$C$L^{-1}$) in downstream of Saigon River where DO concentrations are higher (6 mgO$_2$$L^{-1}$ in dry season and 4 mgO$_2$$L^{-1}$ in rainy season, Figure \@ref(fig:C3-gas-profile)).

### Relationship of urban discharge, eutrophication and GHGs

Highly polluted water discharged in the river can promote eutrophication and biogeochemical processes responsible for GHGs production, such as mineralization of organic matter, denitrification, nitrification [@Brigham2019; @Cotovicz2021; @Li2021]. GHGs concentrations often peak at sampling sites located close to urban discharges. @Marescaux2018 also found a spatial gradient of GHGs in the downstream Seine River (France) and estuarine system under the impact of wastewater treatment plants effluents of Paris. Similarly, the urban discharge had a clear impact on the concentration of $\mathrm{CH_4}$ and $\mathrm{CO_2}$ in both seasons in the Saigon River. As shown in Figure \@ref(fig:C3-gas-profile), $\mathrm{CO_2}$ and $\mathrm{CH_4}$ concentrations in urban and suburban (10 km upstream of HCMC) were 5 and 14 times higher than in non-urban areas, respectively, while a factor of about three times was observed for $\mathrm{N_2O}$. This shows that urban rivers with high availability of organic matter and oxygen deficiency are hotspots for GHGs formation like $\mathrm{CO_2}$ and $\mathrm{CH_4}$. We found $\mathrm{CH_4}$ concentrations in the urban stretch of the Saigon River much higher than those measured in rural areas, i.e., upstream and downstream, as already reported for other highly urbanized watersheds [@Wang2018b].

Regarding $\mathrm{N_2O}$, its concentrations were related to urban domestic effluent of temperate rivers, together with $\mathrm{CO_2}$ and $\mathrm{CH_4}$, e.g., downstream from Paris effluents [@Garnier2009;@Garnier2013; @Nixon1995], or in the Meuse River [@Borges2018]. Several studies indicated that the eutrophic status might not have the same relationship with all the three GHGs [@Grasset2020; @Li2021]. For instance, $\mathrm{CH_4}$ and $\mathrm{CO_2}$ concentrations were reported as often coupled during phytoplankton blooms ($\mathrm{CH_4}$ increasing and $\mathrm{CO_2}$ decreasing), and $\mathrm{N_2O}$ would be decoupled from these others GHGs [@Bartosiewicz2021]. Regression models have been used to explore the effects of eutrophication and other environmental factors (e.g., nutrients) to explain patterns of GHGs concentrations [@Brigham2019; @Galantini2021]. Therefore, we applied multivariate statistical analysis (RDA and HP methods) on the environmental variables and GHGs to identify their relationship and assess the impact of urban discharge on GHGs production (Figure \@ref(fig:C3-RDA-HP)).

```{r C3-RDA-HP, echo=FALSE, out.width='100%', fig.cap='RDA and hierarchical partitioning analysis between explanatory variables (physio-chemical parameters, eutrophication) and responses (GHGs) in the Saigon River in (a) dry and (b) rainy season. The cosines of angles between vectors approximate their relationship (e.g., the approaching angles of 0o and 180o indicate positive and negative relationships). The vector with a longer length indicates a higher impact on the variation of the dataset.', fig.align='center'}
knitr::include_graphics("figures/C3-RDA-HP.pdf")
```

Overall, the ten parameters (physio-chemical parameters, nutrients, eutrophication indicators) accounted for about 80% of the variation of GHGs in both seasons. The two axes of RDA1 and RDA2 contributed nearly 70% to the interpretation of GHGs and about 30% of the unexplained variation (Figure \@ref(fig:C3-RDA-HP)). A stepwise selection was applied after running RDA to help remove non-significant explanatory variables. As a result, there were five explanatory variables, including DO, TSS, TOC, Chl-a and TRIX in the dry season. During the rainy season, the results of RDA determined the impact of DO, TSS, $\mathrm{NH_4^{+}}$, alkalinity and salinity on GHGs concentrations.

#### $\mathrm{N_2O}$ vs. environmental parameters

Several studies found that anthropogenic inputs (with high DIN concentration) correlated significantly with $\mathrm{N_2O}$ production (e.g., @Murray2015;  @Marescaux2018; @Burgos2017). These findings suggested that most eutrophic and polluted estuaries would probably experience an $\mathrm{N_2O}$ supersaturation [@Manjrekar2020]. However, there was no linear correlation between $\mathrm{N_2O}$ and $\mathrm{NH_4^{+}}$, $\mathrm{NO_{3}^{-}}$ in the Saigon River, which received a large number of nutrients from urban discharge.

Based on the RDA results, TOC, Chl-a and eutrophication index had strong relationships with the variation of $\mathrm{N_2O}$ concentration in dry season in the Saigon River (Figure \@ref(fig:C3-RDA-HP)a). High concentrations of pollutants and heavy eutrophication help explain such high concentrations of $\mathrm{N_2O}$. In rainy season, dissolved $\mathrm{N_2O}$ was not detected in both the urban area and downstream of Saigon River, despite the nutrient concentrations were even higher than in dry season (Figure \@ref(fig:C3-gas-profile)). During the rainy season, $\mathrm{N_2O}$ concentrations did not correlate with nutrient concentrations. Instead, alkalinity contributed to 53% to the variation of $\mathrm{N_2O}$. In dry period, the impact of urban wastewater (high TOC, low DO concentrations, eutrophication) corresponds to an increase in $\mathrm{N_2O}$ concentration in the receiving urban area, with alkalinity supporting ammonium oxidation, and hence $\mathrm{N_2O}$ production [@Peng2015]. The alkalinity concentration in dry season is higher than in rainy season, but in low concentration (50 mgCaCO3$L^{-1}$). Based on the statistical approach, the contribution of alkalinity was not enough to explain the significant decrease of $\mathrm{N_2O}$ from dry season to rainy season in urban area of Saigon River. Eutrophication was the factor having the most significant impact on $\mathrm{N_2O}$ in this study. TRIX or Chl-a was significantly correlated with $\mathrm{N_2O}$ during the dry season. From upstream (UP01) to urban area (SG05), Chl-a concentration increased steadily from 0.5 to 15 $\mu$g$L^{-1}$, $\mathrm{N_2O}$ concentration also increased from about 2 to 30 $\mu$gN $L^{-1}$ (Figure \@ref(fig:C3-gas-profile)). This result was similar to previous studies with the confirmation of the effect on eutrophication on $\mathrm{N_2O}$ production in rivers and estuaries; e.g, Rivière du Nord in Canada [@Galantini2021], Xiaoyue River in China [@Wang2020], or Elbe estuary in Germany [@Brase2017].

#### $\mathrm{CO_2}$ vs. environmental parameters

Studies dealing with the impact of urbanization on $\mathrm{CO_2}$ production shown that the $\mathrm{CO_2}$ concentration was higher in the rivers with higher urban coverage [@Tang2021]. In the Saigon River, $\mathrm{CO_2}$ concentrations were supersaturated in both seasons even if its concentration significantly decreased in the downstream section of the estuary (Figure \@ref(fig:C3-gas-profile)).

RDA and HP analysis results show that DO was linked to $\mathrm{CO_2}$ concentration in the Saigon River (>20% independent effects) in both seasons. The DO concentrations frequently ranging from moderate to severe hypoxic conditions (0.5 mg $L^{-1}$), while dissolved organic carbon (DOC) was maintained at a concentration of 3.6 $\pm$ 1.1 mgC$L^{-1}$. In other words, the Saigon River shown all the conditions (polluted and anoxic states, anaerobic/anoxic processes) for decomposing organic carbon and depleting oxygen and producing $\mathrm{CO_2}$ [@Nguyen2019]. In general, $\mathrm{CO_2}$ concentration in both seasons tended to decrease gradually along the estuarine gradient (Figure \@ref(fig:C3-gas-profile)) due to the dilution effect. However, compared to global atmospheric $\mathrm{CO_2}$ concentration, the $\mathrm{CO_2}$ downstream of Saigon River was 2-4 times higher (Figure \@ref(fig:C3-gas-profile)). The Can Gio mangrove at the interface of Saigon River and the coastal zone was considered as a source of $\mathrm{CO_2}$ to the atmosphere, with the highest emissions near the inflowing river. @David2018 concluded that the high $\mathrm{CO_2}$ emissions in this sector were due to high water temperature (26 to 31$^{\circ}$C), high POC content in TSS originating from upstream urban ar, and organic matter degradation in the mangrove.

The variation of $\mathrm{CO_2}$ was also linked to independent effects of TRIX (12%) and TOC (10%) in the dry season (Figure \@ref(fig:C3-RDA-HP)a). Although a negative relationship between Chl-a and $\mathrm{CO_2}$ was considered as uptake of $\mathrm{CO_2}$), in the Saigon River, the trophic index (TRIX) or Chl-a was positively correlated with $\mathrm{CO_2}$ in both seasons. This result was similar to another recent study in a subtropical lake in China [@Sun2021]. These results suggested that a large amount of $\mathrm{CO_2}$ concentrations was due to abundant organic carbon or to the contribution of lateral inputs, overwhelming the $\mathrm{CO_2}$ consumption by phytoplankton through photosynthesis. In rainy season, TSS, salinity, $\mathrm{NH_4^{+}}$ also accounted for 16%, 14%, 12% of independent effects on $\mathrm{CO_2}$ variation, respectively (Figure \@ref(fig:C3-RDA-HP)b). Although HP analysis determined that alkalinity and salinity in the dry season accounted for 15% of independent effects on $\mathrm{CO_2}$), the significant model selection by RDA indicated that they were non-significant explanatory variables in dry season. The POC content in TSS also shown a proportion many times higher in the upstream and urban area than in the downstream ones (Figure \@ref(fig:C3-gas-profile)), which supported the high $\mathrm{CO_2}$ concentrations in these two areas, with a gradual decrease towards the sea. However, the results of RDA, HP analysis, and the monitoring dataset still do not fully explain the high $\mathrm{CO_2}$ concentration in the upstream less polluted area. The upstream area of Saigon River may be affected by $\mathrm{CO_2}$ from allochthonous inputs. e.g., $\mathrm{CO_2}$)-rich sources in groundwater [@Marescaux2018].

Overall, the positive correlation between $\mathrm{CO_2}$ and the eutrophic status in the Saigon River indicated that respiration was significantly higher than photosynthesis, resulting in high $\mathrm{CO_2}$ emissions. The $\mathrm{CO_2}$ concentrations seemed to decrease only under the influence of dilution with coastal water along the estuarine gradient.

#### $\mathrm{CH_4}$ vs. environmental parameters

in the Saigon River, $\mathrm{CH_4}$ concentrations were much higher in urban than non-urban areas, as $\mathrm{N_2O}$. Besides, an interesting finding was the significantly higher $\mathrm{CH_4}$ concentrations in rainy than in dry season (Figure \@ref(fig:C3-gas-profile)). Figures 2 and 3 indicate that the water quality and eutrophication were worst in the extended urban area. The POC, TOC, and Chl-a had higher concentrations in these areas, which explained the higher concentration of $\mathrm{CH_4}$. Among the ten environmental variables explored, the contribution of Chl-a, TOC, DO and TRIX to $\mathrm{CH_4}$ variation in dry season was 23%, 16%, 15%, 14%, respectively, while TSS and nutrients (N, P) accounted for less than 3% (Figure \@ref(fig:C3-season)a).  @DelSontro2019 indicated that the $\mathrm{CH_4}$ emissions exponentially rose with the Chl-a concentration. A study in an urbanized tropical bay (Guanabara Bay, Brazil) found a high correlation between Chl-a and $\mathrm{CH_4}$ [@Cotovicz2016], confirming the high impact of urbanization on eutrophication and $\mathrm{CH_4}$ production in tropical regions.

The $\mathrm{CH_4}$ concentration in the urban section of Saigon River in the rainy season was much higher than in the dry season (Figure \@ref(fig:C3-gas-profile)). $\mathrm{NH_4^{+}}$ and alkalinity accounted for the largest independent effects on 48% and 17% of $\mathrm{CH_4}$ variation (Figure \@ref(fig:C3-RDA-HP)b). $\mathrm{CH_4}$ formation occurs mainly in anoxic/anaerobic conditions, when most oxidants (e.g., $\mathrm{O_{2}}$, $\mathrm{NO_{3}^{-}}$, Mn, $SO_4^{2-}$) are completely depleted (Likens 2010). The $\mathrm{NO_{3}^{-}}$ concentration remained in the range of 0.3 - 0.6 mgN$L^{-1}$ in the Saigon River in both seasons. Therefore, the emission of $\mathrm{CH_4}$ in the water column was probably negligible in comparison to $\mathrm{CH_4}$ production in sediment. Although Saigon River had high TOC concentration of about 6 mgC$L^{-1}$ and low DO (about 1 mgO$_2$$L^{-1}$) during the dry season, the $\mathrm{CH_4}$ concentration remained about 70 times lower than in the rainy season in the urban section. The lower $\mathrm{CH_4}$ in the dry season might be due to methanotrophy, which consumed $\mathrm{CH_4}$. During dry season, there was an increase in salinity in the urban area, leading to the generation of more electron acceptors and a consecutive decrease in methanogenesis. Lower $\mathrm{CH_4}$ concentrations in the dry period could also be attributed to high respiration rates at the roots of the floating aquatic plants (supported by $\mathrm{CH_4}$-oxidising bacteria) such as water hyacinths [@Attermeyer2016; @Nijman2021], while in the rainy season with stronger flushing capacity, macrophytes in the Saigon River were not as dense. The lower $\mathrm{CH_4}$ concentrations in the surface water covered here by water hyacinths in dry season were comparable to the open water areas in a tropical lake in India [@Attermeyer2016]. This study indicated that the coverage of water hyacinths only affected the $\mathrm{CH_4}$ concentration but did not strongly alter the $\mathrm{CO_2}$ concentrations. Besides, @Rosentreter2018 suggested that the distinct $\mathrm{CH_4}$ difference between the two seasons in tropical estuaries was contributed by flushing capacity. Stronger flushing capacity can promote remobilization of bottom organic matter as well as sediment-water process, which increases $\mathrm{CH_4}$ concentration. Indeed, the average discharge in rainy season is ten times higher than in the dry season in the Saigon River [@Nguyen2019]. This suggests that the flushing capacity in the rainy season is much higher than that in the dry season [@Nguyen2021].

## Conclusion of Chapter 3

This research provides comprehensive information about the impact of the urban discharge on eutrophication and GHGs emissions of an urbanized tropical estuary, the Saigon River Estuary. Average concentrations of $\mathrm{CO_2}$, $\mathrm{CH_4}$, and $\mathrm{N_2O}$ were higher than the global GHGs concentration from about 10 to 300 times. Based on global warming potential, the total GHGs concentration in the Saigon River was nearly 15,000 $\mu g$$\mathrm{CO_2}$-eq$L^{-1}$, in which $\mathrm{CO_2}$ contributes to 70-90% of total GHGs. While $\mathrm{CO_2}$ concentration did not differ between the two seasons, the concentration of $\mathrm{N_2O}$ and $\mathrm{CH_4}$ significantly differed between the dry and the rainy seasons. Based on the statistical analysis, the concentrations of three GHGs were linearly correlated with the eutrophication status of the estuary, especially in the dry season when phytoplankton was abundant. The analysis of AOU and $\mathrm{CO_2}$, $\mathrm{N_2O}$ excess allows identifying an important contribution of oxic respiration and nitrification and denitrification in the formation of these GHGs. GHGs fluxes in the Saigon River were comparable to other urbanized estuaries worldwide regardless of climatic conditions. This study suggests that effective control of urban wastewater will minimize eutrophication, which is also an effective solution to reduce the greenhouse effect from estuaries.
