---
output:
  #bookdown::html_document2: default
  #bookdown::word_document2: default
  bookdown::pdf_document2:
    template: templates/brief_template.tex
    citation_package: biblatex
documentclass: book
bibliography: [bibliography/references.bib]
---

```{r , include=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
```

# Biogeochemical functioning of an urbanized tropical estuary: Implementing the generic C-GEM (reactive transport) model {#steady-state-model} 
\chaptermark{Biogeochemical functioning of an urbanized estuary}
\minitoc <!-- this will include a mini table of contents-->


**Abstract**

Estuaries are amongst the most productive ecosystems of the land ocean continuum, but they are also under high anthropic pressures due to coastal urbanization. Too sparse observations have hindered the understanding of complex interactions between water quality and estuarine hydrodynamics and biogeochemical transformations. Until now, estuarine modelling studies have mainly focused on temperate estuarine systems in industrialized countries. This study investigates the responses of a tropical estuary to pollution load from a megacity (Ho Chi Minh City, Southern Vietnam) by applying a one-dimensional, biogeochemical estuarine model (C-GEM). The Saigon River Estuary flows through the megacity of Ho Chi Minh (HCMC) and is subject to episodic hypoxia events due to wastewater inputs from urban discharges. Good agreements are found between simulation outputs and observations for tidal propagation, salinity, total suspended sediment, and water quality variables in dry season in the Saigon River Estuary. C-GEM reproduces the increases in ammonium, total organic carbon, phytoplankton and dissolved oxygen depletion in the urban section of the Saigon River as an impact of untreated wastewaters from HCMC. The steady-state version of C-GEM also reveals the formation of a pollutant cloud (30-km stretch) resulting from the combined effects of tidal fluctuation and low flushing capacity during the dry season. Furthermore, the quantification of the reaction fluxes simulated by the model demonstrates that nitrification is the main process removing $\mathrm{NH_4^{+}}$ from the Saigon River. For the first time in such a type of environment, our study demonstrates the effectiveness of C-GEM at unraveling the complex interplay between biogeochemical reactions and transport in a tropical estuary with a minimized data requirement. This is significant for tropical estuaries in developing countries, where intensive monitoring programs are rare and have thus been rarely the object of modeling investigations.

*Keywords:* biogeochemical modelling, self-purification, Saigon River Estuary, Ho Chi Minh megacity, Vietnam

Published as: 

Nguyen, A. T., Némery, J., Gratiot, N., Garnier, J., Dao, T. S., Thieu, V., & Laruelle, G. G. (2021). Biogeochemical functioning of an urbanized tropical estuary: Implementing the generic C-GEM (reactive transport) model. _Science of the Total Environment_, _784_, 147261.


## Introduction

Estuaries are complex ecosystems under the simultaneous influences of tides and rivers, where nutrient loads are controlled by urban, industrial, and agricultural emissions. Although estuaries have self‐purification capacity supported by the tidal mixing of inland and coastal waters, excessive nutrients loads are likely to cause eutrophication [@Lanoux2013; @Paerl2006]. This phenomenon threatens water resources safety, especially for megacities in developing countries, where the balance between urbanization and environmental management is commonly a difficult tradeoff. In this context, reactive transport models can provide valuable insight to better understand the complex biogeochemical dynamics of such estuaries and help predict the potential response of water quality to anthropic perturbations. Under the pressure of multiple pollution sources (e.g., domestic wastewater, industrials inputs), the main factors controlling water quality in estuaries are biological and physical processes; in particular: (1) biogeochemical processes such as nitrification, denitrification, remineralization, primary production in water columns [@Hu2009; @Vanderborght2007; @Volta2016] as well as adsorption-desorption processes at the water-sediment interface [@Mortimer1999; @Nemery2007; @Vilmin2015]; (2) dilution process between riverine and marine waters [@Etcheber2011; @LajaunieSalla2019; @Lanoux2013]. Understanding these processes is often facilitated by using numerical models that usually require high-frequency data series (Tappin, 2002; Wild-Allen and Rayner, 2014).

Studies on the anthropogenic impact on water quality in estuaries by biogeochemical models are mainly available in temperate estuaries of industrialized countries [@Gawade2017; @Regnier2013].The Carbon Generic Estuarine Model (C-GEM) is a generic one-dimensional, reaction-transport model that takes advantage of the relationship between estuarine geometry and hydrodynamics to minimize data requirements [@Volta2014]. Steady-state simulations of C-GEM have provided accurate descriptions of estuarine hydrodynamics and biogeochemical transformations in several temperate estuaries [@Laruelle2017; @Laruelle2017; @Volta2014; @Volta2016 ]. The biogeochemical reaction network of C-GEM and its parameterization were built upon the study of more than 40 temperate estuaries @Volta2014; @Volta2016. Although several studies have been carried out in tropical estuaries through monitoring and experiments (e.g. @MiguelLucasLavoAntonioJimo2018, @Vipindas2018, @Wu2013), estuarine biogeochemical modelling applications are still missing in many countries in tropical regions such as Southeast Asia, India, Africa, South and Central America [@Volta2016 ; @Regnier2013]. There is now a need to improve this temperate parameterization of C-GEM to allow for the robust simulation of tropical estuaries. This is of particular importance for water resource management in urbanized tropical estuaries such as in the Saigon River Estuary, Vietnam [@Nguyen2019], Recife estuarine system, Brazil [@Noriega2013]; Cochin Estuary, India [@Gupta2009; @Vipindas2018]; Cross River estuary system, Nigeria [@Dan2019]; Cisadane estuary, Indonesia [@Dong2011]. The most notable difference between temperate and tropical systems is the rate of biological processes. The higher temperature of tropical estuaries often increases the biological uptake, excretion of nutrients and microbial activity (e.g., nitrification and denitrification) [@Eyre1999; @Tappin2002]. Another major difference between the two types of estuaries is the flow regime with a direct impact on the retention time of water bodies. In most tropical estuaries, two distinct flow rates and retention time are found during the dry and rainy seasons  [@Dong2011], whereas higher flow rates occur in winter in temperate estuaries. @Vipindas2018 suggested that nitrification rate can increase 10-40 folds when encountering a low flushing rate of tropical estuary during pre-monsoon. Similarly, oxidation rate of organic matter would not be significantly affected by temperature, but by the river discharge magnitude [@Gawade2017; @Wu2013b] . Therefore, the factors controlling the intensity of biogeochemical processes will differ between the temperate and tropical systems.

The Saigon River Estuary (Southern Vietnam) is a tropical estuary located in Southeast Asia. This estuary is considered as a natural collector for domestic wastewaters of about 9 million people living in Ho Chi Minh City (HCMC). Urban canals of HCMC and the Saigon River receive about 90% of untreated domestic wastewater from HCMC [@Nguyen2020]. The high polluted loads regularly lead to an excess of nutrients and intense periods of anoxia, especially during the dry season [@Nguyen2019]. There have been several studies in recent years showing the risk of eutrophication in this estuary, especially in the dry season because of low flushing capacity [@Nguyen2019;@Camenen2021;@Nguyen2021]. In addition, the research results of @Nguyen2019 showed that there was no significant difference between water quality variables in rainy season and dry season (except chlorophyll *a*). Besides that, the phytoplankton abundances were very low in rainy season, but more than 100 times higher in dry season which mainly concentrated in urban section of Saigon River [@Nguyen2021b]. Simultaneously, the risk of estuarine pollution is expected to increase in the coming years, under the impact of megacity’s development. The population of HCMC is indeed expected to reach 23 million by 2050, the Saigon River would then receive three times more pollution than what is currently released by the city [@Nguyen2020]. Understanding the self-purification capacity of this tropical estuary is important for HCMC’s development plans related to water resources safety.

This study aims to quantify the biogeochemical processes and tidal fluctuations controlling the water quality in a tropical estuary. We applied C-GEM to simulate the evolution of water quality over a tidal cycle representative of the dry season in the Saigon River Estuary during the 2014 – 2017 period, as a representative case study of an urban impacted tropical estuary. In the first part of the paper, the characteristics of the study area are described, including those of the estuary (Saigon River Estuary) and of the megacity (Ho Chi Minh City). The second part of the paper presents the protocol for implementing C-GEM model in this tropical estuary, in terms of description of data collection, the model constraints, the fundamental equations for building hydrodynamics, transport modules, biogeochemical reactions, and associated parameters that need to be calibrated. The last part of the paper deals with the quantification of interactions between estuaries and water quality including tidal effects and biogeochemical transformations. Intensity of biogeochemical reactions and fluxes are also calculated and compared with other studies in temperate and tropical estuaries.

## Materials and methods

### Study area

```{r C4-Map-Model, echo=FALSE, out.width='90%', fig.cap='Model domain of Saigon River Estuary and monitoring stations', fig.align='center'}
knitr::include_graphics("figures/C4-Map-Model.pdf")
```

The Saigon River Estuary system is located in Southern Vietnam. The river flows from its source in southeastern Cambodia to the Dau Tieng Reservoir in Vietnam. Dau Tieng Reservoir (270  $km^{2}$ and 1580x106 $m^{3}$) was designed for ﬂood control, domestic and industrial water demands and preventing the saltwater intrusion in the Saigon River [@Ngoc2014]. The Saigon River then flows through the natural boundaries of Tay Ninh province, Binh Duong province and Ho Chi Minh City. From Dau Tieng Reservoir to the estuary mouth, 200 km downstream, the Saigon River (18 $\pm$ 14 $m^{3}$  $s^{-1}$) joins in turn several notable tributaries such as the Thi Tinh River (20 $\pm$ 11 $m^{3}$  $s^{-1}$) and the Dongnai River (632 $\pm$ 446 $m^{3}$  $s^{-1}$), forming Nha Be River, which then splits into two river mouths (Soai Rap River and Long Tau River) flowing into the East Sea of Vietnam (Figure \@ref(fig:C4-Map-Model)). In addition, the Saigon River is connected to an urban river (Vam Thuat River, 4 $m^{3}$  $s^{-1}$) and three urban canals of HCMC (combined discharge of 5.5 $m^{3}$  $s^{-1}$) before the confluence with Dongnai River [@Nguyen2020].

The climate of this estuarine area is tropical monsoon, which has two distinct seasons (dry and rainy season), with a relatively constant temperature (about 29oC). The rainy season lasts from May to November, with an average annual rainfall of 1800 mm [@Nguyen2019]. Under the semi-diurnal tidal mixing, the saline intrusion diffuses up to 90 km upstream [@Aygun2019]. The tidal amplitude ranges from 1m during neap tide to about 4m during spring tide [@Schwarzer2016].

Land use in the watershed of the Saigon River significantly varies from North to South. Upstream of HCMC (North of the city), agricultural activities dominate with paddy rice and rubber tree farms. The central area hosts urban settlements and industrial zones belonging to HCMC. Downstream of the estuary, the Can Gio mangrove system is recognized as a biosphere reserve by UNESCO.

The status of water quality in the Saigon River strongly varies along its path. The water quality status at the upstream and downstream ends of the estuary are considered good, based on Vietnamese regulation on surface water quality [@Nguyen2019]. The water quality status of Saigon River becomes moderate to bad condition after connecting with the Vam Thuat River (an urban river, km 114) and urban canals of HCMC (km 130, @Nguyen2020). While nitrate ($\mathrm{NO_3^{-}}$) concentrations in canals are similar or lower than in the upstream stretch of Saigon River [@Strady2017], the ammonium ($\mathrm{NH_4^{+}}$) and total organic carbon (TOC) concentrations in those canals are many times higher than the threshold of 0.3 mg $\mathrm{NH_4^{+}}$ $L^{-1}$, 4 mg TOC $L^{-1}$ provided by the Vietnamese regulation for surface water – QCVN 08:2015/BTNMT [@Nguyen2020]. However, the concentration of these pollutants at the outlet of the canals decreases several folds before entering the Saigon River supported by three factors: firstly, the high retention and removal capacity of urban rivers and canals of HCMC, secondly, the mud dredging activities and finally, the dilution effect during flood tide [@Nguyen2020]. In the downstream section, the water quality of Saigon River Estuary improves markedly after joining the Dongnai River.

### Data collection

Bathymetry, freshwater inflow, water level, water quality and climate conditions were collected to implement C-CEM.

Mean depth and width for 83 cross-sections along the 200 km long Saigon River Estuary were extracted from bathymetry surveys performed in 2008 and 2016, provided by the Southern Institute of Water Resources Research (SIWRR, Vietnam). Climate condition data (sunshine hours and wind speed) were obtained from the statistical office of HCMC (www.pso.hochiminhcity.gov.vn).


```{r C4-table1-data, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Data collection for model implementation and validation (see Figure 2.21 for the locations of the hydrological and water quality sampling stations)") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH4-table1-Data collection for model implementation.pdf") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

Water quality data were obtained from the Center of Environment Monitoring (CEM, Vietnam) and the Centre Asiatique de Recherche sur l'Eau (CARE, Vietnam). CEM has monitored the water quality in the Saigon River Estuary (bi-weekly) and urban canals of HCMC (monthly) since 2005. The measured variables include temperature (ToC), salinity, total suspended sediment (TSS), dissolved oxygen (DO), ammonium ($\mathrm{NH_4^{+}}$), phosphate ($\mathrm{PO_4^{3-}}$), chemical oxygen demand (COD), biochemical oxygen demand ($\mathrm{BOD_5}$). The CARE laboratory has observed water quality at three locations along the Saigon River and one in the Dongnai River since July 2015 (Figure \@ref(fig:C4-Map-Model)). These locations are the same as those used by the CEM monitoring program, but CARE’s analyses include additional variables such as total nitrogen (TN), total phosphorus (TP), nitrate ($\mathrm{NO_3^{-}}$), total organic carbon (TOC), dissolved silica (DSi) and chlorophyll-a as described in @Nguyen2019.

Hydrological data (water level, velocity, discharge) were measured monthly at 10 stations along the Saigon and Dongnai rivers by the CEM. Once a month, the hydrological monitoring stations track the evolution of water level and velocity over 24 hours (2 tidal cycles) to calculate the net water discharge by removing the influence of tidal fluctuations in the Saigon River Estuary [@Camenen2021].

Implementing a water quality model for the assessment of daily to seasonal fluctuations in river flow was limited by the low frequency of hydrological measurements. The simulation of water quality for conditions representative of an average dry season is therefore a reasonable choice for our study. During the dry season, the upstream flow discharge is low and quite steady [@Camenen2021; @Nguyen2020]. In contrast, a lack of high-resolution data for precipitation, runoff and river discharge during the rainy season would have resulted in reliability issues of the modeling, especially in developing countries such as in Southeast Asia with a low monitoring effort. Besides, the high flow in the rainy season can create very short residence times and high flushing capacity, pollutants thus being mainly washed away from the estuary towards coastal waters.

Table \@ref(tab:C4-table1-data) summarizes the data used for the implementation and validation of the model for the dry season from 2014 to 2017 in the Saigon River Estuary. The 2014 - 2017 period has the most complete data set along the Saigon River and urban canals in this study.

### Model implementation

The generic estuarine model C-GEM was applied in the Saigon River Estuary to reproduce the nutrient dynamics over a generic tidal cycle representative of dry season conditions. The version of C-GEM used in this study was built upon the steady-state version described in @Volta2016 with some parameter values adjusted for tropical monsoon conditions in the Saigon River Estuary. The updated values of the modified parameters (discussed below) fall within the ranges provided by @Volta2016 based on a literature survey.

C-GEM’s application domain spans 200 km of the Saigon River Estuary, from the outlet of Dau Tieng Reservoir (km 0) to the estuary mouth (km 200) (Figure \@ref(fig:C4-Map-Model)). The three main modules of C-GEM (geometry module, hydrodynamics – transport module and biogeochemical module) were implemented following the set-up protocol proposed by @Volta2014. The partial differential equations governing the hydrodynamics module were solved using a finite difference scheme applied along a 1D grid, with a grid size of 2000 m and an integration time step of 300 seconds, insuring the stability of the numerical scheme. A operator-splitting method was used to solve the transport and biogeochemical reactions within a single time step [@Regnier1997]. C-GEM required a spin-up of 180 days to get into a steady-state regime for the Saigon River Estuary. A dynamic steady state was considered reached once two consecutive tidal cycles generated similar hydrological and biogeochemical results that could then be compared with observations.

```{r C4-Model-Depth, echo=FALSE, out.width='90%', fig.cap='Depth and Width of Saigon River Estuary', fig.align='center'}
knitr::include_graphics("figures/C4-Model-Depth.pdf")
```

#### Geometry module

Following Savenije’s work on alluvial estuaries [@Savenije2005], C-GEM uses an idealized representation of the estuarine geometry which can describe the estuarine width by an exponential equation along the estuary gradient.

$$\bar{B}\left(x\right)=B_0\cdot\exp{\left(-\frac{x}{b}\right)}$$

where x is the distance from estuary mouth [m], B0 is the width at estuary mouth [m], b is the width convergence length [m].



```{r C4-table2-physical, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Physical parameters to build the estuarine geometry and hydrodynamic module in the Saigon River Estuary") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH4-table2-Physical parameters to build the estuarine geometry.pdf") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```


C-GEM was designed to account for the interdependency between geometry and hydrodynamics in tidal alluvial estuaries. The width convergence length, which controls the system's shape, is thus a very sensitive parameter for hydrodynamics. Therefore, the value of the convergence length was carefully calibrated against observations (Figure \@ref(fig:C4-Model-Depth)). The impact of freshwater flow on this estuarine morphology was very different between the upstream (before Dongnai River confluence) and downstream sections [@Gugliotta2020]. The estuarine width ranges from 60 m from Dau Tieng Reservoir (km 0) to 350 m (km 140), and rapidly increases to 3690 m at the estuary mouth. This particular geometry justifies the use of a second convergence length to reproduce the entire width profile of the Saigon River Estuary. C-GEM geometry used a short convergence length from the estuary mouth (km 200) to the confluence of the Saigon and Dongnai rivers (km 142). From this confluence to the upstream section (km 142 – km 0), the width of the river only decreases slowly and gradually (with longer convergence length) (Figure \@ref(fig:C4-Model-Depth) and Table \@ref(tab:C4-table2-physical)). Such strategy was previously used with C-GEM for applications to other estuaries [@Laruelle2019;@Volta2016b].

The depth of the Saigon River Estuary gradually increases from 2.5 m at the outlet of Dau Tieng reservoir to 19 m at km 60. The depth of the river is then almost constant over the next 80 km (i.e. until the confluence with the Dongnai River), it then decreases gradually to 9.6 m at the estuary mouth. The idealized geometry in the model captured well the mean estuarine depth and width which were extracted from 83 cross-sections of the bathymetry.

#### Hydrodynamics and transport module

The hydrodynamics of the 200 km long reach along Saigon River Estuary was described by the one-dimensional hydrodynamic module of C-GEM based on the continuity and cross-sectional integrated momentum equations [@Nihoul1976b; @Regnier1998; @Regnier1999]:

$$r_s\frac{\partial A}{\partial t}+\frac{\partial Q}{\partial x}=0$$

$$\frac{\partial U}{\partial t}+U\frac{\partial U}{\partial x}=-g\frac{\partial\xi}{\partial x}-g\frac{U\left|U\right|}{C_h^2H}$$

where$ x$ is distance along the longitudinal axis [m],$ t$ represents time [s], $rs$ is the dimensionless storage water ratio, $A$ is the cross-sectional area [$m^{2}$], $H$ is the instantaneous water depth [m], $Q$ is the cross-sectional discharge [$m^{3}$  $s^{-1}$] and Ch is the Chézy coefficient [$m^{1/2}$  $s^{-1}$].


```{r C4-table3-sediment, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Determination of sediment parameters and adjustment of biological parameters for application of the C-GEM model in the Saigon River Estuary") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH4-table3-Determination of sediment parameters and adjustment.pdf") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

Water elevation at estuary mouth and freshwater discharge at the upstream limit of the model domain were required to solve the hydrodynamics module. Parameters related to the characteristics of the tidal amplitude and period are also required to calculate the water elevation at the estuary mouth to constrain the downstream boundary. Table \@ref(tab:C4-table2-physical) summarizes the parameters used to solve these equations for the Saigon River Estuary. Chézy coefficient was the only parameter being fixed for the calibration of the hydrodynamic module, as it was not measured in the Saigon River Estuary. Chézy coefficient typically ranges from 40 – 60 $m^{1/2}$  $s^{-1}$ along alluvial estuaries [@Savenije2001]. The Chézy coefficient was calibrated based on the comparison between simulated tidal amplitude profiles and mean tidal observations at seven stations along Saigon River Estuary. Calibrated results showed that the Chézy coefficient applied Saigon Estuary ranged from 15 to 60 $m^{1/2}$  $s^{-1}$. The upper reaches of the Saigon River were much rougher than other alluvial estuaries. The sudden change of topography of upstream Saigon River Estuary compared to downstream probably explains this larger difference between Chézy coefficients in these two areas than the values reported by @Savenije2001.

After validating the hydrodynamics module, the transport module was implemented to reproduce the dynamics of dissolved variables (e.g., ammonium, dissolved oxygen) and total suspended sediment. The advection–dispersion equation for C (x, t) given by:

$$\frac{\partial C}{\partial t}+\frac{Q}{A}\frac{\partial C}{\partial x}=\left(AD\frac{\partial D}{\partial x}\right)+P$$

where $C$ is the concentration of solutes or Total Suspended Sediments (TSS), $D$ is the effective dispersion coefficient [$m^{2}$  $s^{-1}$] which is automatically calculated following the geometry of the system, P is the net biogeochemical processes related to the solutes or solid C(x,t).

```{r C4-Model-range, echo=FALSE, out.width='80%', fig.cap='Comparison between observed (red marks) and simulated (red lines) profiles for average tidal amplitude and salinity along the Saigon River Estuary. The vertical gray lines correspond to tributaries or canals. Standard deviations (grey area) of simulation come from hourly fluctuations over tidal cycles.', fig.align='center'}
knitr::include_graphics("figures/C4-Model-range.pdf")
```


The transport module was validated by comparing salinity simulation and observation during the dry season 2014 - 2017. Salinity was used because it is not affected by biogeochemical processes (variable P = 0). In addition, salinity distributions reflect the combined processes such as advection-dispersion and mixing processes along Saigon River Estuary. Salinity simulation results for the 2014-2017 dry season in the Saigon River are depicted in Figure \@ref(fig:C4-Model-range). The salinity distribution agreement between simulation and observation has ensured the precise transport of solutes along the Saigon River over tidal cycles.

#### Biogeochemical module

The biogeochemical module implemented in C-GEM in this study allows assessing the concentrations of eight state variables, namely ammonium ($\mathrm{NH_4^{+}}$), nitrate ($\mathrm{NO_3^{-}}$), phosphate ($\mathrm{PO_4^{3-}}$), total organic carbon (TOC), Silica (DSi), dissolved oxygen (DO), phytoplankton (diatoms and non-diatoms) and TSS. Seven biogeochemical processes are simulated in C-GEM (Figure \@ref(fig:C4-C-GEM)), including oxygen exchange through the air-water interface, aerobic degradation (organic carbon mineralization), nitrification, denitrification, primary production, phytoplankton mortality and TSS erosion/deposition [@Volta2014; @Volta2016].

```{r C4-C-GEM, echo=FALSE, out.width='100%', fig.cap='The state variables (rectangle) and processes (ellipse) in the biogeochemical module of C-GEM applied in the Saigon River Estuary. Red and green lines represent the consumption and production processes', fig.align='center'}
knitr::include_graphics("figures/C4-C-GEM.pdf")
```

C-GEM uses eighteen biogeochemical parameters to calculate the intensity of biogeochemical processes. Their determination is based on a comprehensive literature review of 49 estuarine biogeochemical model applications in temperate regions [@Volta2016]. The primary production process corresponds to the growth of phytoplankton and oxygen production through the uptake of C, N, P, and Si in the water column under soluble form depending on environmental forcings, especially light penetration. The light availability in the water column is affected by TSS, which is calculated by the model through the simulation of erosion/deposition processes. Nitrification consumes DO in the water column, which potentially eliminates a high concentration of ammonium. Denitrification contributes to anaerobic TOC consumption, $\mathrm{PO_4^{3-}}$ production and nitrogen elimination by converting $\mathrm{NO_3^{-}}$ into the inert $\mathrm{N_2}$gas that escapes from the water column into the atmosphere. Mineralization of organic matter (aerobic degradation) consumes TOC and $\mathrm{O_{2}}$ while producing $\mathrm{PO_4^{3-}}$ and $\mathrm{NH_4^{+}}$. Currently the application of C-GEM to shallow, pristine estuarine systems might be critical, when processes are dominated by intense element recycling within the sediments. The current lack of some N processes such as anammox does likely not significantly alter our result because these processes are relatively minor contributors to the nitrogen cycle in tropical estuaries [@Dong2011]. However, the $\mathrm{NO_3^{-}}$ simulation might affected by dissimilatory nitrate reduction to ammonium process in tropical estuaries with low $\mathrm{NO_3^{-}}$ concentration [@Dong2011]. Phosphate release from bottom sediment might have a high impact on the phosphorus simulation at shallow estuaries [@Vilmin2015]. Note that the modular structure of C-GEM allows for a relatively simple implementation of new processes in its biogeochemical module and increasing the complexity of the N and P cycles could be the subject of future studies.

In our simulations, seven of the eighteen biogeochemical parameters used by C-GEM were modified to suit the tropical conditions in the Saigon River Estuary (Table \@ref(tab:C4-table3-sediment)). These reaction rate constants (except aerobic degradation rate constant) are mostly higher than the C-GEM applications for temperate estuaries. In particular, nitrification, denitrification rate constants are 2-12 times higher. However, these parameters were in the ranges reported by @Volta2016b. In addition, they are similar to findings of biogeochemical reactions in tropical estuaries reported by @Miranda2008 or @Vipindas2018. The eleven other biogeochemical parameters were similar, including phytoplankton maintenance rate constant (4.11 x 10-7  $s^{-1}$), phytoplankton excretion constant (0.05), phytoplankton growth constant (0.29), Redfield ratio (C:Si:N:P = 106:15:16:1), Michaelis – Menten constant terms for dissolved silica (1.07), phosphate (0.2), ammonium (228.9), nitrate (26.07), organic carbon (186.25), oxygen in aerobic degradation (31.0), oxygen in nitrification (51.25), dissolved nitrogen (1.13).

The TSS concentration has a direct impact on light extinction and then on phytoplankton development. The sediment parameters (critical shear stress and erosion coefficient) were thus determined before adjusting biological ones though a calibration process only involving hydrological parameters and SPM concentration profiles.


```{r C4-table4-BC, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Boundary conditions used for the steady-state simulation of water quality variables in the Saigon River Estuary 2014-2017") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH4-table4-Boundary conditions for the steady-state.pdf") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

The boundary conditions of variables concentrations were the mean values of observations from 2014 – 2017 in dry season (Table \@ref(tab:C4-table4-BC)).

#### Model evaluation

The model calibration and validation processes were quantitatively evaluated using statistical criteria (Table \@ref(tab:C5-table5-eq)). The parameters used for the evaluation included the coefficient of determination ($\mathrm{R^2}$), root mean square error (RMSE) and the bias percentage (pbias). The performance of hydrodynamics module was assessed by comparing the mean values of simulated variables to observations at eight hydrological stations. The simulation of water quality variables was compared with the mean data at 14 water quality monitoring stations along the Saigon River Estuary (Figure \@ref(fig:C4-Map-Model)).


```{r C5-table5-eq, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Equations, ranges, and optimal values for assessing model performance") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH4-table5-Equations ranges optimal values for assessing.pdf") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

## Results and discussion

### Model simulations

Simulation results included hourly water depth, flow velocity, concentrations of eight water quality variables and intensity of biogeochemical processes over generic tidal cycles along the Saigon River Estuary. Table \@ref(tab:C4-table6-perform) summarizes the statistical comparison between simulated and observed variables during the dry season.

```{r C4-table6-perform, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Model performance statistics in the Saigon River Estuary during an average dry season for the 2014-2017 period") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH4-table6-Model performance statistics in Saigon River Estuary.pdf") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

Simulations using the hydrodynamics and transport modules were only evaluated using average tidal amplitude and salinity profiles collected during the average dry season from 2014 to 2017 (Figure 4). The model demonstrated realistic tidal amplitudes within the Saigon Estuary throughout its entire length ($\mathrm{R^2}$> 0.9, pbias < 3%, Table \@ref(tab:C4-table6-perform)).

The statistical comparison between simulated and observed salinity showed that the model properly captures salinity profiles along the estuary with a $\mathrm{R^2}$ of 0.98. The salinity distribution reveals a strong saline intrusion in the Saigon River which is consistent with previous literature [@Aygun2019]. Salinity gradually decreases from 28 to 0.4 at 120 km from the estuary mouth. Salinity strongly drops (from 28.0 - 5.0) within the first 30 km, then gradually decreases to almost zero at km 120 within HCMC.

```{r C4-Model-quality, echo=FALSE, out.width='100%', fig.cap='Comparison of observed and simulated results of water quality along the Saigon River Estuary over a tidal cycle. The vertical lines are tributaries or canals. Lateral input are described by vertical lines. The black dotted lines are the maximum permissible limit for domestic use, aquatic life based on Vietnamese regulation on surface water quality – QCVN 08:2015/BTNMT.', fig.align='center'}
knitr::include_graphics("figures/C4-Model-quality.pdf")
```

Our results reveal that the model properly reproduces concentrations and spatial trends for most of the biogeochemical state variables over the generic tidal cycles in the dry season. The simulations of TSS, DO, $\mathrm{NH_4^{+}}$, TOC reveal a pretty good agreement with observations based on model performance statistics (Table \@ref(tab:C4-table6-perform)) with $\mathrm{R^2}$> 0.7 and percent bias < 20%. $\mathrm{NO_3^{-}}$ and DSi also fall within the range of observations at three sampling stations along the Saigon River. Phytoplankton is rather well represented by the model despite a high pbias of 178% (Figure \@ref(fig:C4-Model-quality)f). $\mathrm{PO_4^{3-}}$ is the only variable not properly captured by C-GEM (pbias = 39%), but the simulated profiles still remain well within the standard deviations of the observations (Figure \@ref(fig:C4-Model-quality)e).

In the upstream section of the estuary (km 0 to 60-80), most variables are low and almost constant, except DO and DSi (Figure \@ref(fig:C4-Model-quality)). Water quality in this area is highly dependent on water discharge from Dau Tieng Reservoir which is rated in good status according to Vietnamese regulation on surface water quality (QCVN 08: 2015/BTNMT) [@Nguyen2019]. The sudden increase in DSi concentration from km 60 to 80 originates from the higher DSi concentration of TT river tributary (Table \@ref(tab:C4-table4-BC)). The simulation overestimates $\mathrm{PO_4^{3-}}$ concentration and underestimates TOC concentration. The underestimation of TOC may first stem from the intensity of the TOC degradation process, but also from TOC inputs to the system which may not be well considered. $\mathrm{PO_4^{3-}}$ overestimation can be explained by the lack of an important process in C-GEM, such as adsorption within the water column or in the sediment because the current version of the model does not have an explicit benthic module. This assumption is consistent with the study of @Vilmin2015 who confirmed that the concentration of $\mathrm{PO_4^{3-}}$ in estuaries is very sensitive to the adsorption of $\mathrm{PO_4^{3-}}$ onto the surface sediment. The concentration of $\mathrm{PO_4^{3-}}$ can indeed be reduced by 50% in the presence of adsorption in estuaries [@Vilmin2015]. Using a modelling approach, @Hu2009 indicated that $\mathrm{PO_4^{3-}}$ adsorption can account for about 99% of total $\mathrm{PO_4^{3-}}$ removal in the Pearl River, a subtropical estuary in China. @Nguyen2019 also reported that the Saigon River exhibits a high potential of phosphate adsorption onto the sediments ranging from 1 to 64%. However, the current water quality monitoring programs in the Saigon River (by CEM and CARE laboratory) mainly analyze $\mathrm{PO_4^{3-}}$, without measuring particulate inorganic phosphorus (PIP) in routine. $\mathrm{PO_4^{3-}}$ and PIP are indeed two necessary indicators for building the representation of sorption processes [@Billen2007; @Vilmin2015]. The addition of PIP measurements in future monitoring programs can thus support the improvement of C-GEM.

In the urban section (km 80 to 140), the Saigon River Estuary's water quality, which receives domestic wastewater through the connected urban canals of HCMC, was clearly different from the upstream section. The concentrations of $\mathrm{NH_4^{+}}$ and TOC in the canal were much higher than in the upstream area [@Nguyen2020; @Strady2017]. Both simulations and observation data confirm that water quality has been significantly deteriorated in this area. The model successfully simulates DO, $\mathrm{NH_4^{+}}$, $\mathrm{NO_3^{-}}$, TOC, DSi and phytoplankton concentrations in this section. Most of the variables impacted by the urban canal’s inputs present concentrations higher than those simulated upstream in the Saigon River. $\mathrm{NH_4^{+}}$ concentrations brought by the urban river (VT) and canals are ten times higher than those in the upstream sector. The $\mathrm{NO_3^{-}}$ concentrations from urban canals are not significantly different from the upstream concentration (\~0.4 mg $L^{-1}$), and the contribution of the urban VT river was fixed to 0.71 mg $L^{-1}$ (Table \@ref(tab:C4-table4-BC)). The increase in $\mathrm{NO_3^{-}}$ in this urban section reaching more than 1 mg $L^{-1}$ would be thus linked to biogeochemical processes in the water column rather than the $\mathrm{NO_3^{-}}$ anthropogenic contribution. In our simulations, $\mathrm{PO_4^{3-}}$ concentrations are overestimated (Figure \@ref(fig:C4-Model-quality)f) which could likely be attributed to an inefficient $\mathrm{PO_4^{3-}}$ uptake by the phytoplankton to compensate the higher load of $\mathrm{PO_4^{3-}}$ brought by the urban VT river and $\mathrm{PO_4^{3-}}$ production from aerobic degradation and denitrification processes. The decrease of DSi and increase of phytoplankton biomass were well simulated in this section and can be attributed to diatom growth. The abrupt decrease in phytoplankton and a slight increase in DSi between km 130 and 140 (Figure \@ref(fig:C4-Model-quality)f,h) corresponds to a dilution effect resulting from the mixing with the Dongnai River.

Downstream of the confluence with the Dongnai River (km 140-160 to 200), the concentrations of $\mathrm{NH_4^{+}}$, TOC, phytoplankton and DSi gradually decreased towards the estuary mouth (Figure \@ref(fig:C4-Model-quality)). Most of the variables in this area meet the Vietnamese regulation QCVN 08: 2015/BTNMT on surface water quality for domestic use and aquatic life. The water quality in the downstream section of Saigon River Estuary largely depends on the Dongnai River confluence's water quality, its discharge being about 10 times larger than that of Saigon River (392.9 $m^{3}$ $s^{-1}$ against 17.6 $m^{3}$ $s^{-1}$). Therefore, the concentrations of the simulated variables at this confluence (km 140) are almost similar to those from the Dongnai tributary, except TSS that increases from 100 to 180 mg $L^{-1}$ from km 140 to km 180, in dry season (Figure \@ref(fig:C4-Model-quality)g). This increase in TSS concentration is not caused by the Dongnai River, but instead by sediment erosion and resuspension (see section 3.3). Further, the gradually decreasing concentration of TSS from the 180 km to the estuary mouth is due to the dilution with coastal water.

### Tidal effects on pollutant clouds

Saigon River Estuary is dominated by a semi-diurnal tidal regime from the East Vietnam Sea which changes four times a day. Sea water can penetrate inland to more than 120 km during the dry season. Although tides can improve water quality through the mixing of urban and saline waters, the tidal fluctuation jointly diffuses the highly polluted sources into a pollutant cloud along the estuary.

Our simulation over 24-hours (two tidal cycles) shows that there a pollutant cloud of about 30 km remains centered around kilometer 120 regardless of the time of day with low DO (< 2 mg$L^{-1}$) (Figure \@ref(fig:C4-tide)a). At 0h (in a generic day), the contaminated area is located from km 108 to km 138, while it roughly expends from kilometers 98 to 128 after six hours (Figure \@ref(fig:C4-tide)a). These upstream and downstream movements are controlled by the flood and ebb tides in the Saigon River Estuary. During the dry season, tidal oscillation dominates because the freshwater discharge from upstream the river is very small (few tens of $m^{3}$ $s^{-1}$). From the outlet of the urban discharge (km 114) to the confluence with the Dongnai River (140 km), the residual discharge at the Saigon River is about 30 $m^{3}$  $s^{-1}$ during the dry season. A rough estimate of the residual velocity for a segment of 26 km (km 114 to 140 km) using an average cross-section of 5700 $m^{2}$ (width x depth: 300m x 19m) yields a value of 0.45 km $d^{-1}$.


```{r C4-tide, echo=FALSE, out.width='100%', fig.cap='The tidal effect on the transport of the low DO concentration cloud along Saigon River Estuary. (a) DO concentration along Saigon River Estuary, (b) water particle displacement from km 108 during two tidal cycles. The displacement is calculated by multiplying velocity (m $s^{-1}$) and each hour (3600s). This displacement is only considered by the velocity (advection) of the water particle.', fig.align='center'}
knitr::include_graphics("figures/C4-tide.pdf")
```

As a result of our simulations, a water particle could travel a distance of 10 km over six hours during flood tide (Figure \@ref(fig:C4-tide)b). It would then almost return to its original position after six hours of receding tide. The maximum tidal velocity at km 108 could reach 0.8 m  $s^{-1}$ during flooding tide. The residence time of the water body in this urban section was thus estimated to be around 57 days. Low discharge not only makes pollutants unable to flush out during the dry season, but it also increases the risk of eutrophication. The long residence times of water with high concentrations of nutrients facilitate algae's growth in this section before being flushed out of the estuary. As mentioned previously by @Nguyen2019, high concentrations of chlorophyll-a as a eutrophication indicator are only evidenced in the dry season in the urban section of Saigon River Estuary. Similarly, oxygen depletion (DO< 2 mg$L^{-1}$) was also observed in the Gironde Estuary (a west European estuary) during summer with a low river discharge and long residence time of water (from 20 to 86 days) [@Lanoux2013]. However, urban pollution discharges in most temperate European countries have been decreasing since the 2000’s European Union directive, so that low summer water flow is nowadays less degraded [@Romero2016]. In tropical estuaries, despite high pollution levels during the dry season, much larger water flows during the rainy season usually lead to a strong flushing which helps the water quality in estuaries to be reset every year [@Eyre1999].

### Processes and flux budgets

#### Intensity of biogeochemical processes

While tidal mixing is the main control of the model’s state variables concentrations in the downstream section of the estuary, the biogeochemical processes play a much larger role in regulating the biogeochemical dynamics of the system in its upstream and urban sections, especially in the pollutant cloud (Figure \@ref(fig:C4-process)). Once evaluated, C-GEM allows to quantify the intensity of each process controlling the consumption or production of oxygen, nutrients, and organic carbon along the estuary, together with the physical processes such as erosion/deposition of TSS and oxygen exchange. In general, the Saigon River Estuary tends to (i) consume $\mathrm{O_{2}}$ in the urban section and $\mathrm{NH_4^{+}}$, DSi, TOC in the entire estuary; (ii) produce DO in the upstream and downstream sections, TSS in both urban and downstream sections and $\mathrm{NO_3^{-}}$, $\mathrm{PO_4^{3-}}$ and phytoplankton in the whole estuary.

##### Erosion/deposition of TSS

A change in TSS in the urban section and downstream of the confluence with the Dongnai River was simulated (Figure \@ref(fig:C4-process)g). The sudden shallowness in river depth combined with the increase in river width at the Dongnai River confluence would promote erosion process in this section, as previously observed by @Gugliotta2020. This erosion phenomenon was also observed in the coastal Can Gio mangrove (downstream of the Saigon River) which eroded rapidly within four years of observation (2014 – 2017) [@Nguyen2019c]. @Nguyen2019c observed that the erosion—deposition intensity in the dry season was stronger than in the rainy season because of the stronger waves in dry season. The suspended sediment was influenced strongly by the currents in the dry season and had a strong tidal asymmetry. Therefore, the impact of hydrodynamics in the dry season were stronger than the ones in the rainy season[@Nguyen2019c].

##### Oxygen exchange

Oxygen exchange (between surface water and atmosphere) and net primary production (NPP) are the two processes that provide oxygen to the estuary in C-GEM. The oxygen exchange at the air-water interface strongly contributes to the oxygenation of the water column in the Saigon River Estuary while most of other processes (except for NPP) consume oxygen. From upstream to downstream of the Saigon River, the oxygen exchange produces between 4.0 and 8.3 $\mu$M $\mathrm{O_{2}}$ $d^{-1}$ (Figure \@ref(fig:C4-process)a). This amount of oxygen introduces into the water through the air-water interface is quite similar to the amount of oxygen consumed by TOC aerobic degradation. Consumption of oxygen by nitrification process appears slightly smaller in the upstream and downstream sections, but significantly increases in the urban sector. Oxygen production by phytoplankton also increases in the urban section but remains negligible in the upstream and downstream sections compared to the other oxygen production/consumption processes.

```{r C4-process, echo=FALSE, out.width='110%', fig.cap='Intensity of biogeochemical processes and erosion/deposition rates affecting the concentrations of (a) DO, (b) TOC, (c) $\\mathrm{NH_4^{+}}$, (d) $\\mathrm{NO_3^{-}}$, (e) $\\mathrm{PO_4^{3-}}$, (f) Phytoplankton, (g) TSS and (h) DSi. For example, Oxygen exchange, Net primary production and nitrification are affecting DO concentrations. The dashed lines are the net biogeochemical reactions.', fig.align='center'}
knitr::include_graphics("figures/C4-process.pdf")
```

##### Nitrification and denitrification

Nitrification primarily leads to DO depletion and $\mathrm{NH_4^{+}}$ consumption in the Saigon River. The nitrification process is particularly intense in the urban section (from km 98 to km 138). Nitrification could convert a maximum of 9.3 $\mu$M N $d^{-1}$ of $\mathrm{NH_4^{+}}$ into $\mathrm{NO_3^{-}}$ at km 130 (Figure \@ref(fig:C4-process)c, d). In the urban section, $\mathrm{NH_4^{+}}$ consumption by nitrification surpasses the $\mathrm{NH_4^{+}}$ production by TOC degradation. This contributes to explain the decrease of $\mathrm{NH_4^{+}}$ upstream and downstream of the Saigon River. The simulated nitrification rate in the Saigon River is similar to observations at Kochi estuary (0 - 4 $\mu$M N $d^{-1}$), a tropical monsoon estuary in India which receives untreated sewage from urban activities [@Miranda2008]. Further, in another study in the Cochin estuary (southeast Arabian Sea), @Vipindas2018 observed particularly high nitrification rates in the dry season (about 10 $\mu$M N $d^{-1}$) because of the low flushing rate of the estuary, a situation also met in the Saigon estuary.

Denitrification is a process controlling the concentration of $\mathrm{NO_3^{-}}$ by transforming $\mathrm{NO_3^{-}}$ into atmospheric gaseous $\mathrm{N_2}$ (and possibly $\mathrm{N_2O}$, as an intermediate product, although this step is not explicitly modelled by C-GEM) under anoxic conditions. In addition, denitrification forms $\mathrm{PO_4^{3-}}$ under anoxic conditions @Arndt2011 which partly explains the increase of $\mathrm{PO_4^{3-}}$ in the urban section of the Saigon River. Besides, under anoxic condition, a possible release of bottom $\mathrm{PO_4^{3-}}$ through sediment desorption might occur [@Nemery2007; @Vilmin2015]. The simulated denitrification rate in the Saigon River reached a maximum of 3.3 $\mu$M N $d^{-1}$ in the urban section. The intensity of the removal of $\mathrm{NO_3^{-}}$ through denitrification is only one-third of $\mathrm{NO_3^{-}}$ formation through nitrification (Figure \@ref(fig:C4-process)d). This rate is nonetheless higher than what was reported experimentally in some of other tropical regions such as the Barra Bonita Reservoir (1.36–1.77 $\mu$M N $d^{-1}$) [@Abe2003]. According to this study, denitrification processes in tropical conditions often occur with a high intensity because of the fast DO consumption (aerobic respiration) by heterotrophic bacteria using the high labile organic carbon. Such an environmental setting is consistent with the Saigon River's conditions in the urban section where low oxygen concentrations are frequent during the dry season (\~ 1.0 mg$L^{-1}$).

##### Aerobic degradation of organic carbon (aerobic respiration)

The simulated aerobic degradation rate in the Saigon River Estuary ranges from 5 to 6.5 $\mu$M C $d^{-1}$. In addition, in the urban section, the intensity of aerobic degradation does not appear to be as high as other processes when receiving contaminated water from urban canals. This could be explained by the low availability of DO in this section (\~1.0 mg$L^{-1}$, Figure \@ref(fig:C4-tide)) that makes the TOC degradation preferentially anaerobic (i.e., denitrification) as observed in Figure \@ref(fig:C4-process)b. The sum of TOC consumption by the aerobic degradation and denitrification almost equals the TOC production by phytoplankton mortality (8.5 $\mu$M C $d^{-1}$) in urban section. Therefore, a biogeochemical balance is reached in this section which does not reduce TOC concentration. However, TOC's net biogeochemical reactions are negative in upstream and downstream. This would lead to a decrease in TOC concentrations in these two sections. However, according to the observations, the TOC concentration from reservoir to km 80 is almost constant. Other sources of TOC in the upstream section can be the cause of TOC compensation due to the aerobic degradation process. According to a well-water quality survey from 233 wells in agricultural, urban and industrial areas of HCMC, the wells with high concentration of organic carbon are only observed in agricultural areas (upstream of Saigon River) [@Nguyen2020b]. This suggests that agricultural activities in the upstream Saigon River can release high TOC content, affecting groundwater and surface water in this area.

Compared to previous biogeochemical applications of C-GEM [@Volta2016], the aerobic degradation constant rate applied in our study was set at the low end of the range of possible values for this parameter (Table \@ref(tab:C4-table3-sediment)). Such low value strongly diverges with some studies performed in highly urbanized estuaries such as the Recife estuarine system (Brazil). @Noriega2013 also argued that the respiration process was the dominant process in a highly urbanized estuary which was the cause of the constant low oxygen concentration. Unlike the Recife estuarine system, the low oxygen concentrations in the Saigon River would be due to the nitrification of $\mathrm{NH_4^{+}}$ brought by urban canals, as shown by increasing $\mathrm{NO_3^{-}}$ concentration (from about 0.5 to 1.1 mg$L^{-1}$) in the urban section. The TOC degradation might strongly occur in the canal system with a long residence time before discharging to the Saigon River. We assumed that the less labile organic carbon was mainly stay in mainstream of Saigon River while the labile organic carbon was already degraded in the inner canal system. Based on C-GEM application in typical temperate estuaries [@Volta2016], the aerobic degradation rate could reach 20 $\mu$M C $d^{-1}$, which was about two times higher than in our study. Although aerobic degradation was reported to be the main factor leading to oxygen depletion in many studies carried out in other systems, nitrification was the main factor in this tropical estuary.

##### Phytoplankton primary production (NPP)

NPP uses nutrients and produces $\mathrm{O_{2}}$. In the Saigon River, NPP only plays a minor role in eliminating nutrient pollution but still contributes to the oxygen regeneration in the system. In the urban section, oxygen seems to be a scarce element because of its consumption through nitrification and aerobic degradation. However, the intensity of oxygen production through NPP is similar to the oxygen exchange in the urban area (Figure \@ref(fig:C4-process)a), but is negligible upstream and downstream due to low phytoplankton levels. While $\mathrm{NH_4^{+}}$ is consumed by nitrification and NPP, $\mathrm{PO_4^{3-}}$ can only be consumed through primary production in C-GEM. Besides, the production of $\mathrm{PO_4^{3-}}$ by aerobic degradation and denitrification is higher than the $\mathrm{PO_4^{3-}}$ uptake associated with primary production in the Saigon River's urban section, a positive net biogeochemical reaction that led to an increase in $\mathrm{PO_4^{3-}}$ in this section (Figure \@ref(fig:C4-process)e).

NPP is the only process in C-GEM that consumes silica. Silica is a naturally occurring element essential for the development of diatoms which account for about 90% of the total phytoplankton observed in the Saigon River in the dry season [@Nguyen2021b]. According to the Redfield stoichiometric ratio in the Saigon River, DSi was always in excess over nitrogen and phosphorus for diatom's development [@Nguyen2019] and its availability is thus never a limiting factor to diatom’s growth currently. DSi concentrations however, are relatively low in the upstream section (Figure \@ref(fig:C4-Model-quality)h, km 0 to km 60), but they are 2-3 times higher in the urban section (km 80 to km 160), where diatom frustule dissolution might occur in this highly productive section, accentuated by e.g., resuspension by navigation, feeding in return diatom uptake, feeding in return diatom uptake. In addition, the increase in DSi could come from high concentration of DSi of tributaries (see Table \@ref(tab:C4-table4-BC)) and groundwater (not taken as an input by C-GEM). Indeed, according to @Oehler2019, DSi concentration in the water column can be particularly elevated during the ebb tide by the contribution of DSi in groundwater which is greater than in surface water. Further, results of interaction analysis between groundwater and river showed that Saigon River indeed received water from groundwater at upstream and contributed to groundwater at downstream [@Van2018]. In the following downstream section, the concentration of DSi then decreases because of dilution with the sea and the Dongnai River (Figure \@ref(fig:C4-Model-quality)h) . There might be a risk of DSi depletion and the formation of harmful non-diatom phytoplankton if the pollutant cloud (from km 98 to km 138, Figure \@ref(fig:C4-tide)a) further spreads towards the upstream section. The decrease in freshwater inflow and the increase in tidal amplitude could indeed move the pollutant cloud upstream (Park and Song, 2018). However, the freshwater discharge from Dau Tieng Lake (upstream of Saigon River Estuary) is usually maintained at a stable level during the dry season [@Ngoc2014].

##### Comparison with temperate estuaries

The results of our simulations in the Saigon River Estuary highlight some differences between a tropical estuary and other temperate estuaries. C-GEM application for three idealized temperate estuaries (marine estuary, riverine estuary, mixed estuary) showed that $\mathrm{NH_4^{+}}$ consumption rates ranged from 0.2 to 0.3 $\mu$M N $d^{-1}$ [@Volta2016] which is about 30 times smaller than what was simulated in the Saigon Estuary. This is likely the consequence of higher $\mathrm{NH_4^{+}}$ concentrations and higher temperatures that facilitate the nitrification process in the Saigon River. Indeed, the average temperature of this tropical estuary falls within the optimal temperature range for the growth of nitrifying bacteria (25-35$^{\circ}$C) [@Eyre1999]. Besides, the biogeochemical models for typical temperate estuaries showed that denitrification rates reached a maximum of about 1.5 $\mu$M N $d^{-1}$ [@Volta2016], which is two times lower than in the Saigon River. In tropical regions, denitrification is usually much more intense than in areas with relatively lower temperatures, such as temperate regions [@Abe2003].

#### Biogeochemical reaction fluxes

This section assesses the self-purification capacity of this tropical estuary by calculating mainstream fluxes and reaction fluxes of $\mathrm{O_{2}}$, TOC, and $\mathrm{NH_4^{+}}$ in urban and downstream sections of Saigon River Estuary (Figure \@ref(fig:C4-flux)).


```{r C4-flux, echo=FALSE, out.width='100%', fig.cap='           Fluxes and reaction fluxes of (a) DO, (b) TOC, (c) $\\mathrm{NH_4^{+}}$ in urban and downstream sections of Saigon River Estuary in dry season 2014 – 2017. Process rates are in tons', fig.align='center'}
knitr::include_graphics("figures/C4-flux.pdf")
```


In the Saigon River, about 90% of the nutrients from untreated wastewaters are retained in sludge within the canal network of HCMC and partly lost (denitrified into N2) [@Nguyen2020]. These polluted sludges are occasionally removed through canal dredging programs. However, at the output of canals, about 10% of the wastewater still has a very high concentration in nutrients and organic matter, well beyond the thresholds recommended by the Vietnamese regulation on surface water quality QCVN 08:2015/BTNMT.

In the urban section of the Saigon River Estuary, the average fluxes of DO, TOC, and $\mathrm{NH_4^{+}}$ were 8.1 ton$\mathrm{O_{2}}$ $d^{-1}$, 15.5 tonC $d^{-1}$ and 2.2 tonN $d^{-1}$, respectively. Biogeochemical reaction fluxes remove 12% of DO, 1% of TOC and 18% of $\mathrm{NH_4^{+}}$ within the urban section. These results show that the Saigon River Estuary is more effective at transforming/removing nutrients than organic carbon. Nitrification plays a key role in the self-purification of the Saigon River Estuary (Figure \@ref(fig:C4-flux)c). Phytoplankton's contribution to the removal of nutrients is negligible, but NPP contributes to the oxygenation of water in the urban section. Nitrification consumes about 2 ton$\mathrm{O_{2}}$ $d^{-1}$, while NPP provides about 0.75 ton$\mathrm{O_{2}}$ $d^{-1}$. The high abundance of phytoplankton facilitates an increase of TOC through the phytoplankton mortality process. The phytoplankton mortality flux in the urban section reaches 0.30 tonC $d^{-1}$, while the aerobic degradation only removed 0.27 tonC$d^{-1}$ (Figure \@ref(fig:C4-flux)b). This makes the Saigon River Estuary an inefficient TOC filter, which is a strong difference with the behavior of several temperate systems [@Laruelle2017; @Laruelle2019;@Volta2016b] calculated using C-GEM.

In the downstream section, the flux calculation mainly depends on the Dongnai River's fluxes, with a much higher discharge than in the Saigon River. The reaction fluxes are negligible and only reduce the downstream fluxes by 2 - 4%. As discussed earlier, the concentrations of water quality variables in the downstream section of the Saigon River are almost similar to those in the Dongnai River with a good water quality status. Therefore, the biogeochemical reactions in this section are weaker than in the urban section. The most notable difference between these sections is the dominant biogeochemical processes. While nitrification has a strong impact on DO in the urban section, this process is much weaker in the downstream section. Oxygen supply through gas exchange (7.25 ton$\mathrm{O_{2}}$ $d^{-1}$) plays a key role in improving DO in this section (Figure \@ref(fig:C4-flux)a), compensating for the consumption taking place upstream. Besides, TOC degradation is also larger than the TOC produced by phytoplankton mortality. However, the overall net biogeochemical flux is not significant compared to the TOC flux downstream of the Saigon River (Figure \@ref(fig:C4-flux)b).

Compared to other studies, the self-purification capacity of the Saigon River based on biogeochemical processes is relatively small. For example, the Scheldt Estuary (Belgium, Netherlands) in the 1990s received high concentrations of lateral inputs and could consumed 73% $\mathrm{NH_4^{+}}$, 78% TOC and almost all $\mathrm{O_{2}}$ [@Vanderborght2007]. Another modelling study conducted in the Pearl River Estuary, a sub-tropical estuary in China, one of the most densely populated and economically developed regions of the country, showed that the biogeochemical processes could eliminate 50% BOD and 37% of $\mathrm{NH_4^{+}}$ in the estuary [@Hu2009]. The lower reduction capacity in the Saigon River Estuary can be explained by the high retention or removal of pollutants in the inner canal system of HCMC. While the Pearl River Estuary or Scheldt Estuary can remove high amounts of both organic carbon and $\mathrm{NH_4^{+}}$, in the Saigon River Estuary, $\mathrm{NH_4^{+}}$ was mostly left for removal. The high concentration of TOC created by phytoplankton mortality process, mostly biodegradable, would effectively fuel aerobic degradation process within water column and TOC removal. This would mean that urban TOC discharged in the mainstream would be refractory, the biodegradable fraction of TOC being directly consumed in the canals. As a result, the urban canal and river (VT) network within the city plays a role of pretreatment, mineralizing first TOC and lowering oxygen, before the urban effluents reach the river itself where ammonia could find enough oxygen for nitrification.

Overall, our simulations suggest that the urban section of Saigon River mainly plays a role in retaining/eliminating macropollutants, which can be explained by the low flushing and long residence time, especially during the dry season. In the downstream section, the large discharge from the Dongnai River increases the capacity of the system to flush out pollutants into the coastal waters. In this downstream area, biogeochemical reactions then only play a minor role (2-4%) compared to hydro-physical controls.

## Conclusion

Water quality in tropical estuaries is often under great pressure from urban and industrial wastewater sources, especially in developing countries. The lack of extensive monitoring programs in such regions makes difficulties in developing or applying water quality models for water management in these estuaries. In this study, the application of a generic reaction-transport model (C-GEM) allows to describe the dynamics of biogeochemical reactions and tidal effects on the water quality in a tropical estuary (the Saigon River Estuary, Vietnam). The simulation results capture the movement of a pollutant cloud in this estuary which confirms the impact of megapole wastewater on water quality. During the dry season, this cloud (30km stretch) moves up and down within a radius of 10km within the urban section of Saigon River Estuary, under the combined influences of the tidal effect and low discharge. In addition, a long residence time and high nutrient concentrations facilitate the development of phytoplankton which rapidly grows in such conditions. C-GEM allows quantifying the intensity of the different biogeochemical processes controlling nutrient dynamics in the system. Nitrification plays an important role in the Saigon River Estuary, removing about 18% of $\mathrm{NH_4^{+}}$ flux in the urban section. The contribution of phytoplankton growth to the consumption of nutrients was negligible, but significant to the improvement of DO in a water column. Unlike many studies in other estuaries, the Saigon River Estuary only removed TOC originating from phytoplankton. Untreated ammonia-rich effluents were not diluted in the dry season, which indeed favored nitrification. This study indicated that the optimal temperature and abundant nutrients created favorable conditions for the growth of nitrifying bacteria in tropical estuaries. This study highlighted some fundamental differences between sanitation conditions in high- and low-income countries, the latter showing higher vulnerability as furthermore located in tropical zones.

**Acknowledgements**

This study was conducted at the International Joint Laboratory Le CARE and funded by the NUTRIM project EC2CO Bioheffect Structurante Initiative (2017-2018). We acknowledge the Center of Monitoring Of the Department of Natural Resource and Environment for the water quality data. Goulven G. Laruelle is a research associate of the F.R.S-FNRS at the Université Libre de Bruxelles.
