---
output:
  #bookdown::html_document2: default
  bookdown::word_document2: default
  bookdown::pdf_document2: 
    template: templates/brief_template.tex
    citation_package: biblatex
documentclass: book
bibliography: [bibliography/references.bib]
---

# Modeling the seasonal nutrient dynamics and phytoplankton development in the Saigon River tropical estuary
\chaptermark{Modeling the seasonal nutrient dynamics}
\minitoc <!-- this will include a mini **Table** of contents-->

```{r , include=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
```

\noindent Eutrophication in urbanized tropical estuaries of emerging countries is increasing. The understanding of eutrophication dynamics in the Saigon River Estuary has not been achieved because of the sparse observations and consequent lack of biogeochemical modeling application. The distinct differences in phytoplankton densities and eutrophication risk between dry and rainy seasons were implicitly interpreted through multivariate data analysis (Chapter 3). The C-GEM model (steady-state version) described the strong nitrification, denitrification processes, and phytoplankton development during the dry season (Chapter 4). In this chapter, the transient version of the C-GEM model is extended to describe the seasonal evolution of nutrients and phytoplankton under megacity's influence in the Saigon River Estuary from 2017 - 2018. The model was validated for hydrodynamics and water quality variations of longitudinal profiles in two seasons. C-GEM captured the evolution trend of organic carbon and phytoplankton over these two years. The temporal variation is still not achieved for some variables such as phosphate and ammonium upstream and downstream of the Saigon River. The simulated biogeochemical processes and hydrological effects have elucidated the mechanism of phytoplankton succession. The high nutrient availability in the urban section and low upstream and downstream explains for the spatial phytoplankton differences. The distinct seasonal residence time explains the high and low phytoplankton densities in the dry and rainy seasons, respectively. The advantages and disadvantages of C-GEM application for estuaries in developing countries are then discussed to evaluate the effectiveness and significance of the model in supporting tools for water quality management and reducing the risk of eutrophication in urbanized tropical estuaries (Chapter 6). This chapter follows the format for further submission in a scientific journal.

## Introduction

### Risk of phytoplankton blooms at urbanized tropical estuaries

Seasonal variation of phytoplankton has been observed in tropical estuaries, with phytoplankton densities usually very high in the dry season and low in the rainy season [@Nguyen2021b; @Duong2019b; @Damar2019]. Climate change makes the hydrological difference between the two seasons more obvious in tropical estuaries [@Eyre1999; @Tappin2002]. Besides, the rapid urbanization and ineffective water management practices lead to the risk of water pollution in rivers, estuaries and coastal areas. Indeed, the megacities are mainly located in the coastal areas supported by the abundant provision of ecosystems, maritime services [@Vitousek2008; @Hoornweg2017]. The estuaries are also known as the natural treatment plant for urban, agricultural and industrial wastewaters [@Howarth2011; @Billen2011]. Thus, increasing the risk of eutrophication in the dry season in emerging countries is unavoidable.

In contrast to megacities in developed countries, population growth and domestic wastewater volume have increased rapidly in recent years in developing countries, but wastewater treatment systems have not kept pace with urbanization [@Lee2016]. Eutrophication in urbanized estuaries is increasingly reported. For example, urban wastewater of the megacity of Ho Chi Minh City (Vietnam) pollutes the Saigon River Estuary [@Nguyen2019; @Nguyen2021]. Widespread strong hypoxia in near-bottom waters was detected in over half of the Gulf of Thailand during the dry season in 2015 [@Buranapratheprat2021]. In Jakarta Bay (Indonesia), urban wastewater increases eutrophication which has affected the aquatic environment, and the trend of eutrophication is increasing over the 13-year survey 2001–2013 [@Damar2019]. Other tropical countries such as Guanabara Bay (Brazil) and Ivory Coast (Africa) also observed an increase in the risk of eutrophication under the pressure of urbanization [@CarvalhoAguiar2011; @Droge2007]. Currently, 75% of 20 megacities (more than 10M inhabitants) worldwide are located in coastal areas [@WorldOceanReview2017]. It is estimated that in 2050, 122 cities with more than 5M inhabitants will be mostly in coastal areas [@Hoornweg2017]. This estimation warns that the impact of megacities on coastal ecosystems will continue to increase in the coming decades.

### The complex interaction between phytoplankton and controlling factors

Understanding the response mechanism of estuaries to wastewater emissions on phytoplankton dynamics is difficult because of the complex processes of estuaries and their interplay [@Gypens2013; @Lancelot2011]. Tidal estuaries have a large capacity to receive and dilute wastewater with tides and seawater [@Billen2011]. The drivers to the seasonal variation of phytoplankton characterization (abundance, type of species) at estuaries are affected by nutrient limitations, solar radiation, the residence time of the water body [@Billen1994; @Lancelot2011; @Arndt2011; @Dijkstra2019]. Temperature and tidal effects also cause phytoplankton fluctuations in estuarine systems [@Bianchi2007]. Several studies suggest that residence time is the key factor that directly affects the density of phytoplankton [ @Lancelot2011]. For estuaries with long residence time and weak flushing capacity, phytoplankton has enough time to grow before being flushed out of the system. In contrast, in estuaries with short residence and strong flushing capacity, phytoplankton is often pushed out of the system before rapidly growing even with high nutrient concentration. Indeed, the phytoplankton densities only increase when net specific growth rates exceed the residence time of the water [@MiguelLucasLavoAntonioJimo2018]. In tropical estuaries, the causes leading to differences in eutrophication in estuaries often stem from fluctuations in hydrology and weather conditions [@Damar2019; @Buranapratheprat2021]. @Lancelot2011 indicated that subtropical and tropical estuaries often observed a significant difference in phytoplankton densities between dry and rainy seasons due to the intense seasonal monsoons and the occurrence of droughts. However, some studies reported that residence times did not greatly affect phytoplankton densities and species, but rather light penetration intensity influenced by suspended sediments and nutrient availability [@Hoang2018]. In upstream of the estuaries, light availability is generally better than downstream because of shallow water and active sediment resuspension in the mouth of estuaries. Nutrient limitation is also crucial in the studies of phytoplankton development [@Costa2009;@Smith2006; @Arndt2011]. Rivers and estuaries controlled by dam discharge often reduce silica formation, changing the nutrients ratio [@Likens2010].

The preference for the statistical interpretation of the controlling factors is often reported in tropical estuaries. Some results are contradictory even under the same climatic conditions. For instance, at two contrast interpretations for the seasonal phytoplankton difference at Red River Delta (Vietnam) by @Hoang2018 and @Duong2019. The difference interpretation can be due to the nature of the wastewater source with different C, N, P, Si ratios, or the difference of external forcings such as tidal fluctuations, upstream flow control such as by dams. The C:N:P:Si ratio changes can lead to a huge difference in phytoplankton development. For instance, North Atlantic spring bloom in 1989 showed that dissolved silicate (DSi) was rapidly depleted to a minimum before nitrate was depleted. This coincided with a shift in dominant phytoplankton from diatoms to small flagellates [@Sieracki1993]. The deficiency of Si is mainly due to the consumption during diatom bloom. The addition of phytoplankton from natural processes such as weathering is much slower than DSi consumption or dissolution of diatom shells [@Garnier2010; @Arndt2007]. In estuaries, the evolution of phytoplankton species is more complex than in freshwater environments such as rivers and lakes because of the transition between freshwater and saltwater [@Eyre2000]. It seems to be that diatoms are regulated/ interfered by salinity rather than dilution. Freshwater species cannot grow in saline conditions; in contrast, marine diatoms cannot develop in low brackish/ freshwater conditions.

### Estuarine biogeochemical modeling in emerging countries

The modeling application can support the understanding on phytoplankton variation (species and density, biovolume/biomass), along with the dynamics of nutrients (C, N, P, Si), hydrological conditions (e.g., discharge, flow velocity, tides, precipitation), climate (e.g., temperature, light, wind). Theoretically, phytoplankton formation and conversion mechanisms are generally similar across environments such as tropical, temperate regions  [@LeMoal2019]. However, the growth rate and nutrient consumption of phytoplankton are often different in different climates. In the survey of about 40 biogeochemical model applications, the kinetics rates applied to different models often have large fluctuations [@Volta2016]. Even within an estuary, the coefficients applied have varied depending on the model. The model selection and calibration process are thus critical to ensure the effective application and minimize the misjudgement of management solutions. In emerging countries, extensive monitoring programs are rarely implemented. Choosing a suitable model is currently difficult because the over-simplified models do not explain the complex processes in the estuaries.

For example, in Gulf of Mexico, @Turner2014 argued the importance of zooplankton in describing nutrients and phytoplankton dynamics in modeling application. In addition, benthic-pelagic processes can significantly contribute to phytoplankton bloom dynamics in the water column [@Arndt2007]. The application of water quality models to urbanized estuaries of emerging countries is increasingly required to ensure the sustainable development of the region. Saigon River Estuary is a tropical estuary in Southern Vietnam heavily influenced by domestic wastewater from megacities. Similar to other tropical estuaries of developing countries, the application of water quality modeling in assessing eutrophication in this estuary is still limited. In well-mixed estuaries, without the significant contribution of benthic processes, the 1D models allow to reduce the data but still ensure the interpretation of biogeochemical processes in the water column such as Carbon-Generic Estuarine Model (C-GEM). C-GEM is an optimized model to minimize data requirements by taking advantage of the relationship between estuarine geometry and hydrodynamics while ensuring that important biogeochemical processes in the water column and hydrodynamics of estuaries [@Volta2014; @Volta2016]. C-GEM has been applied in several temperate estuaries in Western Europe [@Volta2014; @Volta2016;@Laruelle2019], North America [@Laruelle2019], and recently in a tropical estuary (Saigon River Estuary, Vietnam, @Nguyen2021). Based on the results of C-GEM simulation for the dry season in the Saigon River Estuary, the water quality in this estuary is affected by domestic wastewater from Ho Chi Minh Megacity. However, the steady-state version of C-GEM did not characterize the significant differences in phytoplankton between the rainy and dry seasons. Similarly, the eutrophication risk of tropical estuaries also varies depending on hydrological conditions. The survey results in 2016-2017 on the density of phytoplankton at Saigon River Estuary showed a clear difference between the dry and rainy seasons, with more than 100 times higher density in dry season [@Nguyen2021b]. In 2019-2020, four longitudinal profiles were carried out along the Saigon River Estuary (two times in the rainy season, two times in the dry season). There was a clear difference in the phytoplankton biomass between the two seasons and between the monitoring areas. Therefore, to understand the temporal fluctuations of phytoplankton biomass, the application of the steady-state version of C-GEM is no longer appropriate. The application of the transient version of C-GEM allows the description of these processes on a daily, seasonal and annual resolution with input data requirements consistent with existing data in the Saigon River Estuary.

### Objectives

This study investigates spatio-temporal patterns of nutrients, phytoplankton biomass (chlorophyll-a), organic carbon, nutrients, suspended sediment in the Saigon River Estuary by applying the C-GEM. The objectives are (i) to reproduce the variation of water quality and phytoplankton from 2017 to 2018, (ii) to quantify nutrient cycles and phytoplankton development under hydrological influences and urban wastewater discharge, and (iii) to identify the drivers of increased eutrophication in this urbanized tropical estuary.

## Materials and methods

A detailed description of the study area is presented in Chapter 2 with information about Saigon River Estuary and Ho Chi Minh Megacity. This section only briefly summarizes the research area and presents the difference between applying transient version versus steady-version of the C-GEM model for Saigon River Estuary.

### Study area

### Saigon River Estuary
Saigon River Estuary is located in the south of Vietnam, with a catchment area of 4500 $km^{2}$. From Dau Tieng Reservoir to the estuary mouth, the Saigon River has a length of about 200km. The Saigon River has an average discharge of 50 $m^{3}$ $s^{-1}$ (30$m^{3}$ $s^{-1}$ in dry season and more than 100 $m^{3}$ $s^{-1}$ in rainy season). It confluences with Thi Tinh River (10 and 30$m^{3}$ $s^{-1}$ in dry and rainy season) and Dongnai River (600 and 1200$m^{3}$ $s^{-1}$ in dry and rainy season) at km 40 and km 130, respectively. At km 110 and 126, the Saigon River receives wastewater from Vam Thuat River – urban river (4$m^{3}$ $s^{-1}$) and urban canals (5$m^{3}$ $s^{-1}$). The urban rivers and canals discharge are calculated based on the total amount of domestic wastewater accounted for the province population (approximately 150L/day/inhabitant). The Saigon River Estuary has a funnel shape, with a width of about 5000 m at the estuary mouth and about 300 m at the urban section (from km 80 to km 160) and 50 m at the upstream. Saigon River Estuary is deepest in the urban section with a maximum depth of 20m. At upstream and downstream, there are depths of 5 - 10 m. This estuary is influenced by the semi-diurnal asymmetric tidal regime of the East Sea of Vietnam, with a tidal period of 12.42h and a tidal range of 3m. Salinity at the estuary mouth ranges from 20 to 30 ppt, and can spread about 100km inland in the dry season and about 40km (km 160) in the rainy season. The residence time at the urban section of Saigon River Estuary in the dry and rainy seasons is about 80-120 days and 20-40 days, respectively calculated based on volume/freshwater inflow (Salman, 2021).


```{r C5-Map, echo=FALSE, out.width='100%', fig.cap='The model domain of C-GEM, water quality stations in the Saigon River Estuary. Monthly discharge of Saigon River (at km 130) and Dongnai River are average data from 2014 – 2017. CARE and CEM are two monitoring and research centers, described in section 2.2', fig.align='center'}
knitr::include_graphics("figures/C5-Map.pdf")
```

### Ho Chi Minh Megacity

Most of the length of the Saigon River lies within the territory of Ho Chi Minh City (HCMC). HCMC is the largest economic center of Vietnam, contributing 22% of GDP (in 2020) and having the highest population density in the country (4292 inhabitants$km^{-2}$, 9% national population in 2020). Measures to improve environmental quality in Ho Chi Minh City are not at the same rate as the population growth of HCMC. In 2008, HCMC built the first wastewater treatment plant to meet 10% of the wastewater of about 6.8 million people. By 2018, Ho Chi Minh City had three WWTPs, but only treated about 18% of wastewater from about 8.7 million people. Currently, one WWTP is under construction, while other WWTPs are still in the fundraising stage. More than 80% of urban domestic wastewater from Ho Chi Minh City is discharged directly into the canal system of HCMC, and finally to the Saigon River Estuary. The outlet of these canals belongs to km 120 - 130 of the Saigon River, about 80 km from the estuary mouth. Urban canals have very high concentrations of nutrients, much further beyond the regulated threshold of recommendation of surface water quality in Vietnam (QCVN 08:2008/BTNMT), and higher than concentrations in upstream of the Saigon River Estuary (Figure \@ref(fig:C5-C-GEM)). The phenomenon of mass fish deaths has appeared regularly in urban canals from 2014 – 2021. For example, from April to July 2014 (14 tons of dead fish, [Saigon News, 2014](http://saigonnews.vn/doi-song/108826-ca-chet-hang-loat-tren-kenh-nhieu-loc-thi-nghe.html)), 19/5/2015 (25 tons of dead fish, [Thanhnien News, 2015](https://thanhnien.vn/ca-chet-hang-loat-tren-kenh-nhieu-loc-thi-nghe-post462426.html)), 18/5/2016 (70 tons of dead fish, source: [Vietnamnet News, 2016](http://english.vietnamnet.vn/fms/society/156776/over-70-tons-of-dead-fish-removed-from-saigon-canal.html)), 4-7/4/2017 ([Vietnamnet News, 2017](https://vietnamnet.vn/vn/thoi-su/ca-chet-hang-loat-tren-kenh-nhieu-loc-thi-nghe-364991.html#:~:text=V%E1%BB%9Bt%20g%E1%BA%A7n%2070%20t%E1%BA%A5n%20x%C3%A1c,r%C3%A1c%20%C4%90a%20Ph%C6%B0%E1%BB%9Bc%20ti%C3%AAu%20h%E1%BB%A7y.)), 3/5/2018 ([Saigon Online, 2018](https://www.sggp.org.vn/ca-chet-hang-loat-tren-kenh-nhieu-loc-thi-nghe-516679.html)). From 2019 to 2020, there was no mass fish death, but April 4, 2021, there was a fish death like in previous years (Tuoitre News, 2021). The phenomenon of mass fish deaths in urban canals of HCMC has the common feature that all occur in the dry season, around April-May. However, there is no mass fish death in the Saigon River, where urban canals' wastewater is received, although the Saigon River is often in hypoxia condition and high nutrient concentrations.

### Data collection

Data used for the transient version of C-GEM includes hydrological, water quality, weather condition data. The data are mainly extracted from Vietnamese Center of Environmnetal Monitoring (CEM) and Center Asiatique de Recherche sur l'Eau (CARE). In general, the input data for the model is bi-weekly or monthly monitoring; only tidal elevation and solar radiation are hourly data.

#### Hydrological data

Hydrological data is collected from CEM's monthly measurements. C-GEM model only requires average discharge at the upstream boundary and generic tidal variation at the downstream boundary. The steady-state version of C-GEM, using a generic tidal variation, which is calculated based on the tidal period and tidal range at the estuary mouth. The steady-state C-GEM can only access the dynamics of processes within a tidal cycle. The changes in tidal forcings can affect the contaminant's spread to the upstream area, where domestic water production of HCMC is located. Therefore, the use of hourly tidal elevation would allow a more accurate depiction of tidal variation. The hourly tidal elevation at the estuary mouth is extracted from the global tide gauge data of UHSLC (<https://uhslc.soest.hawaii.edu/datainfo/>, accessed 10/02/2021). CEM monitoring program only provides water level, velocity, and discharge for 48 hours (i.e. 4 tides cycles) of each month. The average discharge in upstream and tributaries are the average of monthly measurement from 2014 to 2017. For model validation, the simulated tidal range are compared to the daily observed tidal range at three stations provided Vietnamese Center of Water Management and Climate Change (WACC) with 1.5 years (1/2017 to 6/2018) of available data.

#### Water quality data

CEM water quality data included concentrations of $\mathrm{BOD_5}$, $\mathrm{PO_4^{3-}}$, $\mathrm{NH_4^{+}}$, DO, and TSS. CARE data were used in addition (Chl-a, TOC, DSi and $\mathrm{NO_3^{-}}$). C-GEM allows for diatoms and non-diatom simulations. The results of the phytoplankton densities survey 2016-2017 showed that diatom accounted for 90% of total phytoplankton density [@Nguyen2021]. In addition, Chl-a was significantly correlated with total phytoplankton density, Chl-a was thus used instead of biomass of diatoms (90% Chl-a) and non-diatom (10% Chl-a). Currently, CARE only observes at four stations (Figure \@ref(fig:C5-Map)), so the data for boundary conditions are supplemented from the longitudinal profile along the Saigon River during the dry and rainy seasons 2019-2020. The survey results showed that the concentrations of TSS, $\mathrm{NH_4^{+}}$, $\mathrm{NO_3^{-}}$, $\mathrm{PO_4^{3-}}$, Chl-a, TOC, DSi in the upstream boundary were not significantly different from the monitoring station at km 76. Therefore, the data of Chl-a, DSi and $\mathrm{NO_3^{-}}$ at the upper boundary (km 0) and the lower boundary (km 200) is applied according to the average data of the four surveys. We assumed that the complex evolution of the variable concentrations in the estuary mainly originates from the interactions of biogeochemical processes and hydrology and external forcings (temperature, solar radiation, wind) rather than from changes in inflow water and saltwater.

#### Weather condition

Climatic conditions data applied to the C-GEM model include solar radiation, wind speed at 10m, and water temperature. The solar radiation was extracted from the data of a weather station of CARE located in HCMC, about 10 km from Saigon River Estuary and applied to the entire estuary. The monthly wind speed is collected from the HCMC Statistical Office data and applied to the whole estuary. Water temperature was used from CARE's bi-weekly monitoring survey.

### Model implementation

The set-up procedure for the transient version C-GEM model in this study is similar to those reported by @Nguyen2021 for the steady-state version, including building idealized geometry and hydrodynamics for Saigon River Estuary. The notable difference in this version is the addition of the P adsorption/desorption process to the biogeochemical module and the use of monthly input data. Calibration to determine the parameters for geometry (convergent length), hydrodynamics (Chézy), sediments (critical shear stress and erosion/deposition coefficient) was performed as in steady-state version. For biogeochemical parameters, three parameters, including aerobic degradation rate, phytoplankton mortality rate, were modified according to the range recommended by [@Volta2016]. Calibration data includes bi-weekly data at three stations PC, BD, BK for the period 11/2017 - 05/2018, a period in which there was no marked change in the concentration of Chl-a. The choice of this calibration period is to minimize the error in overfitting the model because the model's input data is still mostly monthly (e.g., discharge, external loadings from tributaries, canals).

#### Biogeochemical module

The variables and biogeochemical processes applied in the C-GEM model for Saigon River Estuary are shown in Figure \@ref(fig:C5-C-GEM). The processes and their interpretations are shown in Table \@ref(tab:C5-table1-reaction). Compared with the steady-state version [@Nguyen2021], the maximum mortality rate in this version increases from 2%/day to 9%/day. This change stems from solar radiation during 2017-2018, about three times higher than the calculated average values in the steady-state version, based on 9 hours of sunshine duration per day at 60% cloud coverage.

The P sorption process was supplemented because P was identified as a limiting factor for eutrophication at the Saigon River Estuary [@Nguyen2020]. Deficiency of this process is likely to lead to overestimated $\mathrm{PO_4^{3-}}$ concentration and phytoplankton biomass concentration and ultimately to Si depletion leading to a loss of predominance of diatom, which was not found at Saigon River Estuary. The P sorption process includes the adsorption/desorption $\mathrm{PO_4^{3-}}$ onto/from surface suspended sediment, which depends on the concentrations of TSS, $\mathrm{PO_4^{3-}}$, particulate inorganic phosphate (PIP). The simulated $\mathrm{PO_4^{3-}}$ is calculated by on sorption equation based on Langmuir equilibrium. This process requires two equilibrium parameters maximal P sorption capacity (Pac) and half-saturation constant (Kps). These equilibrium parameters can be determined by sorption experiments. These parameters are calibrated based on the P sorption experiments of @Nguyen2019d for Saigon River Estuary in 2017. PIP concentration data are only available in the experiment of @Nguyen2019, but not in the routine monitoring program of CARE and CEM. Therefore, the average concentration of PIP is declared for the upper and lower boundary. The variations of simulated PIP in estuaries depend only on the adsorption/desorption processes process, not contributed by PIP from tributaries and canals.


```{r C5-C-GEM, echo=FALSE, out.width='100%', fig.cap='Biogeochemical processes in C-GEM applied for Saigon River Estuary.', fig.align='center'}
knitr::include_graphics("figures/C5-C-GEM.pdf")
```

Although the P exchange between the particulate phase (PIP) to the liquid phase ($\mathrm{PO_4^{3-}}$) and vice versa is very slow (from one week to one year) and never achieve equilibrium [@Nemery2005], the P sorption process in C-GEM reaches equilibrium at each time step to reduce model complexity. The potential PIP is calculated based on Langmuir equilibrium isotherm. The $\mathrm{PO_4^{3-}}$ at each time step is calculated in the below equation (based on @Vilmin2015).

$$\frac{PIP}{TSS}=P_{ac}\times\frac{PO4}{PO4+K_{ps}}$$

$$\left[PO4\right]\left(t+dt\right)=\left[PO4\right]\left(t\right)-(\left[\mathrm{PIP}\right]_{\mathrm{potential}}-\left[PO4\right]\left(t\right))\times\frac{dt}{dt+K_{dt}}$$


if $PIP_{potential}>PIP$: adsorption, otherwise desorption


```{r C5-table1-reaction, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Biochemical reactions and formulations of the processes applied in C-GEM for Saigon River") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH5-table1-Biochemical reactions and formulations of the processes.pdf") %>%
  landscape() %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

Although there have been changes in kinetic reaction rates compared to the original version of C-GEM to improve suitability for application at Saigon River Estuary, deficiencies of some other processes may affect simulation results. On one hand, benthic-pelagic processes need to be added to accommodate shallow depths upstream. On the other hand, the addition of zooplankton will improve the accuracy of phytoplankton dynamics. However, these processes are not available in this study because of missing data for calibration. The calibration of phytoplankton mortality rate includes the impact of bacteria and zooplankton grazing. Besides, the omission of benthic processes is accepted based on the longitudinal depth profiles in 2019, 2020. The concentrations of DO, Chl-a, pH are not significantly different between surface water and riverbed, indicating a well-mixed estuary (Figure \@ref(fig:C5-profile)). Only TSS increased strongly in the riverbed, but more than 50% TSS concentration did not change significantly with depth except the last 30 km downstream (Figure \@ref(fig:C5-profile)), a well-known turbidity maxima estuarine process. Besides, the biogeochemical module in C-GEM only presents the Si consumption by diatoms but not for the dissolution of diatom shells. In addition, the assessment of the retention capacity of the estuary by C-GEM currently mainly depends on the biogeochemical processes in the water column and tidal effects because the current model does not include the burial process.


```{r C5-profile, echo=FALSE, out.width='110%', fig.cap='Longitudinal depth profiles along Saigon River Estuary in 2019 - 2020.', fig.align='center'}
knitr::include_graphics("figures/C5-profile.pdf")
```

#### Model performance criteria

Simulation efficiency is evaluated through three parameters $\mathrm{R^2}$, RMSE and percent bias (pbias). The spatial simulation results will be compared with the average concentrations in the dry and rainy seasons along the Saigon River 2017-2018. Temporal simulated results are compared with water quality bi-weekly data of CARE and a monthly tidal range of CEM at three monitoring stations at km 86, 130 and 156.

The formulas for model performance evaluation are presented in Table \@ref(tab:C5-table2-indicator). Phytoplankton (sum of diatom and non-diatom) is converted to Chl-a concentrations with an equivalent conversion factor (1 $\mu g$ chl-a is equal to 40 $\mu g$C). The input data of phytoplankton biomass for C-GEM was also converted from Chl-a concentration.

```{r C5-table2-indicator, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Evaluation criteria for model performance") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH5-table2-Evaluation criteria for model performance.pdf") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

## Results and discussion

### Evaluation of model performance

The model performance evaluation is presented in Figure \@ref(fig:C5-range), \@ref(fig:C5-model-profile), \@ref(fig:C5-model-station), and summarized in Table \@ref(tab:C5-table3-validate). The application of a steady-state version of C-GEM in the dry season at Saigon River Estuary has been highly appreciated for its effectiveness in describing spatial fluctuations, describing well the influence of megacity [@Nguyen2021]. In this study, the C-GEM also performs well the spatial trend of water quality along Saigon River Estuary for both dry and rainy seasons. The simulation results are extended to evaluate the effectiveness of temporal variation of water quality at three typical areas of Saigon River Estuary (Table \@ref(tab:C5-table3-validate), Figure \@ref(fig:C5-model-station)).

The current version of C-GEM generally allows a good description of the tidal range and the salinity intrusion in both seasons with $\mathrm{R^2}$ about 0.6 - 0.9, bias 11 - 16%. The water quality simulation is not as good as hydrodynamics. The model only achieves good results for some variables in some periods and some locations. Specifically, for the efficiency of describing the temporal trends ($\mathrm{R^2}$), DSi, Chl-a, TOC have good simulation results in the urban section with $\mathrm{R^2}$ of 0.4 - 0.8. The simulation of DO concentration trend is only acceptable upstream with $\mathrm{R^2}$=0.5. The simulation of $\mathrm{NO_3^{-}}$ reaches $\mathrm{R^2}$ of 0.4 – 0.5 in the rainy season in the upstream and urban sections. C-GEM does not well describe $\mathrm{NH_4^{+}}$, $\mathrm{PO_4^{3-}}$, TSS with only $\mathrm{R^2}$ of about 0.1 – 0.3. Although the simulated results are not appreciated for some parameters, the percentage error such as $\mathrm{NO_3^{-}}$ is still less than 20% during the rainy season. Likewise, the error in the DO simulation is just around 5% in downstream.


```{r C5-table3-validate, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="The performance of the simulated temporal variation by C-GEM in the Saigon River in dry, rainy seasons") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH5-table3-The performance of the simulated temporal variation.pdf") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

The following two subsections discuss the dynamics of the hydrological regime and water quality of the Saigon River, respectively. The advantages and disadvantages of applying C-GEM model in terms of existing data of Saigon River Estuary are also discussed.

#### Hydrodynamics and transport

The simulated tidal fluctuations in the Saigon River by C-GEM are well evaluated. However, the accuracy in describing hydrodynamics processes gradually decreases from the estuary to the upstream. At km 130, the model efficiency has $\mathrm{R^2}$, RMSE, error of 0.9, 0.2 m, 11%, respectively, while at the upstream area of km 86, the efficiency is reduced to 0.6, 0.33m, 16% (Figure \@ref(fig:C5-range)). C-GEM underestimated the tidal range mainly from May to November, during the rainy season at Saigon River Estuary. A shortening of the tidal range from reality can cause errors in assessing pollutant mobility over tidal cycles. Salinity can spread about 70 km from the sea mouth to the mainland in the rainy season, while the simulation results only describe 50 km of saline intrusion. In the dry season, the resulting salinity intrusion is only about 70 km inland, while the monitoring data shows this number can be up to 100 km. This error originates from the great influence of Dongnai River, at km 130 (70 km from the sea mouth) with a freshwater flow of about 300$m^{3}$ $s^{-1}$ (more than ten times that of Saigon River) which greatly increases the flushing capacity from km 130 to estuary mouth. In addition, the simulated depth and cross-sectional area are used to calculate the volume of three sections along the Saigon River to estimate the retention time in these areas. Residence time varies from about 20 days in the rainy season to 120 days in the dry season (Figure \@ref(fig:C5-range)).

```{r C5-range, echo=FALSE, out.width='100%', fig.cap='The observed and simulated tidal range, salinity in the Saigon River from 2017 to 2018. Three dotted lines in the left panels are locations of three stations at km 86, 130 and 156. The grey line indicates Dongnai River. The tidal range observation data are calculated based on hourly water level data from the Vietnamese Center of Water Management and Climate Change (WACC). The residence time is calculated by dividing the simulated volume by the flow of each section.', fig.align='center'}
knitr::include_graphics("figures/C5-range.pdf")
```

Applying C-GEM with hourly data at estuary mouth and monthly freshwater inflow has enabled a good description of the trend and intensity of tidal oscillations at Saigon River Estuary. Compared with high-resolution flow data (10 minutes) from @Camenen2021, monthly monitoring data does not capture the high variations of daily discharge in the Saigon River. The difference can be explained by Dau Tieng Reservoir's activities upstream, markedly controlling the freshwater discharge in the Saigon River [@Camenen2021]. However, the flooding of the Dau Tieng Reservoir mostly occurs in the rainy season with a highly flushing capacity. The use of hourly tidal variation from global gauge data in C-GEM still ensures the assessment of tidal intensity changes to the propagation and diffusion of substances along the estuary. It should be noted that although the model was able to describe contaminant propagation, the lack of daily discharge data at the upstream boundary and tributaries could cause large concentration differences in the dilution equation. @Arndt2009 suggested that the transient behaviours of variables in estuaries are often derived from fluctuating riverine boundary conditions. The failure to describe the peaks in the concentrations of the observed substances is inevitable if using the data of average conditions. Therefore, daily discharges for the upstream and all tributaries are recommended to describe the short-term variations at the estuaries.

#### Nutrients and phytoplankton dynamics

The transient version of C-GEM describes well the water quality trends along the Saigon River under the impact of urban wastewater of HCMC for both dry and rainy seasons (Figure \@ref(fig:C5-model-profile)). Similar to observed data, $\mathrm{NH_4^{+}}$, TOC, Chl-a, DO have significant differences between the three sections, upstream (km 0 – 86), urban (km 86 – 130) and downstream (km 130 – 200) of Saigon River estuary. C-GEM has well described $\mathrm{NH_4^{+}}$, $\mathrm{NO_3^{-}}$, DSi, Chl-a in both seasons with $\mathrm{R^2}$ from 0.7 – 0.9. Note that spatial variations of Chl-a, DSi, $\mathrm{NO_3^{-}}$ are only compared with the observed data of three monitoring stations, while the remaining parameters are compared with 6-7 stations. Although only a few stations represent three areas along the Saigon River, the trend of water quality change under the impact of urban discharge has been clarified.

In the dry season, the observed concentrations of TOC and $\mathrm{NH_4^{+}}$ both increased, leading to a decrease in DO to anoxic conditions (\< 1 mg $\mathrm{O_{2}}$$L^{-1}$). However, the simulation results of C-GEM for DO concentrations are from 2 to 5 mg$L^{-1}$, and $\mathrm{R^2}$ only reaches 0.35, the lowest $\mathrm{R^2}$ compared to the remaining variables. The overestimated DO stems from a rapid increase in phytoplankton density from near zero to more than 60 $\mu g$ Chl-a$L^{-1}$ in the urban section of Saigon River. The phytoplankton bloom increases the DO concentration in the water by photosynthesis, roughly equal to the oxygen consumption by nitrification and organic carbon respiration in the simulation. The simulated phytoplankton in the dry season is almost always in high density because the simulated nutrient source (N, P) is not exhausted. The high nutrients availability allowed the growth of phytoplankton not to be limited by the required nutrients, especially $\mathrm{PO_4^{3-}}$. Based on the C:N:P:Si ratio analysis by @Nguyen2019, $\mathrm{PO_4^{3-}}$ is the limiting factor for phytoplankton development in the Saigon River Estuary. Indeed, there were several times the the $\mathrm{PO_4^{3-}}$ was lower than 0.01 mgP$L^{-1}$ during the dry season. Phytoplankton biomass in the urban section in the dry season can reach more than 130 $\mu g$ Chl-a$L^{-1}$ in the dry season, while many periods are still below the detection threshold. This shows that the intensity of biogeochemical processes and phytoplankton bloom is very strong in the Saigon River. Strong fluctuations of phytoplankton over a short period (days to weeks) are also detected according to studies conducted in other estuaries such as Scheldt Estuary in 2006. Phytoplankton biomass at this estuary reached 150 $\mu g$ Chl-a $L^{-1}$ in summer and can decrease to almost undetected levels in winter [@Gypens2013]. The study at Scheldt Estuary detected phytoplankton depletion under low temperature and DSi depletion. in the Saigon River Estuary, solar radiation is almost unchanged throughout the year and is not limited in DSi concentration. Meanwhile, the C-GEM model applied at Saigon River Estuary has not yet detected these extreme values neither in the observations. The two main causes of overestimation are the lack of other important biogeochemical processes (e.g., benthic processes, zooplankton grazing, burial processes), and the use of average monthly data as input values. Based on simulation results at a subtropical estuary of Vietnam, Day River, which also receives highly untreated domestic wastewater, the burial process played an important role in reducing nutrient concentrations in the water column [@Trinh2012]. Specifically, domestic wastewater with high P concentration at Day River is discharged into this river, but up to 38% of phosphorus is retained in bed sediment [@Trinh2012]. Therefore, beside the consumption of $\mathrm{PO_4^{3-}}$ by phytoplankton uptake and P adsorption, the P burial process is important. Downstream, the sudden decrease in concentration from km 130 is due to the dilution of the Dongnai River. A gradual decrease in concentrations is observed at the estuary due to dilution with seawater (e.g., Chl-a, $\mathrm{NH_4^{+}}$, $\mathrm{NO_3^{-}}$, DSi). The improvement of the downstream oxygen concentration is also contributed by the aeration of the water column.

```{r C5-model-profile, echo=FALSE, out.width='100%', fig.cap='Comparison of simulated and observed water quality along Saigon River. The model performance is calculated based on average values in the dry and rainy season 2017 – 2018.', fig.align='center'}
knitr::include_graphics("figures/C5-model-profile.pdf")
```

The $\mathrm{R^2}$ of spatial simulations is mostly from 0.7 to 0.9, except for TOC with $\mathrm{R^2}$ of 0.4. In rainy season, concentrations of TOC, $\mathrm{NO_3^{-}}$, Chl-a, were not significantly different between the upstream, downstream and urban sections of the Saigon River (Figure \@ref(fig:C5-model-profile)). The concentrations of $\mathrm{NH_4^{+}}$ and DO change significantly when the Saigon River received the domestic wastewater of Ho Chi Minh City. In contrast to ineffective simulated DO in the dry season, the simulation efficiency in the rainy season reached $\mathrm{R^2}$= 0.98 with an error of 7%. The $\mathrm{NH_4^{+}}$ simulation results also have an $\mathrm{R^2}$ of about 0.80 but have an error of 105%. This large error is detected in km 116, where the observed $\mathrm{NH_4^{+}}$ concentration was exceptionally high. The high $\mathrm{NH_4^{+}}$ concentration cloud at approximately 30 km long in the dry season is similar to the dry season. However, as a result of saline intrusion in the rainy season, the dispersion is not as strong as in the dry season. This explains the ineffective spreading of high $\mathrm{NH_4^{+}}$ from km 116 to km 130 , as observed. The underestimation is unlikely from an error in the pollutant diffusivity simulation but perhaps from other biogeochemical processes or another external $\mathrm{NH_4^{+}}$ source which is still unanswered.

$\mathrm{NO_3^{-}}$, DSi, $\mathrm{PO_4^{3-}}$, TSS do not have the spatial trends as Chl-a, $\mathrm{NH_4^{+}}$ when receiving wastewater from urban canals. The C-GEM model has well described this anomaly through the role of processes such as denitrification ($\mathrm{NO_3^{-}}$  $N_2O$, $\mathrm{N_2}$) and $\mathrm{PO_4^{3-}}$ adsorption. The concentration of $\mathrm{PO_4^{3-}}$ in the urban section is not significantly different from other sections. The addition of P sorption processes avoids the overestimated $\mathrm{PO_4^{3-}}$ as compared with the results of @Nguyen2021. According to @Minaudo2018, $\mathrm{PO_4^{3-}}$ adsorption/desorption is an important process in understanding the dynamics of phytoplankton and nutrients. While $\mathrm{PO_4^{3-}}$ adsorption in urban section plays a role in reducing $\mathrm{PO_4^{3-}}$ concentration, the downstream $\mathrm{PO_4^{3-}}$ simulation results are not good. Downstream, there is an increase in $\mathrm{PO_4^{3-}}$ mainly from P desorption due to an increase in TSS deposition. However, this process may not be suitable because observed TSS downstream tends to increase gradually. The sediment erosion and increase of $\mathrm{PO_4^{3-}}$ adsorption is thus reasonable. According to the experiment of @Nguyen2019, at high salinity, P desorption will greatly increase in the Saigon River. This coincides with the desorption phenomenon simulated downstream. However, $\mathrm{PO_4^{3-}}$ downstream always has a lower concentration than $\mathrm{PO_4^{3-}}$ in the urban section in @Nguyen2020 and in this study. Dongnai River with low $\mathrm{PO_4^{3-}}$ concentration strongly impacts $\mathrm{PO_4^{3-}}$ dilution, which can lead to a gradual decrease in $\mathrm{PO_4^{3-}}$ concentration.

### Spatio-temporal evolution of nutrients and phytoplankton biomass

Temporal variation of water quality was analyzed in three sections representing upstream, urban and downstream of Saigon River Estuary. These sections correspond to the three monitoring stations of CARE at km 86 (freshwater, less affected by urban wastewater and tides), km 130 (receiving urban wastewater), km 156 (after the confluence with Dongnai River). Figure \@ref(fig:C5-model-station) depicts spatio-temporal variations of TSS, DO, nutrients ($\mathrm{NH_4^{+}}$, $\mathrm{NO_3^{-}}$, $\mathrm{PO_4^{3-}}$, DSi), organic carbon and phytoplankton biomass (Chl-a) over two years.

Simulated $\mathrm{NH_4^{+}}$, $\mathrm{NO_3^{-}}$, Chl-a, and $\mathrm{PO_4^{3-}}$ have seasonal trends with higher concentrations in the dry season. The most distinct difference of water quality parameters between dry and rainy seasons is phytoplankton biomass, but mainly in the urban section of Saigon River Estuary. In addition, simulation and observation data show that the intensity and timing of phytoplankton peaks are also different between the dry seasons of 2017 and 2018. C-GEM well capture the phytoplankton mass development in two periods in 2017-2018, specifically from late November to late May in dry season. The peak of simulated phytoplankton concentration is only 40 $\mu g$ Chl-a$L^{-1}$ at km 130, approximately two times lower than the maximum value observed in 2017. As discussed in hydrodynamics evaluation, the use of average inflow may interfere with the short-term variations of phytoplankton. However, the model still provides a good description of the seasonal trends of phytoplankton. The phytoplankton mortality rate can explain the phytoplankton underestimation. This process has implicitly considered the process of phytoplankton reduction by bacteria and zooplankton grazing. Statistical results from tropical estuaries in India, show that zooplankton is the main factor controlling phytoplankton density in low flow conditions in dry season [@Bharathi2019]. Fluctuations in the inflow of freshwater may cause fluctuations in zooplankton and phytoplankton in the Saigon River.

```{r C5-model-station, echo=FALSE, out.width='100%', fig.cap='The daily evolution of suspended sediment, dissolved oxygen, nutrients, organic carbon and phytoplankton biomass in three stations along Saigon River Estuary from 2017 to 2018. Blue points are observed data of bi-weekly monitoring.', fig.align='center'}
knitr::include_graphics("figures/C5-model-station.pdf")
```


Monitoring data and simulation both show that there is no seasonal trend of TSS, DO, $\mathrm{PO_4^{3-}}$. The simulated transient variations of TSS have the same signal as the tidal fluctuations. This result is similar to simulation results of higher resolution models at Vietnamese Mekong Delta [@Marchesiello2019; @LeTu2019] or in other temperate estuaries such as Scheldt Estuary [@Arndt2009;@Arndt2011]. Besides, tidal elevation has no seasonal trend but only changes according to the semi-diurnal regime (i.e., two high tides and two low tides per day). Therefore, TSS also has a similar tendency and creates transient variations of $\mathrm{PO_4^{3-}}$ adsorption onto TSS. Using the simple sediment erosion/deposition module allowed a good description of the fluctuation range and concentration of TSS in the Saigon River. However, the model is not effective in describing the changing trend of TSS in the urban area (km 130). Indeed, @Vanderborght2007 argued that extending the reaction-transport model (e.g., C-GEM) to describe sediment dynamics accurately will increase the model complexity and is often out of reach of a simplified model to minimize the input data. However, in this study, the established P adsorption/desorption process reached equilibrium at each time step. The small fluctuation of TSS thus led to a significant change in the concentration of simulated $\mathrm{PO_4^{3-}}$. As a result, the simulated $\mathrm{PO_4^{3-}}$ concentration downstream increased three times from 0.1 to 0.3 mgP$L^{-1}$ during sedimentation (Figure \@ref(fig:C5-model-station)). Simulated $\mathrm{PO_4^{3-}}$ in the urban section regularly increases due to P desorption from sediment, especially in the rainy season and the dry season (from July to January next year). In the upstream, the P sorption process has almost no significant influence because of the small concentration of TSS in this area.

While the C-GEM model accurately depicts pollution levels in the urban section and downstream, the C-GEM simulation results underestimated the $\mathrm{NH_4^{+}}$, TOC concentration upstream. Low simulated TOC concentrations are associated with low phytoplankton densities upstream and lead to lower phytoplankton mortality. According to the observation, concentrations of TOC and Chl-a have a linear correlation in the upstream and urban sections, suggesting that phytoplankton contribute significantly to organic carbon concentration in these areas [@Nguyen2021b]. This statement is described by C-GEM at the phytoplankton peak events in April 2017 and April 2018 at the upstream station (Figure \@ref(fig:C5-model-station)). The concentrations of TOC at these two periods both increased markedly. However, the simulated TOC concentrations were lower than observed in the remaining periods because of the lack of external organic carbon sources. @Nguyen2020b determined that the upstream area of Saigon River Estuary increases organic carbon in groundwater due to the strong agricultural activities. The exchange between freshwater and groundwater upstream of Saigon River Estuary has not yet been implemented in the C-GEM model. Likewise, simulated $\mathrm{NH_4^{+}}$ concentrations were underestimated. However, this result does not affect the low nutrient status assessment in the Saigon River Estuary's upstream.

### Seasonal evolution of primary production and nutrients cycling

Figure \@ref(fig:C5-process) depicts the evolution of the major biogeochemical processes involved in phytoplankton abundance and nutrients cycling. In addition to the description of biogeochemical processes, external forcings are described, including nutrients limitation, light utilization efficiency, oxygen air exchange. The nutrients limitation is calculated based on $\mathrm{NH_4^{+}}$, $\mathrm{NO_3^{-}}$, $\mathrm{PO_4^{3-}}$ and DSi in the case of diatom simulation. The results in this study present nutrients limitations applied for all four parameters. Light utilization efficiency is calculated based on the photosynthetic rate of phytoplankton, light extinction coefficient (depending on background light attenuation and TSS) and water depth. Oxygen air exchange is calculated based on flow velocity, wind speed, water depth and current oxygen concentration.

```{r C5-process, echo=FALSE, out.width='110%', fig.cap='Spatio-temporal evolution of nutrient limitation and the main biogeochemical processes in dry and rainy seasons 2017-2018 along Saigon River Estuary.', fig.align='center'}
knitr::include_graphics("figures/C5-process.pdf")
```
The description of the evolution of nutrient limitation provides important information in describing the nutrient availability for phytoplankton growth and disappearance, especially for transient variations in estuaries and coastal areas [@Arndt2011]. To assess the risk of eutrophication, the nutrient limitation for phytoplankton growth is the keystone [@Costa2009;@Smith2006]. Simulation results of C-GEM show that the nutrient limitation only exists about 30 km to 75 km during the rainy season. In the remaining periods, the availability of nutrients for phytoplankton growth ranges from 40% to over 80%. Especially in the urban section, from km 85 to 150, the nutrient availability for the primary production process is always high, reaching nearly 100% regularly in both the dry and rainy seasons. in the Saigon River, the simulation results show that high nutrient concentrations are readily available for phytoplankton bloom in the urban section of Saigon River regardless of the season. Simulation of Net Primary Production (NPP), i.e., biomass growth of phytoplankton after primary producers' respiration has been removed, is similar to nutrient availability. This result is similar to that observed at other tropical estuaries, e.g., Sul River Estuary (Brazil), where it was consistently found that the nutrient limitation did not vary significantly between seasons despite low or high flushing conditions [@Costa2009]. However, phytoplankton density in the rainy season is much lower than in the dry season [@Costa2009]. in the Saigon River, according to observations and simulations, phytoplankton bloom also only occurs in the dry season (Figure \@ref(fig:C5-model-profile)). This result contrasts the survey in the Red River Delta (subtropical estuary, Northern Vietnam). The density of phytoplankton and zooplankton in the dry season is equal to or lower than in the rainy season [@Hoang2018]. @Hoang2018 observed that the nutrient availability is higher in the rainy season than in the dry season because of the run-off. Besides, temporal variability of phytoplankton biomass in temperate estuaries directly links to seasonal variations in light availability [@Lancelot2011;@Domingues2005]. The light limitation is thus frequently referred to as controlling phytoplankton growth. The light utilization efficiency calculation in the urban section is the lowest compared to other sections regardless of the season. The depth of this area (about 20m) is the main reason for phytoplankton's inefficient use of light. This finding suggests that although both upstream and downstream have higher light utilization efficiency than urban sections thanks to the shallower of the estuary (about 10m), the NPP in these two areas is still many times lower than urban sections because of nutrient limitation.


The right panels of Figure \@ref(fig:C5-process) depict the process involving oxygen concentration in the water, including nitrification, denitrification, organic carbon aerobic degradation and oxygen air exchange. C-GEM also describes strong reaction rates of these processes for the urban section of Saigon River. These processes help to explain the decrease in DO concentration under the influence of high nutrient concentrations from urban canals. Oxygen air exchange is the physical process that is most markedly different from the biogeochemical processes described at Saigon River Estuary. The transfer of oxygen from the atmosphere into the water column is particularly high in the Saigon River. From km 50 to 90, there is no significant contribution of this process compared to other areas. This is mainly due to the decrease in current velocity while the water level of this area is still relatively deep compared to about 5m in the upstream area.

### Driving factors of phytoplankton evolution

According to a synthesis by @Lancelot2011, under adequate nutrient conditions in estuaries, salinity gradient, light availability, and water residence time are factors controlling phytoplankton dynamics. This statement is consistent with urbanized tropical estuaries suffering the domestic wastewater of megacities such as in the Saigon River but existing the short-term variations of phytoplankton (days to weeks) and distinct phytoplankton density between dry and rainy seasons (Figure \@ref(fig:C5-model-profile)).

```{r C5-HP, echo=FALSE, out.width='100%', fig.cap='Schematic representation of the contribution of driving factors to phytoplankton bloom in the Saigon River Estuary in the dry and rainy season. The top panels are calculated based on the normalized factors in upstream, urban and downstream in 12 months. The independent effects on phytoplankton in lower panels are calculated based on the hierarchical partitioning method, calculated for each section.', fig.align='center'}
knitr::include_graphics("figures/C5-HP.pdf")
```

The implementation of model sensitivity analysis allows understanding the mechanisms controlling the dynamics of simulated phytoplankton in the estuary [@Volta2014; @Gypens2013]. This approach can evaluate the interaction of biogeochemical processes and hydrological conditions on phytoplankton growth. @Gypens2013 proposed to assess the tidal effects on phytoplankton by setting all biological activities to zero. However, in the Saigon River Estuary, phytoplankton density downstream was very low, and there was not much impact of marine phytoplankton on the change of phytoplankton density in the urban section [@Nguyen2021b]. In addition, based on the nutrient limitation simulation results, the origin of phytoplankton bloom in the urban section of Saigon River is due to abundant nutrients in this area, while the upstream and downstream areas have relatively low nutrient concentrations. Therefore, we propose determining the independent effects of potential factors (e.g., biogeochemical processes, hydrodynamics, weather conditions) on phytoplankton biomass through the hierarchical partitioning (HP) method. The HP method effectively separated the independent effects of each environmental parameter which could not be done by regression methods [@Nally1996]. Different from the purely statistical studies in determining the driving factors to phytoplankton in tropical estuaries (e.g., @Nguyen2021b and @Hoang2018), this study uses explicit factors such as nutrient limitation, light availability, residence time, salinity from simulation to calculate their independent effects on phytoplankton development.

Figure \@ref(fig:C5-HP) depicts the independent effects of four representative parameters of nutrient availability, hydrological, climatic and environmental conditions affecting phytoplankton biomass. In addition, all these parameters are normalized to compare the cooperative effects of these factors. Note that the effects of these factors on phytoplankton do not follow a linear trend but are as complex as formulas presented in Table \@ref(tab:C5-table1-reaction). This approach aims to explain the significant differences in phytoplankton between the rainy and dry seasons. The sum of normalized factors in the dry season from December to May of the following year was highest, which is reasonable for the observed and simulated phytoplankton in this period. Specifically, nutrient limitation and residence time have the greatest impact on phytoplankton. The nutrient limitation does not significantly differ between the two seasons, while residence time had a marked difference. At upstream and downstream, salinity and light utilization factors are dominant and differ to urban section. This is consistent with the shallow water with more light diffusion upstream. The upstream salinity contribution probably describes the sensitivity of freshwater phytoplankton to saltwater. However, this area is rarely subject to a high rate of saline intrusion.

In the urban section, the results of hierarchical partitioning analysis also show that residence time is the key factor that has the greatest impact on the fluctuation of phytoplankton. This result is similar to the output of a simple box model applied in the Gulf of Thailand, where widespread strong hypoxia and phytoplankton bloom over half of the gulf was observed during the dry season with a residence time of about 40 days. In the rainy season, the residence time was only about three days and strongly flushed pollutants into the sea and improved oxygen levels [@Buranapratheprat2021]. in the Saigon River Estuary, the residence time is about 20 days in the rainy season and 120 days in the dry season (Figure \@ref(fig:C5-range)), creating higher phytoplankton densities. This result contradicts the observation in a subtropical estuary, Red River Delta (Vietnam), as discussed previously [@Hoang2018]. This estuary also receives high nutrient concentrations from domestic wastewater of about 8 million inhabitants (only 20% are treated) in Hanoi City. The nutrient limitation is the cause of this difference. While nutrient availability tends to remain stable during the entire year in the Saigon River [@Nguyen2019], it increases in the Red River [@Hoang2018]. Therefore, the residence time has become a key factor for the development of phytoplankton at Saigon River Estuary.

## Conclusion

The application of C-GEM in characterizing changes in nutrient and phytoplankton concentrations has allowed understanding the mechanisms leading to fluctuations in water quality at Saigon River Estuary. Compared with the steady-state C-GEM version, which only described the spatial variation in the dry season, this study reproduces the distinct phytoplankton biomass between the two seasons, which is questionable for many tropical estuaries. The addition of the $\mathrm{PO_4^{3-}}$ sorption module has reduced the overestimation of $\mathrm{PO_4^{3-}}$ simulation, but the complexity of the sediment dynamics in downstream causes several unexpected behaviours. The model extensively described biogeochemical processes to determine their role in responding to the megacity's domestic wastewater. Nitrification, denitrification, organic carbon degradation, and primary production occurred strongly in the urban section of Saigon River Estuary. Simulation results for two years showed that nutrient availability in the urban section was abundant for phytoplankton growth in both seasons. By applying multivariate data analysis for simulated results, the significant contribution of residence time to the difference in phytoplankton density is exposed. This study confirms the potential of C-GEM in using a minimal amount of input data while providing essential information for eutrophication management. Applying this model in tropical estuaries in emerging countries will be particularly useful in supporting the assessment of the effectiveness of reducing nutrient pollution through the construction of wastewater treatment plants such as in Ho Chi Minh City.

## Conclusion of Chapter 5

Questions about biogeochemical functioning and hydrological impact on nutrients and phytoplankton dynamics at an urbanized tropical estuary are answered in Chapters 4 and 5. The current version of C-GEM application in the Saigon River is now suitable for assessing the several scenarios of HCMC in the future. The next chapter will present the results of water quality simulation according to the development of this megacity. The risk of future eutrophication is thus discussed to recommend additional solutions to meet the region's sustainable development.
