---
output:
  #bookdown::html_document2: default
  #bookdown::word_document2: default
  bookdown::pdf_document2: 
    template: templates/brief_template.tex
    citation_package: biblatex
documentclass: book
bibliography: [bibliography/references.bib]
---
  
# Eutrophicaiton assessment in an urbanized estuary based on past, present and future nutrient loadings: A modeling approach
\chaptermark{Eutrophication assessment}
\minitoc <!-- this will include a mini table of contents-->

```{r , include=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
```

\noindent In the context of densification of urban centers and climate change, coastal megacities in emerging countries have to face new issues about water management. With more than 9 million inhabitants, Ho Chi Minh City (HCMC), the economic capital of Vietnam, is in full demographic and economic expansion. The city does not have water treatment networks that meet urgent needs. Indeed, in 2020, less than 15% of the population is connected to a sewerage network and treatment plant. The deterioration of the water quality of the Saigon River crossing the city is dramatic. Therefore, Ho Chi Minh City authorities plan to build new sanitation networks and ten new wastewater treatment plants (WWTPs) over the next fifteen years, drawing inspiration from industrialized countries' centralized water management models. However, the relevance of this extremely expensive choice arises and is not necessarily adapted to the particular conditions of this megacity in tropical monsoon zone and under the influence of tidal cycles. Assessing the assimilative pollutant capacity of the Saigon River and the requirement of wastewater treatment capacity is extremely complex at tidal estuaries. The significant difference in eutrophication risk between the dry and rainy seasons was observed in the Saigon River Estuary under the variation condition of hydrology, climate, and urban discharge (Chapter 4, 6). Using the estuarine biogeochemical model (C-GEM), this study aims at assessing the water quality and eutrophication of the Saigon River in response to four scenarios of HCMC development from the past to the future (the 2000s to 2050). The scenarios are implemented based on population growth, new water management infrastructures in HCMC and projected climatic conditions (temperature, hydrology) in the next decades. The results of four scenarios show that the construction of new WWTPs is supported. However, the rapid population growth in HCMC may overcome the WWTPs' capacity, which returns the water quality in the Saigon River to bad condition.

Chapter 6 uses the information presented in the conference proceedings:

Nguyen, Truong An, et al. "Evaluating the response of water quality to pollutant loading in the Saigon River system (Vietnam): modelling scenarios by C-GEM, an estuarine biogeochemical model." Water, Megacities and Global Change. 2020.

## Introduction

### Risk of eutrophication in coastal megacities

Accounting for more than 10% of the world's population, megacities have been attracting people and increasing population densities themselves. Around 70% of the megacities are distributed along estuaries and coastal areas known to provide numerous ecosystem services for the development of the surrounding region [@Glasow2013; @GRIDArendalandUNEP2016] . Estuaries are also considered a high self-purification system, with the ability of transformation or pollutant removal [@Wilk2018]. However, the rapid population growth coupled with the lack of investment in water environmental protection in the megacities of the developing countries threatens the safety of water resources [@Choudhury2010]. In the 21st century, water pollution from megacities is well known in Southeast Asia countries [ @TranNgoc2016; @Choudhury2010]. According to Hoonrweg & Pope, 2014, the rate of population growth and the formation of megacities by 2050 will be dominated by countries in Asia in tropical regions. Comparing the nutrients export between 2000 and 2050 projection shows that the total amount of nitrogen exported by rivers to the sea in South Asia will be many times higher than in other regions such as Europe, North America, Africa, Australia [@Lee2016]. The rapid development of agriculture, industry, and urbanization of emerging countries will significantly increase the nutrients loading into the estuaries and, ultimately, the ocean. This indicates that the pressures of population growth and urbanization for tropical estuaries in emerging countries will intensify in the future. Facing the threat of water safety, these megacities have been implementing environmental rehabilitation projects, particularly the construction of municipal wastewater treatment plants (WWTPs) [@Sajor2009; @TranNgoc2016].

### Ho Chi Minh Megacity, Vietnam

Located in Vietnam's most rapid urbanization area, Saigon River Estuary (Southern Vietnam) has been affected by urban wastewater from Ho Chi Minh City (HCMC) for many years. Canals in HCMC and the Saigon River received about 90% of untreated domestic wastewater in 2016 [@Nguyen2020]. The high polluted loads regularly lead to an excess of nutrients and an intense period of anoxia in the Saigon River, especially during the dry season in the Saigon River [@Nguyen2019]. Water pollution also threatens water production for the region. Over 90% of the water supply used in HCMC is taken from two raw water collection stations on the Saigon River and Dongnai River [@TranNgoc2016]. Under the impact of tides from the East Sea of Vietnam, some water production stations have to shut down several times because of saline intrusion and polluted water from the inner city of HCMC in the Saigon River (Le *et al.*, 2012). The risk of estuarine pollution in the Saigon River is expected to increase in the coming years, under the impact of megacity's development. Following the actual trend, the population of HCMC is expected to reach 23 million by 2050, and the Saigon River will then receive three times more pollution [@Nguyen2020]. In response to the current and future major sources of urban wastewater, the HCMC is implementing an ambitious environmental sanitation project for the urban canals and Saigon River, i.e., more than ten WWTPs to treat about 80% of domestic wastewater in HCMC by 2025. However, because of the huge cost of the project (about 2 billion USD), most of WWTPs are still in stage of calling for capital investments. In this context of rapid changes in pollutant loadings to the Saigon River Estuary (i.e., new WWTPs, increased population), the use of models capable of describing the biogeochemical functioning of aquatic ecosystems and simulating the potential impact of improved urban water treatment is crucial.

### Tools for eutrophication management

Managing estuaries involves managing human activities, which involve complex socio-economic-political issues [@Atkins2011]. As a result, the management of estuaries and coasts has a long track record of a few successes and many failures. According to @Wolanski2014, the essence the anthropogenic problems come from three sources: (1) materials that are put into the estuaries, (2) materials taken out from the estuaries, and (3) wider problems such as climate change. The failure to control "materials that are put into the estuaries" clearly lacks a balance between wastewater volume and wastewater treatment capacity. In addition, there is the effect of a lack of control of products with a high risk of eutrophication. The failure to quantify the "materials taken out from the estuaries" is due to the lack of intensive monitoring and biogeochemical model application that allows the calculation of retention capacity and transformation of nutrients.

In tropical estuaries, phytoplankton blooms and eutrophication risk often experience short-term variation (hours to days) and long-term variation (months to seasons) depending on hydrology, climate, and the duration of anthropogenic nutrient loading [@Livingston2000]. To understand the level of eutrophication and the water quality of urbanized estuaries, it is necessary to consider the aspect of nutrient loadings (i.e., control of waste sources) and consider other aspects such as season, climate and tidal effects. Chapter 5 exposed the importance of nutrient limitation and residence time of water body to phytoplankton growth in a tropical estuary. The accurate assessment of several nutrient loadings scenarios on water quality is thus required an estuarine biogeochemical model to describe transient variations, seasonal effects of estuaries.

Applying biogeochemical models in developing scenarios in urbanized tropical estuaries is not as widespread as in developed countries. Some lessons on applying the model in assessing pollution scenarios in urbanized tropical estuaries are as prominent as in Bay of Santa Catarina in Brazil [@Cabral2019], Shenzhen Bay in China [@Zhou2020]. Shenzhen Bay is an urbanized estuary heavily influenced by anthropogenic activities, which have influenced eutrophication since the 1990s. Therefore, the government has developed and applied several scenarios to alleviate water degradation and improve coastal ecosystems. The survey results and nutrient limitation simulation show that eutrophication in Shenzhen Bay was limited by phosphorus; the government thus decided to improve the effectiveness of WWTP to reduce phosphorus first to alleviate the serious aquatic situation immediately. The results show that since 2005 the trophic status at this bay has improved, and then the several practices to reduce the nitrogen concentration and keep Shenzhen Bay water quality in good condition as currently [@Zhou2020]. At the Bay of Santa Catarina, based on over 30 years of monitoring results and a simple model of eutrophication evaluation, this urbanized estuary is susceptible to eutrophication due to domestic wastewater. Modeling results show that trophic status ranges from moderate to eutrophic in this bay. The main cause of the eutrophication of this area was the low rate of domestic wastewater treatment (27%) and the treatment technology of WWTPs, which only removes 20-50% of nitrogen while nitrogen is the limiting factor for eutrophication in this system. The results of this study suggest that improving WWTPs technology and expanding capacity will be an effective solution in eutrophication management [@Cabral2020]. Overall, these two examples show that upgrading WWTPs is inevitable to ensure the water security of urbanized estuaries. In addition, identification of the controlling factors that affect eutrophication through estuarine modeling is necessary.

The Carbon Generic Estuarine Model (C-GEM) is a generic one-dimensional, reaction-transport model which takes advantage of the relationship between estuarine geometry and hydrodynamics to minimize data requirements (Volta *et al.*, 2014). C-GEM has provided accurate descriptions of estuarine hydrodynamics and biogeochemical transformations in several estuaries, especially in temperate regions [@Volta2014; @Volta2016b; @Laruelle2017; @Laruelle2009] . The model allows the assessment of the estuary response to the simultaneous effects of point sources (e.g., domestic, industrial wastewater) and the effects of the estuary tidal regime [@Volta2014]. This modeling approach is particularly relevant to the urban estuary of developing countries where there is a limited extensive monitoring program. This study aims to evaluate the water quality in the Saigon River Estuary under the impact of urban wastewater from HCMC's development by C-GEM application. The research results allow us to evaluate the efficiency of the construction of WWTPs according to the direction of HCMC's development.

## Materials and methods

### Study area

The Saigon River catchment with 4717 $km^{2}$ is located in Southern Vietnam. The whole Saigon River has a length of 280 km within Vietnam. The river originates in Cambodia and is firstly controlled at Dau Tieng Reservoir (270 $km^{2}$ and 1580x106 $m^{3}$), designed for ﬂood and saline intrusion control, irrigation, domestic, agricultural and industrial demands (Ngoc, Hiramatsu and Harada, 2014). From Dau Tieng Reservoir to the estuarine mouth (200 km), the Saigon River (net discharge 18 $\pm$ 14 $m^{3}$ $s^{-1}$) in turn joins notable tributaries such as the Thi Tinh River (TT River, 20 $\pm$ 11 $m^{3}$ $s^{-1}$) and the Dongnai River (632 $\pm$ 446 $m^{3}$ $s^{-1}$), forming Nha Be River, it then splits into two distributaries (Soai Rap River and Long Tau River) flowing into East Sea of Vietnam (Figure \@ref(fig:C5-sampling)). In addition, the Saigon River is also connected to an urban river (Vam Thuat River, 4 $m^{3}$ $s^{-1}$) and three urban canals (total net discharge, 5.5 $m^{3}$ $s^{-1}$) of HCMC before the confluence with Dongnai River [@Nguyen2020]. The typical climate of this region is tropical monsoon which has two distinct seasons (dry and rainy seasons), with a relatively constant temperature (about 28$^{\circ}$C). The rainy season lasts from June to November, with an average annual rainfall of 1800 mm, of which 80% fall in the rainy season (Nguyen *et al.*, 2019). The status of water quality in the Saigon Estuary River is spatially different. Upstream and downstream of the estuary, water quality status is considered good. The quality status of the water became moderate to bad right after connecting with the Vam Thuat River (an urban river) and urban canals of HCMC [@Nguyen2019].


```{r C5-sampling, echo=FALSE, out.width='100%', fig.cap='Map of Saigon River Estuary, monitoring stations and HCMC population density in 2019.', fig.align='center'}
knitr::include_graphics("figures/C2-CEMsite.pdf")
```

Ho Chi Minh City's development is intrinsically linked with the Saigon River. This megacity's economy has developed considerably over the last decades, and HCMC was the third most dynamic city in the world in 2020, followed by Hyderabad and Bengaluru, India (City Momentum Index, 2020). In 2015, the Department of Natural Resources and Environment reported that land use in HCMC is dominated by agricultural activities in the north (41% of HCMC's area), an urban settlement in the center (23%) and mangrove forest to the south (26%). The population growth rate in Ho Chi Minh City is about 3.48% per year (Ho Chi Minh City Statistical Yearbook, 2017), leading to the city has the highest population density in Vietnam. However, the development of water treatment systems is not in line with that population growth. In the early 2000s, HCMC discharged about 532000 $m^3 d^{-1}$ domestic wastewater of about 6 million inhabitants (80−130 liters per capita per day) and 51300 $m^3 d^{-1}$ industrial wastewater of 18 industrial parks mainly untreated (Le *et al.*, 2012).

Consequently, most water quality variables exceeded Vietnamese surface water quality standards (QCVN 08:2015/BTNMT). In addition, the influence of tides causes saline intrusion and pollutant transport from an urban area to the upstream area. This caused a temporary shutdown of drinking water treatment plants in the dry season in 2004 and 2005  (Le *et al.*, 2012).

### Urban canals and WWTPs

Figure \@ref(fig:C5-sampling) depicts the canal systems in HCMC with the main canals of Tham Luong Canal- Vam Thuat River, Nhieu Loc - Thi Nghe canal, Ben Nghe canal. These canals systems were built in the late 19th century during the French colonial period. Until now, domestic water and rainwater have been collected and transported together by the sewer system and by urban canals and creeks [@TranNgoc2016].

Before 2006, all domestic wastewater was discharged directly into canals and the Saigon River. The current WWTPs system in Ho Chi Minh City still does not meet the total domestic wastewater volume of about nine million people (about 2.5 million$m^{3}$  $d^{-1}$, in 2020). Therefore, most domestic wastewater is discharged into the canal system of HCMC, with 12% treated wastewater in 2021 (Figure \@ref(fig:C6-population)). Since 2010, the Vietnamese government has called for the construction of 12 WWTPs to treat about 3 million$m^{3}$ $d^{-1}$ by 2025 (Vietnam Prime Ministerial Decision Jan. 2010, n24/QĐ-TTg). However, there are currently only three WWTPs in operation; one WWTP is under construction, while other WWTPs are still in the fundraising stage. The total estimated cost is over 2 billion USD for the remaining WWTPs. Therefore, we estimate that in 2025 HCMC can treat about 780,000 $m^{3}$ $d^{-1}$, equivalent to 27% of the total wastewater volume of about 2.8 million people. By 2050, WWTPs in HCMC may treat 3,000,000 $m^{3}$ $d^{-1}$, equivalent to 54% of the wastewater of about 16 million people (Figure \@ref(fig:C6-population)).

```{r C6-population, echo=FALSE, out.width='100%', fig.cap='           Population in HCMC from 1979 to 2020, predicted to 2050, and WWTP capacity. Population data from 1979 to 2019 is from GSO Vietnam; prediction population is based on UN World Urbanization Prospects with a growth rate of 2.7\\% in 2020 and 1.3\\% in 2050. Total domestic wastewater discharge is calculated based on water consumption per capita', fig.align='center'}
knitr::include_graphics("figures/C6-population.pdf")
```

### Nutrient loadings from urban canals

Chapter 5 described that phytoplankton blooms could only exist in the urban section (about 50 km of the 200 km length of the river) of the Saigon River. Nutrient limitations upstream and downstream have limited the growth of phytoplankton, and the water quality of these two areas is assessed as good. Therefore, the scenarios for the Saigon River water quality mainly focus on the urban section affected by domestic wastewater of HCMC. More specifically, the quality of wastewater from urban canals is a decisive influence on the water quality of the Saigon River. Correctly determining the quality of the input data applied in the model will ensure the simulation results' meaning.

```{r C6-canal-retention, echo=FALSE, out.width='100%', fig.cap='The nutrients concentration along canal axis based on parabola best-fit equation. The concentration is a function of distance, with the initial concentration at pollutant sources is. The blue points with standard deviation are observed data of 5 sampling sites in Ben Nghe canal in 2014-2017.', fig.align='center'}
knitr::include_graphics("figures/C6-canal-retention.pdf")
```

The data used in this study are mainly from the Vietnam Center of Environmental Monitoring (CEM, Vietnam) and Center Asiatique de Recherche sur l'Eau (CARE). The monitoring program of these two centers has been detailed in previous chapters. In this section, CEM monitoring program is repeated to explain the calculation method to supplement data used for simulations in the past period without sufficient monitoring data.

The monitoring program by CEM, Vietnam, started in the 2000s and was almost completed in 2012 with the addition of many monitoring sites along the river and at the canal systems of HCMC. The data of the 2000s only allowed to assess the water quality at a few monitoring points along the Saigon River while the water quality of the canals was not monitored in routine. Since 2014, the canal monitoring has been improved with more complete sampling sites with monthly measurements. Therefore, it is difficult to quantify the impact of urban discharge with very sparse data in the 2000s. However, 2014-2017 allows one to estimate the retention capacity of the nutrient pollutants in the canals. The use of nutrient retention capacity at canals can be applied to calculate nutrients of past and future scenarios. Figure \@ref(fig:C6-canal-retention) shows the gradual decrease in average concentrations of $\mathrm{NH_4^{+}}$, $\mathrm{PO_4^{3-}}$ from the inner city to the outlet of the canals where the Saigon River is received. This study considers retention capacity in urban canals to transform, remove pollutants in the water column, and retain pollutants in the sediment (burial process). Besides, the dilution capacity of the Saigon River and during flooding tides, nutrient concentrations in the canals gradually decrease over the distance. Therefore, in the absence of data, nutrient concentrations were calculated according to nutrient emission per capita, and the parabola best-fit equation was applied to estimate the concentration of nutrients entering the Saigon River. According to the synthesis of Nguyen et al., 2020, the nutrient emission per capita human in HCMC is about 11 gN/capita $d^{-1}$ and 2.4 gP/capita $d^{-1}$ with water use of 150L $d^{-1}$/capita. This estimate allows estimating the concentration of TN and TP in domestic wastewater in HCMC. However, C-GEM model needs to declare the concentrations of $\mathrm{NH_4^{+}}$ and $\mathrm{PO_4^{3-}}$ at the outlet of urban canals. Based on the Japan International Cooperation Agency (JICA) report in 2020 on the quality of inlet and outlet wastewater of WWTPs in HCMC, the $\mathrm{NH_4^{+}}$/TN and $\mathrm{PO_4^{3-}}$/TP ratios are estimated. Finally, nutrient concentrations at the urban outlet were calculated using the parabola best-fit equation as a function of canal distance (Figure \@ref(fig:C6-canal-retention)). These parabola best-fit equations were set up based on the non-linear least squares to fit the estimated and observed data. To simplify the determination of the nutrient retention capacity of canals, we hypothesized that the parabola equations already include processes such as transformation, removal and burial of nutrients according to the distance from the pollutant source. Since 2008, the sludges in urban canals have been regularly dredged to remove pollutants and improve water quality. Therefore, the parabola best-fit equations are suitable for the condition after 2008. Before 2008, nutrient retention capacity is estimated to be about 50% smaller.

### Description of scenarios

The calibrated and validated C-GEM model was used to evaluate the responses of the Saigon River Estuary under the impact of changes in nutrient loadings from the 2000s to 2050 according to HCMC's development plan and projections of climate change in 2050. The four scenarios in turn were tested by the C-GEM model are presented in Table \@ref(tab:C6-table-scenarios).

These scenarios were based on plans to develop WWTPs in HCMC and forecasts of temperature and sea level rise in the Saigon River (MONRE et al., 2016). The number of WWTPs and their capacity were after the decision n24/QD-TTg approved by the Prime Minister of Vietnam on January 6th, 2010. Under this plan, a total of 12 WWTPs will ensure efficient domestic wastewater treatment for 12 basins in HCMC. Four scenarios (2005, 2015, 2025, 2050) allow predicting the effect of WWTPs on water quality in the Saigon Estuary under the impact of increased waste loads (population growth in HCMC). Therefore, the model parameters in all three scenarios remain the same except for the wastewater source from urban discharge and climate parameters for the year 2050.

```{r C6-canal-flux, echo=FALSE, out.width='100%', fig.cap='The $\\mathrm{NH_4^{+}}$, $\\mathrm{PO_4^{3-}}$ concentrations and fluxes at the outlet of canals. The nutrient concentrations are calculated based on the nutrient emission per capita and then applied the parabola best-fit equation to get the concentration at the outlet of the canals.', fig.align='center'}
knitr::include_graphics("figures/C6-canal-flux.pdf")
```

**Scenario 1:** Without WWTPs in 2005.

This period represents the 2000s when HCMC did not have WWTPs. All domestic wastewater was discharged directly into the canal system in this period. The simulation during this period allows us to assess the pollution level of the Saigon River in extreme conditions. The nutrient concentration in canals was higher than the recommended concentration threshold for surface water quality (QCVN 08: 2015/BTNMT, National technical regulation on surface water quality). {NH_4^{+}}$ and $\mathrm{PO_4^{3-}}$ in 2005 were 7.0 mg$\mathrm{NH_4^{+}}$ $L^{-1}$ and 0.4 mg$\mathrm{PO_4^{3-}}$ $L^{-1}$, respectively. Based on QCVN 08: 2015/BTNMT, the nutrient concentrations were many times higher than the recommended concentrations for domestic usage of 0.3 mg $L^{-1}$ $\mathrm{NH_4^{+}}$ and 0.2 mg $L^{-1}$ $\mathrm{PO_4^{3-}}$, respectively. The input $\mathrm{PO_4^{3-}}$ concentration was about 4.0 mg $L^{-1}$. However, most of the $\mathrm{PO_4^{3-}}$ was retained in the canal before being discharged into the Saigon River. Meanwhile, $\mathrm{NH_4^{+}}$ did not have a high retention efficiency in the canal as $\mathrm{PO_4^{3-}}$, leading to a much higher $\mathrm{NH_4^{+}}$ concentration.

**Scenario 2:** With two WWTPs in 2015.

This scenario represents the present condition of Saigon River Estuary, with the complete data to validate the model. C-GEM uses the average data of the dry and rainy season from 2014 to 2017 to calibrate some parameters in biogeochemical module and evaluate the effectiveness of the model. This period also marks HCMC's efforts when two WWTPs effectively treat about 10% of urban wastewater.

**Scenario 3:** With four WWTPs in 2025 and an increase in the population.

This scenario aims to evaluate the effectiveness of the two new WWTPs according to the current WWTP construction ability. The HCMC government aimed to achieve 12 WWTPs by 2025 to treat 80% of domestic wastewater. However, we believe that only four WWTPs can be achieved in this period because the remaining WWTPs are still in the fundraising stage. In addition, during this period, a temperature increase of 0.5$^{\circ}$C is also applied to represent the effects of global warming on sea temperature in southern Vietnam (Vietnam Ministry of Natural Resources and Environment – MONRE, 2016).

**Scenario 4:** With twelve WWTPs in 2050, increasing population and climate change.

By 2050, the population of the HCMC is forecasted to increase almost three times compared to the population in the 2000s. The remaining WWTPs have been in the capital raising phase, including the socialization of investment from 2018. The proportion of domestic wastewater connected to 12 WWTPs can be 54% in 2050. C-GEM model has also adjusted parameters related to the impact of climate change. Tidal amplitude can be determined to increase by about 20 mm by 2050 [@Bindoff2019] . In addition, MONRE et al., 2016 suggested that the surface temperature of the seawater of Southern Vietnam can be increased to 1.5$^{\circ}$C by 2050 [@MONRE2016]. An increase in temperature can contribute to the increased activity of biogeochemical processes in the estuary.

```{r C6-table-scenarios, echo=F, warning=F}
temp.mat = data.frame(image="")
colnames(temp.mat) <- NULL
temp.mat %>%
  kbl(format="latex",booktabs = T,escape = F,caption="Four scenarios for C-GEM application in Saigon River Estuary") %>%
  kable_paper(full_width = T,latex_options = c("scale_down","striped", "hold_position"), stripe_index =1) %>%
  column_spec(1, image = "excel/CH6-table1-Four scenarios for C-GEM application in Saigon River Estuary.pdf") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

### Model validation

C-GEM is a one-dimensional, reactive transport model. This model was developed to minimize the data required while ensuring the accurate description of estuarine hydrodynamics, salt transport and biogeochemistry [@Volta2014]. Three main modules of C-GEM (geometry, hydrodynamics, and transport - reaction modules) were implemented following the set-up protocol proposed by @Volta2014. The hydrodynamic module was solved by using a finite difference scheme applied along a one-dimensional grid, with a grid size of 2000 m and an integration time step of 300 seconds. Transport-biogeochemical reaction module was solved by the operator-splitting method within a single time step [@Regnier1997]. The application of C-GEM in the Saigon River Estuary required a spin-up of 180 days to reach a steady-state condition. The C-GEM version in this study was adapted from the transient version of @Volta2016 to simulate the water quality along 200 km of Saigon River Estuary.

This study applies a transient version of C-GEM (Chapter 5) to describe the spatial variation between dry and rainy seasons. The description of the model is detailed in Chapter 2 – Methods and Materials.

The current version of C-GEM in this study can simulate the concentrations of eight state variables, namely ammonium ($\mathrm{NH_4^{+}}$), nitrate ($\mathrm{NO_3^{-}}$), phosphate ($\mathrm{PO_4^{3-}}$), total organic carbon (TOC), Silica (DSi), dissolved oxygen (DO), phytoplankton (diatoms and non-diatoms) and Total Suspended Solid (TSS). Seven biogeochemical processes are considered in C-GEM, including oxygen exchange, aerobic degradation (organic carbon mineralization), nitrification, denitrification, primary production, phytoplankton mortality and TSS erosion/deposition. The rate constants of these processes were adjusted in the limit of their values from experimental or literature determination based on 49 estuarine biogeochemical model applications [@Volta2016].

Water quality data to calibrate biogeochemical parameters were obtained from bi-weekly monitoring data of Center of Environment Monitoring (CEM, Vietnam) and Center Asiatique de Recherche sur l'Eau (CARE, Vietnam) from 2014 - 2017. The sampling and analytical methods were described by Nguyen et al. (2020). Most of the observed data were BOD5 but not the TOC. However, there were three stations (at km 88, 130 and 156) where we measured BOD5 and TOC. Therefore, TOC used in the model was the converted BOD5 concentrations using a relationship equation.

```{r C6-model-validation, echo=FALSE, out.width='100%', fig.cap='Comparison of observed and simulated water quality variables along Saigon River Estuary.', fig.align='center'}
knitr::include_graphics("figures/C6-model-validation.pdf")
```

The water quality simulation results were compared with monitoring data of 13 stations along Saigon River Estuary in the dry and rainy seasons 2016 - 2017 (Figure \@ref(fig:C6-model-validation)). Although the current C-GEM model allows for the simulation of eight environmental variables, only four (DO, TOC, $\mathrm{NH_4^{+}}$, $\mathrm{PO_4^{3-}}$) are discussed in this study because they are of particular interest for the municipal WWTPs construction plan in HCMC. In addition, the two monitoring programs of CARE and CEM provided the complete data sets of these variables for the calibration and validation process in C-GEM. Statistical analysis for model performance shows that the simulation results and observation data of 13 monitoring stations have a coefficient correlation ($\mathrm{R^2}$) from 0.6 to 0.9; the percentage bias is less than 20% for most of the variables (except $\mathrm{PO_4^{3-}}$). $\mathrm{PO_4^{3-}}$ simulation has a percent bias of 37% (0.02 mgP$L^{-1}$), which is just equivalent to the experimental error of $\mathrm{PO_4^{3-}}$ (0.01 mgP$L^{-1}$, standard colorimetric methods). These simulation results show that C-GEM model has a good ability to simulate water quality at Saigon River. In addition, C-GEM allows determining the intensity of biogeochemical processes such as nitrification, organic carbon aerobic degradation affecting the concentration of these environmental parameters. Nitrification is the process with the largest consumption of dissolved oxygen in the Saigon River when it receives a high concentration of $\mathrm{NH_4^{+}}$ from urban canal discharge (results are not shown in this study).

## Results and discussion

### Responses of the tropical estuary to nutrient loadings

Figure \@ref(fig:C6-scenarios) depicts the simulation results of water quality at the Estuary Saigon River under the four scenarios driven by the construction of WWTPs in HCMC. In general, the simulations show significant efficiency in the construction of WWTPs, especially for $\mathrm{NH_4^{+}}$ and TOC. According to QCVN 08:2015/BTNMT, river water quality in Vietnam can be classified into four groups according to different concentration ranges. The river water quality is assessed as "very good" (grade A1) suitable for domestic use, aquatic life. "Good" (grade A2) can be used for domestic use but requires appropriate water treatment. "Moderate" (grade B1) and "Bad" (grade B2) are respectively used for irrigation and transport purposes.

In upstream, from km 0 to km 80, simulation results are almost unchanged for all environmental parameters, which is explained by the lack of input changes in this area (see vertical lines in Figure \@ref(fig:C6-scenarios) for the location of external inputs). However, there is a significant decrease of the phytoplankton concentration in upstream which may be almost zero. This difference comes from the increase in temperature in the two scenarios, 2025 and 2050. Phytoplankton mortality in the C-GEM model is a function of temperature. While nutrient limitation prevails upstream (Chapter 5), an increase in temperature leads to phytoplankton depletion due to increased phytoplankton mortality.

In the urban section, from km 80 to km 150, this area is adversely affected by domestic wastewater from HCMC. All scenarios have clear differences in the concentrations of environmental variables in this area, especially DO, $\mathrm{NH_4^{+}}$, Chl-a. Indeed, as forecasted by MONRE, the construction of two more WWTPs will significantly improve water quality in the Saigon River. In addition, the improvement in DO concentration enhances the nitrification efficiency, which further helps to increase the $\mathrm{NH_4^{+}}$ consumption. Unfortunately, $\mathrm{NH_4^{+}}$ concentrations in the two future scenarios will not improve as much as TOC. This can be explained by the WWTPs treatment efficiency of TN of only about 40-50%, while the TOC can be removed at 85% (Metcalf and Eddy/AECOM, 2014). Therefore, to improve the quality of DO and $\mathrm{NH_4^{+}}$, there is a need to improve the efficiency of WWTPs in nutrient treatment, especially for nitrification $\mathrm{NH_4^{+}}$ abatement. Under the 2050 scenario, all environmental variables in the Saigon River would show signs of deterioration. The concentrations of DO could be almost equivalent to the observations in 2015. DO would decrease from 2.4 to 1.4 mgDO.$L^{-1}$ between 2025 and 2050. Although the connection rate with WWTPs reached 61% during this period, nutrients from 16 million inhabitants would cause significant DO consumption by nitrification and aerobic degradation. For the year 2050, three of four considered variables would return to a bad state compared to 2025, although their concentrations would be improved compared to the present. Meanwhile, $\mathrm{PO_4^{3-}}$ concentrations would always remain within the allowed concentration threshold of QCVN 08:2015/BTNMT regardless of the scenarios.

\begin{landscape}
\vfill
```{r C6-scenarios, echo=FALSE, out.width='85%', fig.cap='Simulated results of water quality variable along Saigon River Estuary in three scenarios. According to Vietnamese regulation on surface water quality, the background colors (red, orange, green, blue) are the concentration ranges for water quality assessment (QCVN 08:2015/BTNMT. The trophic status is classified according to the recommendation of Dodds and Welch (2000).', fig.align='center'}
knitr::include_graphics("figures/C6-scenarios.pdf")
```
\end{landscape}

In downstream, the water quality is mainly affected by seawater and tidal variation. The 2050 scenario assesses the impact of an increase in the tidal range of 20 cm. Although all four scenarios were declared with the same boundary conditions, all four scenarios showed no difference in downstream concentrations. The impact of a small tidal variation on water quality is thus weak. Therefore, there is no effect on the urban section and upstream areas by increasing 20 cm of tidal range. In addition, in the two future scenarios, there is an impact of temperature on phytoplankton biomass. According to the monitoring results of Nguyen et al., [-@Nguyen2019], the concentration of Chl-a downstream is always low in the rainy season. Simulation results for two years, 2017-2018, also reached the same conclusion (Chapter 5). However, an increase in temperature of 1.5oC probably made phytoplankton growth superior to short residence time during the rainy season. The concentration of Chl-a in both the urban section and downstream of Saigon River may increase (Figure \@ref(fig:C6-scenarios)).

The 2005 scenario clearly shows the role of estuary in reducing pollutant concentrations because, during this period, 100% of wastewater was discharged into canals and rivers without treatment. As reported in C-GEM model, CH4 concentration in the canal was 7 mg$\mathrm{NH_4^{+}}$ $L^{-1}$. It was reduced by 30% to about 5 mg$\mathrm{NH_4^{+}}$ $L^{-1}$ at Saigon River in the dry season and about 80% in the rainy season. According to Nguyen et al., 2021, nitrification is the most important role in reducing nutrient contamination in the Saigon River. At Pearl River Estuary, a sub-tropical estuary in China, 37% of $\mathrm{NH_4^{+}}$ of external loadings from urban discharge was also removed by the biogeochemical processes at this estuary [@Hu2009].

The response of Saigon River Estuary to different nutrient loadings is seen according to Chl-a concentration. In C-GEM model, concentrations of TOC, $\mathrm{NH_4^{+}}$ and $\mathrm{PO_4^{3-}}$ are all declared and force their concentrations in the urban section (at km 110 and km 130), while the concentration of Chl-a is not bound. Therefore, the change in Chl-a concentration in the Saigon River water column reflects the estuary's obvious response to different nutrient loadings from urban discharge. The change in Chl-a concentration is most surprising in the urban section in 2050 when the trophic status may return to the eutrophic state such as the 2000s (no WWTPs) and 2015 (with 2 WWTPs). Trophic status during the rainy season has also changed from oligotrophic to mesotrophic status. This situation has never been seen in the observation in the rainy seasons from 2015 to 2020. Besides, $\mathrm{PO_4^{3-}}$ is a parameter that always reaches "very good" or "good" status in any scenario. The main cause of this phenomenon is that $\mathrm{PO_4^{3-}}$ has a high capacity to adsorb onto sediment right from urban canals.

### Role of WWTPs in eutrophication management

Successful lessons in the management or recovery of estuaries have been observed in many developed countries, such as in European rivers [@Garnier2018]. Until now, the main solution is the reduction of pollutant concentrations entering estuaries [@Wolanski2014]. For instance, the results of a 40-year survey at the Seine river and estuary (France) have shown the evolution and effectiveness of freshwater quality policies [@Romero2016]. Two decades ago, the Seine estuary used to have summer anoxic episodes, but the extension of WWTPs and agricultural policies has kept the estuary almost always in good condition since 2010. Besides the construction of WWTPs, policies such as the ban of phosphates in household detergents have also reduced $\mathrm{PO_4^{3-}}$ concentrations significantly since the 1990s [@Romero2016]. There are few successes in mitigating eutrophication in urbanized tropical estuaries (e.g., Shenzhen Bay in China, Zhou et al., 2020) by increasing the capacity of WWTPs.

While the assessment of nutrient concentrations in the 2005 scenario showed estuary responses to high nutrient concentrations of urban discharge, the comparison of concentrations in the Saigon River in 2015, 2025 and 2050 scenarios with 2005 shows the role of WWTPs in pollution control. $\mathrm{NH_4^{+}}$ concentrations in urban section in the scenarios 2005, 2015, 2025 and 2050 are 5 mg $L^{-1}$, 2.5 mg $L^{-1}$, 1.5 mg $L^{-1}$ and 2 mg $L^{-1}$, respectively (Figure \@ref(fig:C6-scenarios)). The $\mathrm{NH_4^{+}}$ loadings from canals in the four scenarios are 5670, 4611, 6100 and 12780 kg$\mathrm{NH_4^{+}}$ $d^{-1}$ (Figure \@ref(fig:C6-canal-flux)). The reduction in $\mathrm{NH_4^{+}}$ loading from 2005 to 2015 demonstrates the role of WWTPs and the improvement of canal retention through sludge dredging. Therefore, in the first three scenarios with the support of WWTPs, nutrient loadings from the canals into the Saigon River are not significantly different. Simulated concentrations of TOC and $\mathrm{NH_4^{+}}$ in the 2015 and 2025 scenarios are markedly lower than in 2005. The concentrations of $\mathrm{NH_4^{+}}$ and TOC in 2025 can be reduced respectively by 52% and 37% compared to 2015. Thanks to the increase of the total treatment capacity of WWTPs in HCMC (from 10% to 27%), TOC concentrations can be judged as good or very good in 2025. Likewise, DO under the 2025 scenario changes from a bad to a moderate quality in both seasons.

Reducing the inputs from wastewater in the inner city is the main factor in reducing $\mathrm{NH_4^{+}}$ concentrations in the river. However, in the 2050 scenario, with the population increase from 6 million in 2005 compared to 16 million in 2050, the total nutrient loading increases significantly even concentration of $\mathrm{NH_4^{+}}$ from the canal in 2050 is lower than in previous years. As a result, $\mathrm{NH_4^{+}}$ concentrations have increased again, suggesting that 12 WWTPs in 2050 still do not allow for $\mathrm{NH_4^{+}}$ reduction. The increase in $\mathrm{NH_4^{+}}$ concentration has resulted in DO depletion in the water column because the Saigon River is an intense system with nitrification. Therefore, the efficient collection and treatment of a large amount of wastewater from about 16 million people in 2050 is a major challenge for HCMC. In addition, $\mathrm{NH_4^{+}}$ treatment efficiency is not as high as TOC.

In terms of eutrophication control, WWTPs do not play a key role as there is no significant decrease of Chl-a in the Saigon River in all scenarios. The control of eutrophication requires determining limiting factors for phytoplankton development. For instance, the Loire River, Seine River had Chl-a regularly exceeding 120 $\mu g$ $L^{-1}$ during the 1980s. Chl-a levels were controlled when limiting nutrient loading was controlled, $\mathrm{PO_4^{3-}}$ in the case of the Seine and Loire River with concentrations of about 0.05 mg$\mathrm{PO_4^{3-}}$ $L^{-1}$ [@Garnier2018]. According to @Nguyen2019, $\mathrm{PO_4^{3-}}$ is also limiting for phytoplankton growth in the Saigon River. Although the concentration of $\mathrm{PO_4^{3-}}$ has decreased many times from inner urban canals before entering the Saigon River, its concentration are still about 0.4 mg$\mathrm{PO_4^{3-}}$ $L^{-1}$ (in the 2000s), 0.2 mg$\mathrm{PO_4^{3-}}$ $L^{-1}$ (in 2010s), 0.15 mg$\mathrm{PO_4^{3-}}$ $L^{-1}$ (in 2050). The oxygen concentration has improved compared to the 2000s, and $\mathrm{PO_4^{3-}}$ is within the recommended concentration according to QCVN 08:2015/BTNMT. The risk of eutrophication by 2050 may occur if $\mathrm{PO_4^{3-}}$ concentrations are not controlled. While the increase in temperature and the growing intensity of phytoplankton cannot be controlled, the reduction of pollutant sources of $\mathrm{PO_4^{3-}}$ is crucial. Therefore, the role of WWTPs or alternative solutions in mitigating eutrophication under the 2050 scenario is extremely important.

## Conclusion

The application of the C-GEM model at the Saigon Estuary well reproduces the water quality parameters such as DO, $\mathrm{NH_4^{+}}$, $\mathrm{PO_4^{3-}}$ and TOC. After calibration and validation based on the Saigon River observed data in the dry seasons 2014 - 2017, C-GEM allows evaluating the changes of water quality in the Saigon River under various conditions of pollution input in dry and rainy season in 2005, 2015, 2025, and 2050. Model results for two future scenarios show a marked effect in improving river water quality by WWTPs construction, especially under scenario 2025. This result shows that the construction of WWTPs is essential for the sustainable development of HCMC. However, population growth from about 6 million inhabitants in 2005 to 16 million in 2050 can make the water quality return to bad condition, as in 2005 and 2015. Although the construction of new WWTPs could treat around 60% of the population in Ho Chi Minh City, the water quality in the Saigon River would not meet the water quality standards on surface water quality QCVN 08:2015/BTNMT for $\mathrm{NH_4^{+}}$. Under the 2050 scenario with climate change factors, the increase in temperature stimulates the growth of phytoplankton, the ineffectiveness of removing $\mathrm{PO_4^{3-}}$ as a limiting factor may lead to the return of eutrophic status in the Saigon River. To solve this problem, improvement of nutrient processing technology in WWTPs and implementing alternative urban wastewater management (Nature-based Solutions) are probably the effective solution for eliminating oxygen shortages or nutrient pollution (e.g., $\mathrm{NH_4^{+}}$) phytoplankton bloom in the urban section of the Saigon River.

## Conclusion of Chapter 6

The efforts in applying the C-GEM model and calibration, validation in the dry season (Chapter 4) and expanding the seasonal assessment (Chapter 5) have ensured the applicability of C-GEM in the evaluation of scenarios at Saigon River Estuary. This chapter has assessed the urgent need to reduce pressure from pollution sources of Ho Chi Minh megacity on water quality and eutrophication in the Saigon River. With the help of an estuarine biogeochemical model (C-GEM), factors for improving water quality and minimizing eutrophication were identified. The extended application of the model will provide important information in ensuring the sustainable development of urbanized tropical estuaries.
